---
title: "CBC data analysis"
author: "Zehao Wu"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Libraries

```{r libraries}

library(tidyverse)
library(readxl)
library(writexl)
library(janitor)
library(factoextra)
library(skimr)
library(naniar)
library(ggplot2)
library(ggpubr)
library(emmeans)
library(lme4)
library(lmerTest)
library(DHARMa)
library(scales)
library(haven)
library(broom.mixed)
library(readr)

```

# Setup: Colors, Paths, and Vectors

```{r setup_colors}

color_black      <- "#000000"
color_mustard    <- "#E69F00"
color_light_blue <- "#56B4E9"
color_green      <- "#009E73"
color_yellow     <- "#F0E442"
color_dark_blue  <- "#0072B2"
color_orange     <- "#D55E00"
color_purple     <- "#CC79A7"

ibm_yellow <- "#ffb000"
ibm_orange <- "#fe6100"
ibm_reddish_purple <- "#dc267f"
ibm_purple <- "#785ef0"
ibm_blue <- "#648fff"

```

```{r setup_paths}

base_path <- "/Users/zehao/Documents/Research Projects/Graham Group Thesis Research (Summer 2025)"
data_path <- file.path(base_path, "Metadata")
clean_data_path <- file.path(base_path, "Cleaned data")
figure_path <- file.path(base_path, "Figures")
cbc_figure_path <- file.path(figure_path, "CBC Figures")
analysis_directory <- file.path(base_path, "Analysis Results")

dir.create(cbc_figure_path, 
           showWarnings = FALSE, 
           recursive = TRUE)

```

```{r setup_vectors}

cbc_number <- c("neu_number",
                "lym_number",
                "mon_number",
                "eos_number",
                "bas_number")

cbc_percent <- c("neu_percent",
                 "lym_percent",
                 "mon_percent",
                 "eos_percent",
                 "bas_percent")

```

# Defining functions

```{r helper_functions}

data_saver <- function(data_name,
                       path = "") {
  write_csv(x = data_name,
            file = path)
}

plot_saver <- function(filename,
                       plot,
                       path = cbc_figure_path,
                       width = 6,
                       height = 4,
                       dpi = 300) {
  ggsave(filename = filename,
         plot = plot,
         path = path,
         width = width,
         height = height,
         units = "in",
         dpi = dpi)
}

cbc_summarizer <- function(data,
                           group_name) {
  data %>%
    summarise(neu_percent = mean(neu_percent, na.rm = TRUE) / 100,
              lym_percent = mean(lym_percent, na.rm = TRUE) / 100,
              mon_percent = mean(mon_percent, na.rm = TRUE) / 100,
              eos_percent = mean(eos_percent, na.rm = TRUE) / 100,
              bas_percent = mean(bas_percent, na.rm = TRUE) / 100) %>%
    mutate(group = group_name) %>%
    pivot_longer(cols = neu_percent:bas_percent,
                 names_to = "cell_type",
                 values_to = "cbc_percent")
}

emm_plotter_long <- function(emm_df,
                              raw_data,
                              y_col,
                              y_label,
                              title = NULL,
                              floor_zero = FALSE) {
  if (floor_zero) {
    emm_df <- emm_df %>% mutate(lower.CL = pmax(lower.CL, 0))
  }

  ggplot(emm_df, aes(x = phase,
                     y = response,
                     color = t_muris_dose,
                     fill = t_muris_dose,
                     group = t_muris_dose)) +
    geom_point(size = 3,
               shape = 21,
               color = "black",
               position = position_dodge(width = 0.5)) +
    geom_errorbar(aes(ymin = lower.CL,
                      ymax = upper.CL),
                  width = 0.2,
                  alpha = 0.7,
                  position = position_dodge(width = 0.5)) +
    geom_point(data = raw_data,
               aes(x = phase,
                   y = .data[[y_col]],
                   fill = t_muris_dose,
                   color = t_muris_dose),
               position = position_jitterdodge(jitter.width = 0.1,
                                               dodge.width = 0.5),
               alpha = 0.3,
               size = 1.2,
               inherit.aes = FALSE) +
    geom_vline(xintercept = 1.5,
               linetype = "dashed",
               color = "grey40",
               alpha = 0.6) +
    annotate("text",
             x = 1.5,
             y = Inf,
             label = "Infection",
             vjust = 1.5,
             hjust = -0.1,
             size = 3,
             fontface = "italic",
             alpha = 0.6) +
    facet_grid(lab_or_rw ~ strain) +
    scale_fill_manual(values = c("Zero" = color_dark_blue,
                                 "Low"  = color_mustard,
                                 "High" = color_orange),
                      name = expression(bolditalic("T. muris") ~ bold("Dosage"))) +
    scale_color_manual(values = c("Zero" = color_dark_blue,
                                  "Low"  = color_mustard,
                                  "High" = color_orange),
                       name = expression(bolditalic("T. muris") ~ bold("Dosage"))) +
    scale_y_continuous(name = y_label,
                       expand = expansion(mult = c(0.05, 0.3))) +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "white",
                                          color = "black",
                                          linewidth = 0.8),
          strip.text = element_text(size = 10,
                                    face = "bold"),
          panel.border = element_rect(color = "black",
                                      fill = NA,
                                      linewidth = 0.8),
          panel.grid = element_blank(),
          plot.title = element_text(hjust = 0.5,
                                    face = "bold"),
          axis.title.x = element_text(face = "bold"),
          axis.title.y = element_text(face = "bold")) +
    xlab("Phase") +
    ggtitle(title)
}

cbc_plotter <- function(cbc_data,
                        cbc_error,
                        y_label,
                        title = NULL) {
  ggplot(cbc_data, aes(x = phase,
                       y = number,
                       fill = t_muris_dose,
                       color = t_muris_dose,
                       group = t_muris_dose)) +
  geom_point(position = position_dodge(width = 0.5),
             size = 1.5,
             alpha = 0.3) +
  geom_line(data = cbc_error,
            aes(y = mean, group = t_muris_dose),
            linewidth = 0.5,
            alpha = 0.6,
            position = position_dodge(width = 0.5)) +
  geom_point(data = cbc_error,
             aes(y = mean),
             position = position_dodge(width = 0.5),
             size = 2.5,
             alpha = 1) +
  geom_errorbar(data = cbc_error,
                aes(y = mean,
                    ymin = mean - se,
                    ymax = mean + se),
                position = position_dodge(width = 0.5),
                width = 0.15,
                alpha = 0.6) +
  geom_vline(xintercept = 1.6,
             linetype = "dashed",
             color = "grey40",
             alpha = 0.6) +
  annotate("text",
           x = 1.6,
           y = Inf,
           label = "Infection",
           vjust = 1.5,
           hjust = -0.1,
           size = 3,
           fontface = "italic",
           alpha = 0.6) +
  facet_grid(lab_or_rw ~ strain) +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "Low"  = color_mustard,
                               "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dosage"))) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "Low"  = color_mustard,
                                "High" = color_orange),
                     name = expression(bolditalic("T. muris") ~ bold("Dosage"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5,
                                  face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold")) +
  xlab("Phase") +
  ylab(y_label) +
  ggtitle(title)
}

```

# Loading Data

```{r load_data}

# Mouse CBC data
cbc_data_1 <- read_excel(file.path(data_path, 
                                   "DCV_25SF.W-3thruNSF1trapout.XLSX"))
cbc_data_2 <- read_excel(file.path(data_path,
                                   "DCV_25SF.W3_NSF2.XLSX"))
meta_data   <- read_csv(file.path(data_path, 
                                  "clean_meta_data.csv"))

# CDC NHANES human data
nhanes_data_2017_18 <- read_xpt(file.path(data_path, 
                                          "CBC_J_NHANES_2017_2018.xpt"))
nhanes_data_2021_23 <- read_xpt(file.path(data_path, 
                                          "CBC_L_NHANES_2021_2023.xpt"))

```

# Cleaning Mouse CBC Data

```{r clean_mouse_cbc}

clean_cbc_data_1 <- cbc_data_1 %>%
  clean_names() %>%
  filter(id != "BG",
         !id %in% c("25SF.W5.1330.0.1",
                    "25SF.W3.1311.0",
                    "25SF.17.1380.-3",
                    "25SF.W5.1330.0",
                    "25SF.W5.1330.2")) %>%
  separate(col = id,
           into = c("experiment",
                    "cage_wedge",
                    "mouse_id",
                    "week",
                    "repeat"),
           sep = "\\.",
           fill = "right",
           extra = "merge",
           remove = FALSE) %>%
  select(mouse_id,
         all_of(cbc_number),
         neu_percent_2,
         lym_percent_2,
         mon_percent,
         eos_percent_2,
         bas_percent_2,
         cage_wedge,
         week,
         `repeat`) %>%
  rename(neu_percent = neu_percent_2,
         lym_percent = lym_percent_2,
         eos_percent = eos_percent_2,
         bas_percent = bas_percent_2) %>%
  mutate(week = as.integer(week),
         cage_wedge = factor(cage_wedge),
         mouse_id = as.character(mouse_id),
         across(all_of(c(cbc_number, cbc_percent)), ~ as.numeric(.)))

clean_cbc_data_2 <- cbc_data_2 %>%
  clean_names() %>%
  filter(id != "BG") %>%
  separate(col = id,
           into = c("experiment",
                    "cage_wedge",
                    "mouse_id",
                    "week",
                    "repeat"),
           sep = "\\.",
           fill = "right",
           extra = "merge",
           remove = FALSE) %>%
  select(mouse_id,
         all_of(cbc_number),
         neu_percent_2,
         lym_percent_2,
         mon_percent,
         eos_percent_2,
         bas_percent_2,
         cage_wedge,
         week,
         `repeat`) %>%
  rename(neu_percent = neu_percent_2,
         lym_percent = lym_percent_2,
         eos_percent = eos_percent_2,
         bas_percent = bas_percent_2) %>%
  mutate(week = as.integer(week),
         cage_wedge = factor(cage_wedge),
         mouse_id = as.character(mouse_id),
         across(all_of(c(cbc_number, cbc_percent)), ~ as.numeric(.)))

```

# Combining Mouse CBC Data and Adding Metadata

```{r combine_and_join}

cleaned_meta_data <- meta_data %>%
  mutate(mouse_id = as.character(mouse_id))

cleaned_combined_cbc_data <- bind_rows(clean_cbc_data_1,
                                       clean_cbc_data_2) %>%
  select(-`repeat`) %>%
  left_join(cleaned_meta_data, by = "mouse_id") %>%
  mutate(cage_wedge = tolower(cage_wedge)) %>%
  mutate(week = as.integer(week),
         strain = factor(strain),
         lab_or_rw = factor(lab_or_rw),
         t_muris_dose = factor(t_muris_dose),
         mouse_id = factor(mouse_id)) %>%
  mutate(strain = relevel(factor(strain), ref = "C57BL/6J"),
         t_muris_dose = relevel(factor(t_muris_dose), ref = "Zero")) %>%
  filter(!is.na(lab_or_rw)) %>%
  mutate(phase = case_when(
    week %in% c(-4, -3) ~ "Pre-RW",
    week == 0           ~ "During RW",
    week %in% c(3, 4)   ~ "Post-RW"
  ) %>% factor(levels = c("Pre-RW", "During RW", "Post-RW")))

cbc_endpoint <- cleaned_combined_cbc_data %>%
  filter(phase == "Post-RW")

# [132, 193] was removed because it has different naming. Check how to fix this issue soon.

```

```{r save_mouse_cbc}

data_saver(data_name = cleaned_combined_cbc_data,
           path = file.path(clean_data_path, "cleaned_combined_cbc_data.csv"))

```

# Cleaning NHANES Human Data

```{r clean_nhanes}

cleaned_nhanes_data_2017_18 <- nhanes_data_2017_18 %>%
  clean_names() %>%
  rename(respondent_id = seqn,
         lym_percent = lbxlypct,
         mon_percent = lbxmopct,
         neu_percent = lbxnepct,
         bas_percent = lbxbapct,
         eos_percent = lbxeopct,
         lym_number  = lbdlymno,
         neu_number  = lbdneno,
         mon_number  = lbdmono,
         bas_number  = lbdbano,
         eos_number  = lbdeono) %>%
  select(respondent_id,
         all_of(cbc_number),
         all_of(cbc_percent)) %>%
  drop_na() %>%
  mutate(respondent_id = as.character(respondent_id),
         across(all_of(c(cbc_number, cbc_percent)), ~ as.numeric(.)))

cleaned_nhanes_data_2021_23 <- nhanes_data_2021_23 %>%
  clean_names() %>%
  rename(respondent_id = seqn,
         lym_percent = lbxlypct,
         mon_percent = lbxmopct,
         neu_percent = lbxnepct,
         bas_percent = lbxbapct,
         eos_percent = lbxeopct,
         lym_number  = lbdlymno,
         neu_number  = lbdneno,
         mon_number  = lbdmono,
         bas_number  = lbdbano,
         eos_number  = lbdeono) %>%
  select(respondent_id,
         all_of(cbc_number),
         all_of(cbc_percent)) %>%
  drop_na() %>%
  mutate(respondent_id = as.character(respondent_id),
         across(all_of(c(cbc_number, cbc_percent)), ~ as.numeric(.)))

data_saver(data_name = cleaned_nhanes_data_2017_18,
           path = file.path(clean_data_path, "cleaned_nhanes_data_2017_18.csv"))

data_saver(data_name = cleaned_nhanes_data_2021_23,
           path = file.path(clean_data_path, "cleaned_nhanes_data_2021_23.csv"))

```

# Long-form Pivot (Mouse CBC)

```{r pivot_long}

cleaned_combined_cbc_data_long <- cleaned_combined_cbc_data %>%
  pivot_longer(cols = c(all_of(cbc_percent),
                        all_of(cbc_number)),
               names_to = c("cell_type",
                            ".value"),
               names_pattern = "(.*)_(number|percent)") %>%
  filter(!is.na(lab_or_rw))

data_saver(data_name = cleaned_combined_cbc_data_long,
           path = file.path(clean_data_path, "cleaned_combined_cbc_data_long.csv"))

```

# Composition Plot (Human vs. Mouse)

```{r composition_plot_data}

human_17_18 <- cbc_summarizer(cleaned_nhanes_data_2017_18, "Human_2017_18") %>%
  mutate(strain = "Human")

human_21_23 <- cbc_summarizer(cleaned_nhanes_data_2021_23, "Human_2021_23") %>%
  mutate(strain = "Human")

mean_mouse <- cbc_endpoint %>%
  group_by(strain, t_muris_dose, lab_or_rw) %>%
  summarise(neu_percent = mean(neu_percent, na.rm = TRUE) / 100,
            lym_percent = mean(lym_percent, na.rm = TRUE) / 100,
            mon_percent = mean(mon_percent, na.rm = TRUE) / 100,
            eos_percent = mean(eos_percent, na.rm = TRUE) / 100,
            bas_percent = mean(bas_percent, na.rm = TRUE) / 100,
            .groups = "drop") %>%
  mutate(group = paste(strain, t_muris_dose, lab_or_rw)) %>%
  pivot_longer(cols = neu_percent:bas_percent,
               names_to = "cell_type",
               values_to = "cbc_percent")

bar_plot_input <- bind_rows(human_17_18,
                            human_21_23,
                            mean_mouse) %>%
  mutate(group = factor(group,
                        levels = c(
                          "Human_2017_18",
                          "Human_2021_23",
                          "C57BL/6J Zero Lab",
                          "B10.BR Zero Lab",
                          "C57BL/6J Low Lab",
                          "B10.BR Low Lab",
                          "C57BL/6J High Lab",
                          "B10.BR High Lab",
                          "C57BL/6J Zero RW",
                          "B10.BR Zero RW",
                          "C57BL/6J Low RW",
                          "B10.BR Low RW",
                          "C57BL/6J High RW",
                          "B10.BR High RW")),
         cell_type = factor(cell_type,
                            levels = c("neu_percent",
                                       "lym_percent",
                                       "mon_percent",
                                       "eos_percent",
                                       "bas_percent")))

data_saver(data_name = bar_plot_input,
           path = file.path(clean_data_path, "input_data_cbc_composition.csv"))

```

```{r composition_plot}

comp_plot <- ggplot(bar_plot_input,
                    aes(x = group,
                        y = cbc_percent,
                        fill = cell_type)) +
  geom_col(position = "stack",
           color = "black",
           linewidth = 0.5,
           alpha = 0.85) +
  scale_fill_manual(values = c("neu_percent" = color_dark_blue,
                               "lym_percent" = color_orange,
                               "mon_percent" = color_green,
                               "eos_percent" = color_mustard,
                               "bas_percent" = color_purple),
                    labels = c("neu_percent" = "Neutrophils",
                               "lym_percent" = "Lymphocytes",
                               "mon_percent" = "Monocytes",
                               "eos_percent" = "Eosinophils",
                               "bas_percent" = "Basophils"),
                    name = "Immune Cell Type") +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  scale_x_discrete(labels = c("Human_2017_18" = "Human 2017-18",
                               "Human_2021_23" = "Human 2021-23")) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        plot.title = element_text(hjust = 0.5,
                                  face = "bold"),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   size = 8)) +
  xlab("Population") +
  ylab("CBC Composition")

comp_plot

plot_saver("cbc_composition_plot.png",
           plot = comp_plot)

```

# Trajectory Plots (Longitudinal)

```{r lym_trajectory}

lym_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(cell_type == "lym")

standard_error_lym <- lym_plot_data %>%
  group_by(phase, t_muris_dose, strain, lab_or_rw) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n),
            .groups = "drop")

lym_plot <- cbc_plotter(lym_plot_data,
                        standard_error_lym,
                        expression(bold("Lymphocyte Count (10"^3~"cells/"*mu*"L)")),
                        "Lymphocyte Count Over Time")

lym_plot

plot_saver("lym_number.png", plot = lym_plot)

```

```{r neu_trajectory}

neu_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(cell_type == "neu")

standard_error_neu <- neu_plot_data %>%
  group_by(phase, t_muris_dose, strain, lab_or_rw) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n),
            .groups = "drop")

neu_plot <- cbc_plotter(neu_plot_data,
                        standard_error_neu,
                        expression(bold("Neutrophil Count (10"^3~"cells/"*mu*"L)")),
                        "Neutrophil Count Over Time")

neu_plot

plot_saver("neu_number.png", plot = neu_plot)

```

```{r mon_trajectory}

mon_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(cell_type == "mon")

standard_error_mon <- mon_plot_data %>%
  group_by(phase, t_muris_dose, strain, lab_or_rw) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n),
            .groups = "drop")

mon_plot <- cbc_plotter(mon_plot_data,
                        standard_error_mon,
                        expression(bold("Monocyte Count (10"^3~"cells/"*mu*"L)")),
                        "Monocyte Count Over Time")

mon_plot

plot_saver("mon_number.png", plot = mon_plot)

```

```{r eos_trajectory}

eos_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(cell_type == "eos")

standard_error_eos <- eos_plot_data %>%
  group_by(phase, t_muris_dose, strain, lab_or_rw) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n),
            .groups = "drop")

eos_plot <- cbc_plotter(eos_plot_data,
                        standard_error_eos,
                        expression(bold("Eosinophil Count (10"^3~"cells/"*mu*"L)")),
                        "Eosinophil Count Over Time")

eos_plot

plot_saver("eos_number.png", plot = eos_plot)

```

```{r bas_trajectory}

bas_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(cell_type == "bas")

standard_error_bas <- bas_plot_data %>%
  group_by(phase, t_muris_dose, strain, lab_or_rw) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n),
            .groups = "drop")

bas_plot <- cbc_plotter(bas_plot_data,
                        standard_error_bas,
                        expression(bold("Basophil Count (10"^3~"cells/"*mu*"L)")),
                        "Basophil Count Over Time")

bas_plot

plot_saver("bas_number.png", plot = bas_plot)

```

# Longitudinal Models and EMM Plots (Per Strain)

## Neutrophil

```{r neu_log_models}

neu_lmer_log_b6 <- lmer(log(neu_number) ~ lab_or_rw * phase * t_muris_dose +
                          (1 | mouse_id) + (1 | cage_wedge) + (1 | cohort),
                        data = filter(cleaned_combined_cbc_data, strain == "C57BL/6J"))

print(summary(neu_lmer_log_b6), correlation = FALSE)

neu_b6_sim_res <- simulateResiduals(fittedModel = neu_lmer_log_b6, n = 1000)
plot(neu_b6_sim_res)

emm_neu_b6 <- emmeans(neu_lmer_log_b6,
                      specs = c("t_muris_dose", "phase"),
                      by = "lab_or_rw",
                      type = "response")
emm_neu_b6
pairs(emm_neu_b6)

neu_lmer_log_b10 <- lmer(log(neu_number) ~ lab_or_rw * phase * t_muris_dose +
                           (1 | mouse_id) + (1 | cage_wedge) + (1 | cohort),
                         data = filter(cleaned_combined_cbc_data, strain == "B10.BR"))

print(summary(neu_lmer_log_b10), correlation = FALSE)

neu_b10_sim_res <- simulateResiduals(fittedModel = neu_lmer_log_b10, n = 1000)
plot(neu_b10_sim_res)

emm_neu_b10 <- emmeans(neu_lmer_log_b10,
                       specs = c("t_muris_dose", "phase"),
                       by = "lab_or_rw",
                       type = "response")
emm_neu_b10
pairs(emm_neu_b10)

```

```{r neu_emm_plot}

emm_neu_combined <- bind_rows(
  as.data.frame(emm_neu_b6)  %>% mutate(strain = "C57BL/6J"),
  as.data.frame(emm_neu_b10) %>% mutate(strain = "B10.BR")
) %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "Low", "High")),
         strain = factor(strain, levels = c("C57BL/6J", "B10.BR")))

emm_neu_plot <- emm_plotter_long(
  emm_df   = emm_neu_combined,
  raw_data = cleaned_combined_cbc_data,
  y_col    = "neu_number",
  y_label  = expression(bold("Neutrophil Count (10"^3~"cells/"*mu*"L)")),
  title    = "Estimated Marginal Means of Neutrophil Count with 95% CI"
)

emm_neu_plot

plot_saver("emm_neu_number.png", emm_neu_plot)

```

## Lymphocyte

```{r lym_log_models}

lym_lmer_log_b6 <- lmer(log(lym_number) ~ lab_or_rw * phase * t_muris_dose +
                          (1 | mouse_id) + (1 | cage_wedge) + (1 | cohort),
                        data = filter(cleaned_combined_cbc_data, strain == "C57BL/6J"))

print(summary(lym_lmer_log_b6), correlation = FALSE)

lym_b6_sim_res <- simulateResiduals(fittedModel = lym_lmer_log_b6, n = 1000)
plot(lym_b6_sim_res)

emm_lym_b6 <- emmeans(lym_lmer_log_b6,
                      specs = c("t_muris_dose", "phase"),
                      by = "lab_or_rw",
                      type = "response")
emm_lym_b6
pairs(emm_lym_b6)

lym_lmer_log_b10 <- lmer(log(lym_number) ~ lab_or_rw * phase * t_muris_dose +
                           (1 | mouse_id) + (1 | cage_wedge) + (1 | cohort),
                         data = filter(cleaned_combined_cbc_data, strain == "B10.BR"))

print(summary(lym_lmer_log_b10), correlation = FALSE)

lym_b10_sim_res <- simulateResiduals(fittedModel = lym_lmer_log_b10, n = 1000)
plot(lym_b10_sim_res)

emm_lym_b10 <- emmeans(lym_lmer_log_b10,
                       specs = c("t_muris_dose", "phase"),
                       by = "lab_or_rw",
                       type = "response")
emm_lym_b10
pairs(emm_lym_b10)

```

```{r lym_emm_plot}

emm_lym_combined <- bind_rows(
  as.data.frame(emm_lym_b6)  %>% mutate(strain = "C57BL/6J"),
  as.data.frame(emm_lym_b10) %>% mutate(strain = "B10.BR")
) %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "Low", "High")),
         strain = factor(strain, levels = c("C57BL/6J", "B10.BR")))

emm_lym_plot <- emm_plotter_long(
  emm_df   = emm_lym_combined,
  raw_data = cleaned_combined_cbc_data,
  y_col    = "lym_number",
  y_label  = expression(bold("Lymphocyte Count (10"^3~"cells/"*mu*"L)")),
  title    = "Estimated Marginal Means of Lymphocyte Count with 95% CI"
)

emm_lym_plot

plot_saver("emm_lym_number.png", emm_lym_plot)

```

## Monocyte

```{r mon_log_models}

mon_lmer_log_b6 <- lmer(log(mon_number) ~ lab_or_rw * phase * t_muris_dose +
                          (1 | mouse_id) + (1 | cage_wedge) + (1 | cohort),
                        data = filter(cleaned_combined_cbc_data, strain == "C57BL/6J"))

print(summary(mon_lmer_log_b6), correlation = FALSE)

mon_b6_sim_res <- simulateResiduals(fittedModel = mon_lmer_log_b6, n = 1000)
plot(mon_b6_sim_res)

emm_mon_b6 <- emmeans(mon_lmer_log_b6,
                      specs = c("t_muris_dose", "phase"),
                      by = "lab_or_rw",
                      type = "response")
emm_mon_b6
pairs(emm_mon_b6)

mon_lmer_log_b10 <- lmer(log(mon_number) ~ lab_or_rw * phase * t_muris_dose +
                           (1 | mouse_id) + (1 | cage_wedge) + (1 | cohort),
                         data = filter(cleaned_combined_cbc_data, strain == "B10.BR"))

print(summary(mon_lmer_log_b10), correlation = FALSE)

mon_b10_sim_res <- simulateResiduals(fittedModel = mon_lmer_log_b10, n = 1000)
plot(mon_b10_sim_res)

emm_mon_b10 <- emmeans(mon_lmer_log_b10,
                       specs = c("t_muris_dose", "phase"),
                       by = "lab_or_rw",
                       type = "response")
emm_mon_b10
pairs(emm_mon_b10)

```

```{r mon_emm_plot}

emm_mon_combined <- bind_rows(
  as.data.frame(emm_mon_b6)  %>% mutate(strain = "C57BL/6J"),
  as.data.frame(emm_mon_b10) %>% mutate(strain = "B10.BR")
) %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "Low", "High")),
         strain = factor(strain, levels = c("C57BL/6J", "B10.BR")))

emm_mon_plot <- emm_plotter_long(
  emm_df   = emm_mon_combined,
  raw_data = cleaned_combined_cbc_data,
  y_col    = "mon_number",
  y_label  = expression(bold("Monocyte Count (10"^3~"cells/"*mu*"L)")),
  title    = "Estimated Marginal Means of Monocyte Count with 95% CI"
)

emm_mon_plot

plot_saver("emm_mon_number.png", emm_mon_plot)

```

## Eosinophil

```{r eos_log_models}

eos_lmer_log_b6 <- lmer(log1p(eos_number) ~ lab_or_rw * phase * t_muris_dose +
                          (1 | mouse_id) + (1 | cage_wedge) + (1 | cohort),
                        data = filter(cleaned_combined_cbc_data, strain == "C57BL/6J"))

print(summary(eos_lmer_log_b6), correlation = FALSE)

eos_b6_sim_res <- simulateResiduals(fittedModel = eos_lmer_log_b6, n = 1000)
plot(eos_b6_sim_res)

emm_eos_b6 <- emmeans(eos_lmer_log_b6,
                      specs = c("t_muris_dose", "phase"),
                      by = "lab_or_rw",
                      type = "response")
emm_eos_b6
pairs(emm_eos_b6)

eos_lmer_log_b10 <- lmer(log1p(eos_number) ~ lab_or_rw * phase * t_muris_dose +
                           (1 | mouse_id) + (1 | cage_wedge) + (1 | cohort),
                         data = filter(cleaned_combined_cbc_data, strain == "B10.BR"))

print(summary(eos_lmer_log_b10), correlation = FALSE)

eos_b10_sim_res <- simulateResiduals(fittedModel = eos_lmer_log_b10, n = 1000)
plot(eos_b10_sim_res)

emm_eos_b10 <- emmeans(eos_lmer_log_b10,
                       specs = c("t_muris_dose", "phase"),
                       by = "lab_or_rw",
                       type = "response")
emm_eos_b10
pairs(emm_eos_b10)

```

```{r eos_emm_plot}

emm_eos_combined <- bind_rows(
  as.data.frame(emm_eos_b6)  %>% mutate(strain = "C57BL/6J"),
  as.data.frame(emm_eos_b10) %>% mutate(strain = "B10.BR")
) %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "Low", "High")),
         strain = factor(strain, levels = c("C57BL/6J", "B10.BR")))

emm_eos_plot <- emm_plotter_long(
  emm_df     = emm_eos_combined,
  raw_data   = cleaned_combined_cbc_data,
  y_col      = "eos_number",
  y_label    = expression(bold("Eosinophil Count (10"^3~"cells/"*mu*"L)")),
  title      = "Estimated Marginal Means of Eosinophil Count with 95% CI",
  floor_zero = TRUE
)

emm_eos_plot

plot_saver("emm_eos_number.png", emm_eos_plot)

```

## Basophil

```{r bas_log_models}

bas_lmer_log_b6 <- lmer(log1p(bas_number) ~ lab_or_rw * phase * t_muris_dose +
                          (1 | mouse_id) + (1 | cage_wedge) + (1 | cohort),
                        data = filter(cleaned_combined_cbc_data, strain == "C57BL/6J"))

print(summary(bas_lmer_log_b6), correlation = FALSE)

bas_b6_sim_res <- simulateResiduals(fittedModel = bas_lmer_log_b6, n = 1000)
plot(bas_b6_sim_res)

emm_bas_b6 <- emmeans(bas_lmer_log_b6,
                      specs = c("t_muris_dose", "phase"),
                      by = "lab_or_rw",
                      type = "response")
emm_bas_b6
pairs(emm_bas_b6)

bas_lmer_log_b10 <- lmer(log1p(bas_number) ~ lab_or_rw * phase * t_muris_dose +
                           (1 | mouse_id) + (1 | cage_wedge) + (1 | cohort),
                         data = filter(cleaned_combined_cbc_data, strain == "B10.BR"))

print(summary(bas_lmer_log_b10), correlation = FALSE)

bas_b10_sim_res <- simulateResiduals(fittedModel = bas_lmer_log_b10, n = 1000)
plot(bas_b10_sim_res)

emm_bas_b10 <- emmeans(bas_lmer_log_b10,
                       specs = c("t_muris_dose", "phase"),
                       by = "lab_or_rw",
                       type = "response")
emm_bas_b10
pairs(emm_bas_b10)

```

```{r bas_emm_plot}

emm_bas_combined <- bind_rows(
  as.data.frame(emm_bas_b6)  %>% mutate(strain = "C57BL/6J"),
  as.data.frame(emm_bas_b10) %>% mutate(strain = "B10.BR")
) %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "Low", "High")),
         strain = factor(strain, levels = c("C57BL/6J", "B10.BR")))

emm_bas_plot <- emm_plotter_long(
  emm_df     = emm_bas_combined,
  raw_data   = cleaned_combined_cbc_data,
  y_col      = "bas_number",
  y_label    = expression(bold("Basophil Count (10"^3~"cells/"*mu*"L)")),
  title      = "Estimated Marginal Means of Basophil Count with 95% CI",
  floor_zero = TRUE
)

emm_bas_plot

plot_saver("emm_bas_number.png", emm_bas_plot)

```
