---
title: "CBC data analysis"
author: "Zehao Wu"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r}

library(tidyverse)
library(readxl)
library(writexl)
library(janitor)
library(factoextra)
library(skimr)
library(naniar)
library(ggplot2)
library(ggpubr)
library(emmeans)
library(lme4)
library(lmerTest)
library(DHARMa)
library(scales)
library(haven)
library(broom.mixed)
library(readr)

```

# Setting the colors (Okabe-Ito palette and IBM palette)

```{r setup_colors}

color_black <- "#000000"
color_mustard <- "#E69F00"
color_light_blue <- "#56B4E9"
color_green <- "#009E73"
color_yellow <- "#F0E442"
color_dark_blue <- "#0072B2"
color_orange <- "#D55E00"
color_purple <- "#CC79A7"

ibm_yellow <- "#ffb000"
ibm_orange <- "#fe6100"
ibm_reddish_purple <- "#dc267f"
ibm_purple <- "#785ef0"
ibm_blue <- "#648fff"

```

# Setting the paths

```{r setup_paths}

base_path       <- "/Users/zehao/Documents/Research Projects/Graham Group Thesis Research (Summer 2025)"
data_path       <- file.path(base_path, "Metadata")
clean_data_path <- file.path(base_path, "Cleaned data")
figure_path     <- file.path(base_path, "Figures")
cbc_figure_path <- file.path(figure_path, "CBC Figures")
dir.create(cbc_figure_path, showWarnings = FALSE, recursive = TRUE)

```

Loading the data from my directory

```{r load the data}

# mouse data

cbc_data_1 <- read_excel(file.path(data_path, "DCV_25SF.W-3thruNSF1trapout.XLSX"))
cbc_data_2 <- read_excel(file.path(data_path, "DCV_25SF.W3_NSF2.XLSX"))
meta_data  <- read_csv(file.path(data_path, "clean_meta_data.csv"))

# CDC human data

nhanes_data_2017_18 <- read_xpt(file.path(data_path, "CBC_J_NHANES_2017_2018.xpt"))
nhanes_data_2021_23 <- read_xpt(file.path(data_path, "CBC_L_NHANES_2021_2023.xpt"))

```

Setting vector for cleaner syntax

```{r}

# setting vectors
# I will use these later instead of typing everyting over and over again.

cbc_number <- c("neu_number", 
                "lym_number", 
                "mon_number", 
                "eos_number", 
                "bas_number")

cbc_percent <- c("neu_percent", 
                 "lym_percent", 
                 "mon_percent", 
                 "eos_percent", 
                 "bas_percent")

```

# Writing a plotting function

```{r}

cbc_plotter <- function(cbc_data, 
                        cbc_error, 
                        y_label, 
                        title = NULL) {
  ggplot(cbc_data, aes(x = phase,
                       y = number,
                       fill = t_muris_dose,
                       color = t_muris_dose,
                       group = interaction(t_muris_dose, 
                                           strain))) +
  geom_point(position = position_dodge(width = 0.5),
             size = 1.5, 
             alpha = 0.3) +
  geom_line(data = cbc_error,
            aes(y = mean, group = t_muris_dose),
            linewidth = 0.5,
            alpha = 1,
            position = position_dodge(width = 0.5)) +
  geom_point(data = cbc_error,
             aes(y = mean),
             position = position_dodge(width = 0.5),
             size = 2.5,
             alpha = 1) +
  geom_errorbar(data = cbc_error,
                aes(y = mean,
                    ymin = mean - se,
                    ymax = mean + se),
                position = position_dodge(width = 0.5),
                width = 0.3,
                alpha = 0.6) +
  facet_wrap(~ strain) +
  scale_fill_manual(values = c("Zero" = color_dark_blue, 
                               "Low" = color_mustard, 
                               "High" = color_orange),
                    name = expression(italic("T. muris")~"Dosage")) +
  scale_color_manual(values = c("Zero" = color_dark_blue, 
                               "Low" = color_mustard, 
                               "High" = color_orange),
                     name = expression(italic("T. muris")~"Dosage")) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white", 
                                        color = "black", 
                                        linewidth = 0.8),
        strip.text = element_text(size = 10, 
                                  face = "bold"),
        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold")) +
  xlab("Phase of the Experiment") +
  ylab(y_label) +
  ggtitle(title)
}

```




Clean data for ease of analysis

```{r}
clean_cbc_data_1 <- cbc_data_1 %>% 
  clean_names() %>%
  # clean_names() just makes all the names more R friendly automatically.
  filter(id != "BG", 
         !id %in% c("25SF.W5.1330.0.1", 
                    "25SF.W3.1311.0", 
                    "25SF.17.1380.-3", 
                    "25SF.W5.1330.0", 
                    "25SF.W5.1330.2")) %>%
  # I want to filter out all the background runs we did (BG), and also remove some samples that had no valid readings like ***** or something. 
  # TODO understand the id != and the !id %in% stuff.
  separate(col = id,
           into = c("experiment", 
                    "cage_wedge", 
                    "mouse_id", 
                    "week", 
                    "repeat"),
           sep = "\\.",
           fill = "right",
           extra = "merge",
           remove = FALSE) %>%
  # Since we formatted the id names with the experiments, cage/wedge, mouse id, the week of the sample, and whether it is a repeat. I need ot extract that information by telling R that I want to seperate the column id into that string of vectors. I also need to tell that the names are separated by .
  # TODO understanding what is fill, extra, and remove is about.
    select(mouse_id, 
         all_of(cbc_number), 
         neu_percent_2, 
         lym_percent_2, 
         mon_percent, 
         eos_percent_2, 
         bas_percent_2,
         cage_wedge,
         week,
         `repeat`) %>%
  # I'm selecting all of the important data columns.
  # For some reason the original data set had two of every cbc readings. I only kept on row.
  rename(neu_percent = neu_percent_2, 
         lym_percent = lym_percent_2, 
         mon_percent = mon_percent,
         eos_percent = eos_percent_2, 
         bas_percent = bas_percent_2) %>%
  # I'm renaming the values. The format is new name = old name
  mutate(week = as.integer(week),
         cage_wedge = factor(cage_wedge),
         mouse_id = as.character(mouse_id),
         across(all_of(c(cbc_number, cbc_percent)),
                ~ as.numeric(.)))
# I'm making sure that the values in each columns are read as the correct data type.
# TODO figure out what is a factor and character. 
# TODO figure out what is the last function doing with the ~

clean_cbc_data_2 <- cbc_data_2 %>% 
  clean_names() %>%
  filter(id != "BG") %>%
  separate(col = id,
           into = c("experiment", 
                    "cage_wedge", 
                    "mouse_id", 
                    "week", 
                    "repeat"),
           sep = "\\.",
           fill = "right",
           extra = "merge",
           remove = FALSE) %>%
    select(mouse_id, 
         all_of(cbc_number), 
         neu_percent_2, 
         lym_percent_2, 
         mon_percent, 
         eos_percent_2, 
         bas_percent_2,
         cage_wedge,
         week,
         `repeat`) %>%
  rename(neu_percent = neu_percent_2, 
         lym_percent = lym_percent_2, 
         mon_percent = mon_percent,
         eos_percent = eos_percent_2, 
         bas_percent = bas_percent_2) %>%
  mutate(week = as.integer(week),
         cage_wedge = factor(cage_wedge),
         mouse_id = as.character(mouse_id),
         across(all_of(c(cbc_number, cbc_percent)),
                ~as.numeric(.)))
```

# Combining all CBC data and added meta data

```{r combining all cbc data for mice and add the meta data information}

cleaned_meta_data <- meta_data %>%
  mutate(mouse_id = as.character(mouse_id))

cleaned_combined_cbc_data <- bind_rows(clean_cbc_data_1, 
                                       clean_cbc_data_2) %>%
  select(-`repeat`) %>%
  left_join(cleaned_meta_data, by = "mouse_id") %>%
  mutate(week = as.integer(week),
         strain = factor(strain),
         lab_or_rw = factor(lab_or_rw),
         t_muris_dose = factor(t_muris_dose),
         mouse_id = factor(mouse_id)) %>%
  mutate(strain = relevel(factor(strain), ref = "C57BL/6J"), 
         t_muris_dose = relevel(factor(t_muris_dose), ref = "Zero")) %>%
  filter(!is.na(lab_or_rw)) %>%
  mutate(phase = case_when(
    week %in% c(-4, -3) ~ "Pre-RW",
    week == 0           ~ "During RW",
    week %in% c(3, 4)   ~ "Post-RW"
  ) %>% factor(levels = c("Pre-RW", "During RW", "Post-RW")))
  
```

```{r writing a function to save the data sets cleaned}

data_saver <- function(data_name,
                       path = "") {
  
write_csv(x = data_name,
          file = path)
  }

data_saver(data_name = cleaned_combined_cbc_data,
           path = file.path(clean_data_path, "cleaned_combined_cbc_data.csv"))

#[132, 193] was removed because it has different naming. Check how to fix this issue soon.

```

# Cleaning CBC's NHANES human data

```{r cleaning the human data from CDC NHANES}

cleaned_nhanes_data_2017_18 <- nhanes_data_2017_18 %>%
  clean_names() %>% 
  drop_na() %>%
  rename(respondent_id = seqn,
         lym_percent = lbxlypct,
         mon_percent = lbxmopct,
         neu_percent = lbxnepct,
         bas_percent = lbxbapct,
         eos_percent = lbxeopct,
         lym_number = lbdlymno,
         neu_number = lbdneno,
         mon_number = lbdmono,
         bas_number = lbdbano,
         eos_number = lbdeono) %>%
  select(respondent_id, 
         all_of(cbc_number), 
         cbc_percent) %>%
  mutate(respondent_id = as.character(respondent_id),
         across(all_of(c(cbc_number, cbc_percent)),
                ~as.numeric(.)))

cleaned_nhanes_data_2021_23 <- nhanes_data_2021_23 %>%
  clean_names() %>% 
  drop_na() %>%
  rename(respondent_id = seqn,
         lym_percent = lbxlypct,
         mon_percent = lbxmopct,
         neu_percent = lbxnepct,
         bas_percent = lbxbapct,
         eos_percent = lbxeopct,
         lym_number = lbdlymno,
         neu_number = lbdneno,
         mon_number = lbdmono,
         bas_number = lbdbano,
         eos_number = lbdeono) %>%
  select(respondent_id, 
         all_of(cbc_number), 
         cbc_percent) %>%
  mutate(respondent_id = as.character(respondent_id),
         across(all_of(c(cbc_number, cbc_percent)),
                ~as.numeric(.)))

```

Save the clean data set

```{r saving the data set}

data_saver(data_name = cleaned_nhanes_data_2017_18,
           path = file.path(clean_data_path, "cleaned_nhanes_data_2017_18.csv"))

data_saver(data_name = cleaned_nhanes_data_2021_23,
           path = file.path(clean_data_path, "cleaned_nhanes_data_2021_23.csv"))

```

Make the data into long form for ggplot

```{r make the columns into long format for all}

cleaned_combined_cbc_data_long <- cleaned_combined_cbc_data %>%
  pivot_longer(cols = c(all_of(cbc_percent), 
                        all_of(cbc_number)),
               names_to = c("cell_type", 
                            ".value"),
               names_pattern = "(.*)_(number|percent)") %>% 
  filter(!is.na(lab_or_rw))

cleaned_nhanes_data_2017_18_long <- cleaned_nhanes_data_2017_18 %>% 
  pivot_longer(cols = all_of(cbc_percent),
               names_to = "cell_type",
               values_to = "percent") %>%
  mutate(cell_type = sub("_percent$", 
                         "",
                         cell_type))
  
cleaned_nhanes_data_2021_23_long <- cleaned_nhanes_data_2021_23 %>% 
  pivot_longer(cols = all_of(cbc_percent),
               names_to = "cell_type",
               values_to = "percent") %>%
  mutate(cell_type = sub("_percent$", 
                         "", 
                         cell_type))

```

# Save the long form

```{r saving}

data_saver(data_name = cleaned_combined_cbc_data_long,
           path = file.path(clean_data_path, "cleaned_combined_cbc_data_long.csv"))

data_saver(data_name = cleaned_nhanes_data_2017_18_long,
           path = file.path(clean_data_path, "cleaned_nhanes_data_2017_18_long.csv"))

data_saver(data_name = cleaned_nhanes_data_2021_23_long,
           path = file.path(clean_data_path, "cleaned_nhanes_data_2021_23_long.csv"))
```

```{r plot the pca plot}

#For color use these codes
#000000,
#E69F00, (Mustard Yellow)
#56B4E9, (Light Blue)
#009E73, (Green)
#F0E442, (Light Yellow)
#0072B2, (Dark Blue)
#D55E00, (Orange)
#CC79A7, (Purple)

```

Writing a function to do diagnostics

```{r}

qq_plotter <- function(model,
                       input,
                       condition) { 
  # I'm stating the variables that I want to input
  qqnorm(resid(model)) 
  # I'm saying that the variable model goes into resid() and then ran as qqnorm
  qqline(resid(model))
  # same thing is happening here
  plot(resid(model) ~ input[[condition]])
  # I'm plotting the resid(model) and split by the conditions given.
  # So you have two values on the x axis.
}

```

Note to self: residuals are "leftovers" that the lmer model can't explain.
Essentially it is actual value - modeled value. 
Why is it important: the lmer model assumes that all residuals are just noise.
So, if there are any patterns in the residuals,  then there might be some
issues with the model.

```{r running diagnostics in neutrophils}

neu_lmer <- lmer(neu_number ~ lab_or_rw * strain * t_muris_dose +
                   (1 | cage_wedge) + (1 | cohort) + (1 | mouse_id),
                 data = cleaned_combined_cbc_data)

# NOTE: This untransformed model is a diagnostic stepping stone only.
# Run DHARMa to confirm whether log transform is needed (expect non-uniform residuals here).
neu_sim_res <- simulateResiduals(fittedModel = neu_lmer, n = 1000)
plot(neu_sim_res)
testUniformity(neu_sim_res)
testDispersion(neu_sim_res)
testOutliers(neu_sim_res)
plotResiduals(neu_sim_res, form = cleaned_combined_cbc_data$lab_or_rw)
plotResiduals(neu_sim_res, form = cleaned_combined_cbc_data$t_muris_dose)
plotResiduals(neu_sim_res, form = cleaned_combined_cbc_data$strain)

ranef(neu_lmer)
VarCorr(neu_lmer)
summary(neu_lmer)

#Need to do log transform according to the results of the diagnostics

```
# Log transforming the neutrophil data

```{r log model of the neu}

neu_lmer_log <- lmer(log(neu_number) ~ lab_or_rw * strain * t_muris_dose +
                   (1 | cage_wedge) + (1 | cohort) + (1 | mouse_id),
                 data = cleaned_combined_cbc_data)

# DHARMa diagnostics for the primary log-transformed neutrophil model
neu_log_sim_res <- simulateResiduals(fittedModel = neu_lmer_log, n = 1000)
plot(neu_log_sim_res)
testUniformity(neu_log_sim_res)
testDispersion(neu_log_sim_res)
testOutliers(neu_log_sim_res)
plotResiduals(neu_log_sim_res, form = cleaned_combined_cbc_data$lab_or_rw)
plotResiduals(neu_log_sim_res, form = cleaned_combined_cbc_data$t_muris_dose)
plotResiduals(neu_log_sim_res, form = cleaned_combined_cbc_data$strain)

ranef(neu_lmer_log)
VarCorr(neu_lmer_log)

```
# Diagonostics to determine whether transformation is needed for lymphocyte

```{r lymphocytes diagnostics}
lym_lmer <- lmer(lym_number ~ lab_or_rw * strain * t_muris_dose +
                   (1 | cage_wedge) + (1 | cohort) + (1 | mouse_id),
                 data = cleaned_combined_cbc_data)

# NOTE: This untransformed model is a diagnostic stepping stone only.
lym_sim_res <- simulateResiduals(fittedModel = lym_lmer, n = 1000)
plot(lym_sim_res)
testUniformity(lym_sim_res)
testDispersion(lym_sim_res)
testOutliers(lym_sim_res)
plotResiduals(lym_sim_res, form = cleaned_combined_cbc_data$lab_or_rw)
plotResiduals(lym_sim_res, form = cleaned_combined_cbc_data$t_muris_dose)
plotResiduals(lym_sim_res, form = cleaned_combined_cbc_data$strain)

ranef(lym_lmer)
VarCorr(lym_lmer)
```

# Log transforming the lymphocyte data
```{r log model of the lym}

lym_lmer_log <- lmer(log(lym_number) ~ lab_or_rw * strain * t_muris_dose +
                   (1 | cage_wedge) + (1 | cohort) + (1 | mouse_id),
                 data = cleaned_combined_cbc_data)

summary(lym_lmer_log)

# DHARMa diagnostics for the primary log-transformed lymphocyte model
lym_log_sim_res <- simulateResiduals(fittedModel = lym_lmer_log, n = 1000)
plot(lym_log_sim_res)
testUniformity(lym_log_sim_res)
testDispersion(lym_log_sim_res)
testOutliers(lym_log_sim_res)
plotResiduals(lym_log_sim_res, form = cleaned_combined_cbc_data$lab_or_rw)
plotResiduals(lym_log_sim_res, form = cleaned_combined_cbc_data$t_muris_dose)
plotResiduals(lym_log_sim_res, form = cleaned_combined_cbc_data$strain)

ranef(lym_lmer_log)
VarCorr(lym_lmer_log)

```

# Diagonostics to determine whether transformation is needed for monocytes

```{r monocytes diagnostics}
mon_lmer <- lmer(mon_number ~ lab_or_rw * strain * t_muris_dose +
                   (1 | cage_wedge) + (1 | cohort) + (1 | mouse_id),
                 data = cleaned_combined_cbc_data)

# NOTE: This untransformed model is a diagnostic stepping stone only.
mon_sim_res <- simulateResiduals(fittedModel = mon_lmer, n = 1000)
plot(mon_sim_res)
testUniformity(mon_sim_res)
testDispersion(mon_sim_res)
testOutliers(mon_sim_res)
plotResiduals(mon_sim_res, form = cleaned_combined_cbc_data$lab_or_rw)
plotResiduals(mon_sim_res, form = cleaned_combined_cbc_data$t_muris_dose)
plotResiduals(mon_sim_res, form = cleaned_combined_cbc_data$strain)

ranef(mon_lmer)
VarCorr(mon_lmer)
```

# Log transforming the monocyte data

```{r log model of the mon}

mon_lmer_log <- lmer(log(mon_number) ~ lab_or_rw + strain * t_muris_dose +
                   (1 | cage_wedge) + (1 | cohort) + (1 | mouse_id),
                 data = cleaned_combined_cbc_data)

print(summary(mon_lmer_log), 
      correlation = FALSE)

# DHARMa diagnostics for the primary log-transformed monocyte model
# NOTE: mon_lmer_log uses a reduced fixed-effects formula (no 3-way interaction)
mon_log_sim_res <- simulateResiduals(fittedModel = mon_lmer_log, n = 1000)
plot(mon_log_sim_res)
testUniformity(mon_log_sim_res)
testDispersion(mon_log_sim_res)
testOutliers(mon_log_sim_res)
plotResiduals(mon_log_sim_res, form = cleaned_combined_cbc_data$lab_or_rw)
plotResiduals(mon_log_sim_res, form = cleaned_combined_cbc_data$t_muris_dose)
plotResiduals(mon_log_sim_res, form = cleaned_combined_cbc_data$strain)

ranef(mon_lmer_log)
VarCorr(mon_lmer_log)

```

# Diagonostics to determine whether transformation is needed for eosinophils

```{r eosinophils diagnostics}

eos_lmer <- lmer(eos_number ~ lab_or_rw * strain * t_muris_dose +
                   (1 | cage_wedge) + (1 | cohort) + (1 | mouse_id),
                 data = cleaned_combined_cbc_data)

# NOTE: This untransformed model is a diagnostic stepping stone only.
eos_sim_res <- simulateResiduals(fittedModel = eos_lmer, n = 1000)
plot(eos_sim_res)
testUniformity(eos_sim_res)
testDispersion(eos_sim_res)
testOutliers(eos_sim_res)
plotResiduals(eos_sim_res, form = cleaned_combined_cbc_data$lab_or_rw)
plotResiduals(eos_sim_res, form = cleaned_combined_cbc_data$t_muris_dose)
plotResiduals(eos_sim_res, form = cleaned_combined_cbc_data$strain)

ranef(eos_lmer)
VarCorr(eos_lmer)
```

# Log transforming the eosinophil data (log1p to handle zeros)

```{r log model of the eos}

eos_lmer_log <- lmer(log1p(eos_number) ~ lab_or_rw * strain * t_muris_dose +
                   (1 | cage_wedge) + (1 | cohort) + (1 | mouse_id),
                 data = cleaned_combined_cbc_data)

# NOTE: log1p(eos_number) is used instead of log() because eos_number can be zero.
# TODO use glmer here

print(summary(eos_lmer_log),
      correlation = FALSE)

# DHARMa diagnostics for the primary log1p-transformed eosinophil model
eos_log_sim_res <- simulateResiduals(fittedModel = eos_lmer_log, n = 1000)
plot(eos_log_sim_res)
testUniformity(eos_log_sim_res)
testDispersion(eos_log_sim_res)
testOutliers(eos_log_sim_res)
plotResiduals(eos_log_sim_res, form = cleaned_combined_cbc_data$lab_or_rw)
plotResiduals(eos_log_sim_res, form = cleaned_combined_cbc_data$t_muris_dose)
plotResiduals(eos_log_sim_res, form = cleaned_combined_cbc_data$strain)

ranef(eos_lmer_log)
VarCorr(eos_lmer_log)

```

# Diagonostics to determine whether transformation is needed for basophils

```{r basophils}
bas_lmer <- lmer(bas_number ~ lab_or_rw * strain * t_muris_dose +
                   (1 | cage_wedge) + (1 | cohort) + (1 | mouse_id),
                 data = cleaned_combined_cbc_data)

# NOTE: This untransformed model is a diagnostic stepping stone only.
# bas_number can include zeros, which is why log1p is used in the primary model.
bas_sim_res <- simulateResiduals(fittedModel = bas_lmer, n = 1000)
plot(bas_sim_res)
testUniformity(bas_sim_res)
testDispersion(bas_sim_res)
testOutliers(bas_sim_res)
plotResiduals(bas_sim_res, form = cleaned_combined_cbc_data$lab_or_rw)
plotResiduals(bas_sim_res, form = cleaned_combined_cbc_data$t_muris_dose)
plotResiduals(bas_sim_res, form = cleaned_combined_cbc_data$strain)

ranef(bas_lmer)
VarCorr(bas_lmer)

```

# Log transforming the basophil data

```{r log model of the bas}

bas_lmer_log <- lmer(log1p(bas_number) ~ lab_or_rw * strain * t_muris_dose +
                   (1 | cage_wedge) + (1 | cohort) + (1 | mouse_id),
                 data = cleaned_combined_cbc_data)

print(summary(bas_lmer_log), 
      correlation = FALSE)

# DHARMa diagnostics for the primary log1p-transformed basophil model
# NOTE: log1p(bas_number) is used instead of log() because bas_number can be zero
bas_log_sim_res <- simulateResiduals(fittedModel = bas_lmer_log, n = 1000)
plot(bas_log_sim_res)
testUniformity(bas_log_sim_res)
testDispersion(bas_log_sim_res)
testOutliers(bas_log_sim_res)
plotResiduals(bas_log_sim_res, form = cleaned_combined_cbc_data$lab_or_rw)
plotResiduals(bas_log_sim_res, form = cleaned_combined_cbc_data$t_muris_dose)
plotResiduals(bas_log_sim_res, form = cleaned_combined_cbc_data$strain)

ranef(bas_lmer_log)
VarCorr(bas_lmer_log)

```
# Saving the results as a csv file

```{r}
analysis_directory <- file.path(base_path, "Analysis Results")

models <- list(neu = neu_lmer_log,
               lym = lym_lmer_log,
               mon = mon_lmer_log,
               eos = eos_lmer_log,
               bas = bas_lmer_log)

all_coef <- lapply(names(models), 
                   function(cell) 
                     {all_dat <- tidy(models[[cell]], 
                                      conf.int = TRUE, 
                                      effects = "fixed") 
                     all_dat$cell_type <- cell
                     return(all_dat)}) %>%
  bind_rows()

write_csv(all_coef, 
          file.path(analysis_directory, "cbc_lmer_coefficients.csv"))


all_coef %>%
  group_split(cell_type) %>%
  lapply(function(individual_cell) { stopifnot(length(unique(individual_cell$cell_type)) == 1)
    cell <- unique(individual_cell$cell_type)
    write_csv(individual_cell, 
              file.path(analysis_directory, 
                        paste0("cbc_lmer_coefficients_", 
                               cell, 
                               ".csv")))})

```


# ggplot visualization of lym vs time for rw only

```{r plot the lym over time relationship RW ONLY}
#Filter for the correct data to plot
lym_rw_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(lab_or_rw == "RW",
         cell_type == "lym") %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")),
         phase = factor(phase, levels = c("Pre-RW", "During RW", "Post-RW")))
#Calculate the standard error and the mean
standard_error_lym_rw <- lym_rw_plot_data %>%
  group_by(phase, 
           t_muris_dose, 
           strain) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n)) %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", 
                                          "Low", 
                                          "High")))
```

```{r}

plot_saver <- function(filename,
                            plot,
                            path = cbc_figure_path,
                            width = 6,
                            height = 4,
                            dpi = 300) {
  ggsave(filename = filename,
         plot = plot,
         path = path,
         width = width,
         height = height,
         units = "in",
         dpi = dpi)
}

```


```{r}
lym_rw_plot <- cbc_plotter(lym_rw_plot_data,
                           standard_error_lym_rw,
                           expression(bold("Lymphocyte Count (10"^3~"cells/"*mu*"L)")),
                           "Lymphocyte Count Over Time in Rewilded Mice")

lym_rw_plot

plot_saver("lym_number_rw.png",
                plot = lym_rw_plot)
```

# ggplot visualization of lym vs time for lab only

```{r plot lym over time LAB ONLY}

#Filter for the correct data to plot
lym_lab_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(lab_or_rw == "Lab", cell_type == "lym") %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")),
         phase = factor(phase, levels = c("Pre-RW", "During RW", "Post-RW")))
#Calculate the standard error and the mean
standard_error_lym_lab <- lym_lab_plot_data %>%
  group_by(phase, 
           t_muris_dose, 
           strain) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n)) %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")))

```


```{r}
lym_lab_plot <- cbc_plotter(lym_lab_plot_data,
                            standard_error_lym_lab,
                            expression(bold("Lymphocyte Count (10"^3~"cells/"*mu*"L)")),
                            "Lymphocyte Count Over Time in Lab Mice")

lym_lab_plot

plot_saver("lym_number_lab.png",
                plot = lym_lab_plot)
```

# ggplot visualization of neu vs time for rw only

```{r neu over time in RW ONLY}

#Filter for the correct data to plot
neu_rw_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(lab_or_rw == "RW", cell_type == "neu") %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")),
         phase = factor(phase, levels = c("Pre-RW", "During RW", "Post-RW")))
#Calculate the standard error and the mean
standard_error_neu_rw <- neu_rw_plot_data %>%
  group_by(phase, 
           t_muris_dose, 
           strain) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n)) %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")))
```

```{r}

neu_rw_plot <- cbc_plotter(neu_rw_plot_data,
                           standard_error_neu_rw,
                           expression(bold("Neutrophil Count (10"^3~"cells/"*mu*"L)")),
                           "Neutrophil Count Over Time in Rewilded Mice")

neu_rw_plot

plot_saver("neu_number_rw.png",
                plot = neu_rw_plot)
```


# ggplot visualization of neu vs time for lab only

```{r plot neu over time LAB ONLY}

#Filter for the correct data to plot
neu_lab_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(lab_or_rw == "Lab", cell_type == "neu") %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")),
         phase = factor(phase, levels = c("Pre-RW", "During RW", "Post-RW")))
#Calculate the standard error and the mean
standard_error_neu_lab <- neu_lab_plot_data %>%
  group_by(phase,
           t_muris_dose,
           strain) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n)) %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")))



neu_lab_plot <- cbc_plotter(neu_lab_plot_data,
                            standard_error_neu_lab,
                            expression(bold("Neutrophil Count (10"^3~"cells/"*mu*"L)")),
                            "Neutrophil Count Over Time in Lab Control Mice")

neu_lab_plot

plot_saver("neu_lab_plot.png",
                plot = neu_lab_plot)

```

# ggplot visualization of mon vs time for rw only

```{r plot mon over time RW ONLY}

#Filter for the correct data to plot
mon_rw_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(lab_or_rw == "RW", cell_type == "mon") %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")),
         phase = factor(phase, levels = c("Pre-RW", "During RW", "Post-RW")))
#Calculate the standard error and the mean
standard_error_mon_rw <- mon_rw_plot_data %>%
  group_by(phase,
           t_muris_dose,
           strain) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n)) %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")))



mon_rw_plot <- cbc_plotter(mon_rw_plot_data,
                           standard_error_mon_rw,
                           expression(bold("Monocyte Count (10"^3~"cells/"*mu*"L)")),
                           "Monocyte Count Over Time in Rewilded Mice")

mon_rw_plot

plot_saver("mon_over_time_rw.png",
                plot = mon_rw_plot)
```

# ggplot visualization of mon vs time for lab only

```{r plot mon over time LAB ONLY}

#Filter for the correct data to plot
mon_lab_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(lab_or_rw == "Lab", cell_type == "mon") %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")),
         phase = factor(phase, levels = c("Pre-RW", "During RW", "Post-RW")))
#Calculate the standard error and the mean
standard_error_mon_lab <- mon_lab_plot_data %>%
  group_by(phase,
           t_muris_dose,
           strain) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n)) %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")))



mon_lab_plot <- cbc_plotter(mon_lab_plot_data,
                            standard_error_mon_lab,
                            expression(bold("Monocyte Count (10"^3~"cells/"*mu*"L)")),
                            "Monocyte Count Over Time in Lab Control Mice")

mon_lab_plot

plot_saver("mon_over_time_lab.png",
                plot = mon_lab_plot)
```

# ggplot visualization of eos vs time for rw only

```{r plot eos over time RW ONLY}

#Filter for the correct data to plot
eos_rw_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(lab_or_rw == "RW", cell_type == "eos") %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")),
         phase = factor(phase, levels = c("Pre-RW", "During RW", "Post-RW")))
#Calculate the standard error and the mean
standard_error_eos_rw <- eos_rw_plot_data %>%
  group_by(phase,
           t_muris_dose,
           strain) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n)) %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")))



eos_rw_plot <- cbc_plotter(eos_rw_plot_data,
                           standard_error_eos_rw,
                           expression(bold("Eosinophil Count (10"^3~"cells/"*mu*"L)")),
                           "Eosinophil Count Over Time in Rewilded Mice")

eos_rw_plot

plot_saver("eos_over_time_rw.png",
                plot = eos_rw_plot)
```

# ggplot visualization of eos vs time for lab only

```{r plot eos over time LAB ONLY}

#Filter for the correct data to plot
eos_lab_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(lab_or_rw == "Lab", cell_type == "eos") %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")),
         phase = factor(phase, levels = c("Pre-RW", "During RW", "Post-RW")))
#Calculate the standard error and the mean
standard_error_eos_lab <- eos_lab_plot_data %>%
  group_by(phase,
           t_muris_dose,
           strain) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n)) %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")))



eos_lab_plot <- cbc_plotter(eos_lab_plot_data,
                            standard_error_eos_lab,
                            expression(bold("Eosinophil Count (10"^3~"cells/"*mu*"L)")),
                            "Eosinophil Count Over Time in Lab Control Mice")

eos_lab_plot

plot_saver("eos_lab_plot.png",
                plot = eos_lab_plot)
```

# ggplot visualization of bas vs time for rw only

```{r plot bas over time RW ONLY}

#Filter for the correct data to plot
bas_rw_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(lab_or_rw == "RW", cell_type == "bas") %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")),
         phase = factor(phase, levels = c("Pre-RW", "During RW", "Post-RW")))
#Calculate the standard error and the mean
standard_error_bas_rw <- bas_rw_plot_data %>%
  group_by(phase,
           t_muris_dose,
           strain) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n)) %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")))



bas_rw_plot <- cbc_plotter(bas_rw_plot_data,
                           standard_error_bas_rw,
                           expression(bold("Basophil Count (10"^3~"cells/"*mu*"L)")),
                           "Basophil Count Over Time in Rewilded Mice")

bas_rw_plot

plot_saver("bas_rw_plot.png",
                plot = bas_rw_plot)
```

# ggplot visualization of bas vs time for lab only

```{r plot bas over time LAB ONLY}

#Filter for the correct data to plot
bas_lab_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(lab_or_rw == "Lab", cell_type == "bas") %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")),
         phase = factor(phase, levels = c("Pre-RW", "During RW", "Post-RW")))
#Calculate the standard error and the mean
standard_error_bas_lab <- bas_lab_plot_data %>%
  group_by(phase,
           t_muris_dose,
           strain) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n)) %>%
  mutate(t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", "Low", "High")))



bas_lab_plot <- cbc_plotter(bas_lab_plot_data,
                            standard_error_bas_lab,
                            expression(bold("Basophil Count (10"^3~"cells/"*mu*"L)")),
                            "Basophil Count Over Time in Lab Control Mice")

bas_lab_plot

plot_saver("bas_lab_plot.png",
                plot = bas_lab_plot)
```
