---
title: "CBC data analysis"
author: "Zehao Wu"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

# First time run order: Spleen → CBC → Spleen
# Requires clean_meta_data.csv from the spleen analysis markdown

# Libraries

```{r}

library(tidyverse)
library(readxl)
library(writexl)
library(janitor)
library(factoextra)
library(skimr)
library(naniar)
library(ggplot2)
library(ggpubr)
library(emmeans)
library(lme4)
library(lmerTest)
library(DHARMa)
library(scales)
library(haven)
library(broom.mixed)
library(readr)
library(kableExtra)
library(sjPlot) # does not work when knitting to PDF

```

# Setting the colors (Okabe-Ito palette and Paul Tol's Bright)

```{r}

color_black <- "#000000"
color_mustard <- "#E69F00"
color_light_blue <- "#56B4E9"
color_green <- "#009E73"
color_yellow <- "#F0E442"
color_dark_blue <- "#0072B2"
color_orange <- "#D55E00"
color_purple <- "#CC79A7"

color_grey <- "#BBBBBB"
color_dark_purple <- "#AA3377"
color_light_purple <- "#EE6677"
color_sand <- "#CCBB44"
color_cyan <- "#66CCEE"
color_olive <- "#228833"
color_navy <- "#4477AA"

```

# Setting up paths

```{r}

base_path <- "/Users/zehao/Documents/Research Projects/Graham Group Thesis Research (Summer 2025)"
data_path <- file.path(base_path, "Metadata")
clean_data_path <- file.path(base_path, "Cleaned data")
figure_path <- file.path(base_path, "Figures")
cbc_figure_path <- file.path(figure_path, "CBC Figures")

dir.create(cbc_figure_path, 
           showWarnings = FALSE, 
           recursive = TRUE)

```

# Setting up vectors

```{r}

cbc_number <- c("neu_number",
                "lym_number",
                "mon_number",
                "eos_number",
                "bas_number")

cbc_percent <- c("neu_percent",
                 "lym_percent",
                 "mon_percent",
                 "eos_percent",
                 "bas_percent")

```

# Defining functions

```{r}

data_saver <- function(data_name,
                       path = "") {
  write_csv(x = data_name,
            file = path)
}

plot_saver <- function(filename,
                       plot,
                       path = cbc_figure_path,
                       width = 8,
                       height = 6,
                       dpi = 300) {
  ggsave(filename = filename,
         plot = plot,
         path = path,
         width = width,
         height = height,
         units = "in",
         dpi = dpi)
}

svg_saver <- function(filename,
                      plot,
                      path = cbc_figure_path,
                      width = 8,
                      height = 6) {
  ggsave(filename = filename,
         plot = plot,
         path = path,
         width = width,
         height = height,
         units = "in",
         device = "svg")
}

cbc_summarizer <- function(data,
                           group_name) {
  data %>%
    summarise(neu_percent = mean(neu_percent, 
                                 na.rm = TRUE) / 100,
              lym_percent = mean(lym_percent, 
                                 na.rm = TRUE) / 100,
              mon_percent = mean(mon_percent, 
                                 na.rm = TRUE) / 100,
              eos_percent = mean(eos_percent, 
                                 na.rm = TRUE) / 100,
              bas_percent = mean(bas_percent, 
                                 na.rm = TRUE) / 100) %>%
    mutate(group = group_name) %>%
    pivot_longer(cols = neu_percent:bas_percent,
                 names_to = "cell_type",
                 values_to = "cbc_percent")
}

emm_plotter <- function(emm_df,
                        cbc_input,
                        y_col,
                        y_label,
                        title = NULL,
                        floor_zero = FALSE) {
  if (floor_zero) {
    emm_df <- emm_df %>% mutate(lower.CL = pmax(lower.CL, 0))
  }

  ggplot(emm_df, aes(x = phase,
                     y = response,
                     color = t_muris_dose,
                     fill = t_muris_dose,
                     group = t_muris_dose)) +
    geom_point(data = cbc_input,
               aes(x = phase,
                   y = .data[[y_col]],
                   fill = t_muris_dose,
                   color = t_muris_dose),
               position = position_jitterdodge(jitter.width = 0.1,
                                               dodge.width = 0.5),
               alpha = 0.3,
               size = 1.2,
               inherit.aes = FALSE) +
    geom_line(position = position_dodge(width = 0.5),
              linewidth = 0.7,
              alpha = 0.7) +
    geom_errorbar(aes(ymin = lower.CL,
                      ymax = upper.CL),
                  width = 0.2,
                  alpha = 0.7,
                  position = position_dodge(width = 0.5)) +
    geom_point(size = 3,
               shape = 21,
               color = "black",
               position = position_dodge(width = 0.5)) +
    geom_vline(xintercept = 2.5,
               linetype = "dashed",
               color = "grey40",
               alpha = 0.6) +
    annotate("text",
             x = 2.5,
             y = Inf,
             label = "Infection",
             vjust = 1.5,
             hjust = -0.1,
             size = 3,
             fontface = "italic",
             alpha = 0.6) +
    facet_grid(lab_or_rw ~ strain,
               labeller = labeller(lab_or_rw = c("Lab" = "Lab",
                                                  "RW" = "Rewilded"))) +
    scale_fill_manual(values = c("Zero" = color_dark_blue,
                                 "Low"  = color_mustard,
                                 "High" = color_orange),
                      name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
    scale_color_manual(values = c("Zero" = color_dark_blue,
                                  "Low"  = color_mustard,
                                  "High" = color_orange),
                       name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
    scale_y_continuous(name = y_label) +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "white",
                                          color = "black",
                                          linewidth = 0.8),
          strip.text = element_text(size = 10,
                                    face = "bold"),
          panel.border = element_rect(color = "black",
                                      fill = NA,
                                      linewidth = 0.8),
          panel.grid = element_blank(),
          plot.title = element_text(hjust = 0.5,
                                    face = "bold"),
          axis.title.x = element_text(face = "bold"),
          axis.title.y = element_text(face = "bold"),
          axis.text = element_text(color = "black"),
          axis.ticks = element_line(color = "black",
                                    linewidth = 0.3),
          axis.ticks.length = unit(0.15, "cm")) +
    xlab("Phase") +
    ggtitle(title)
}

cbc_plotter <- function(cbc_data,
                        cbc_error,
                        y_label,
                        title = NULL) {
  ggplot(cbc_data, aes(x = phase,
                       y = number,
                       fill = t_muris_dose,
                       color = t_muris_dose,
                       group = t_muris_dose)) +
  geom_point(position = position_dodge(width = 0.5),
             size = 1.5,
             alpha = 0.3) +
  geom_line(data = cbc_error,
            aes(y = mean, group = t_muris_dose),
            linewidth = 0.5,
            alpha = 0.6,
            position = position_dodge(width = 0.5)) +
  geom_point(data = cbc_error,
             aes(y = mean),
             position = position_dodge(width = 0.5),
             size = 2.5,
             alpha = 1,
             shape = 21,
             color = "black") +
  geom_errorbar(data = cbc_error,
                aes(y = mean,
                    ymin = mean - se,
                    ymax = mean + se),
                position = position_dodge(width = 0.5),
                width = 0.15,
                alpha = 0.6) +
  geom_vline(xintercept = 2.5,
             linetype = "dashed",
             color = "grey40",
             alpha = 0.6) +
  annotate("text",
           x = 2.5,
           y = Inf,
           label = "Infection",
           vjust = 1.5,
           hjust = -0.1,
           size = 3,
           fontface = "italic",
           alpha = 0.6) +
  facet_grid(lab_or_rw ~ strain,
             labeller = labeller(lab_or_rw = c("Lab" = "Lab",
                                                "RW" = "Rewilded"))) +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "Low"  = color_mustard,
                               "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "Low"  = color_mustard,
                                "High" = color_orange),
                     name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5,
                                  face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm")) +
  xlab("Phase") +
  ylab(y_label) +
  ggtitle(title)
}

```

# Loading Data

```{r}

# Mouse CBC data
cbc_data_1 <- read_excel(file.path(data_path, 
                                   "DCV_25SF.W-3thruNSF1trapout.XLSX"))
cbc_data_2 <- read_excel(file.path(data_path,
                                   "DCV_25SF.W3_NSF2.XLSX"))
meta_data   <- read_csv(file.path(clean_data_path,
                                  "clean_meta_data.csv"))

# CDC NHANES human data
nhanes_data_2017_18 <- read_xpt(file.path(data_path, 
                                          "CBC_J_NHANES_2017_2018.xpt"))
nhanes_data_2021_23 <- read_xpt(file.path(data_path, 
                                          "CBC_L_NHANES_2021_2023.xpt"))

```

# Cleaning Mouse CBC Data

```{r}

clean_cbc_data_1 <- cbc_data_1 %>%
  clean_names() %>%
  filter(id != "BG",
         !id %in% c("25SF.W5.1330.0.1",
                    "25SF.W3.1311.0",
                    "25SF.17.1380.-3",
                    "25SF.W5.1330.0",
                    "25SF.W5.1330.2")) %>%
  separate(col = id,
           into = c("experiment",
                    "cage_wedge",
                    "mouse_id",
                    "week",
                    "repeat"),
           sep = "\\.",
           fill = "right",
           extra = "merge",
           remove = FALSE) %>%
  select(mouse_id,
         all_of(cbc_number),
         neu_percent_2,
         lym_percent_2,
         mon_percent,
         eos_percent_2,
         bas_percent_2,
         cage_wedge,
         week,
         `repeat`) %>%
  rename(neu_percent = neu_percent_2,
         lym_percent = lym_percent_2,
         eos_percent = eos_percent_2,
         bas_percent = bas_percent_2) %>%
  mutate(week = as.integer(week),
         cage_wedge = factor(cage_wedge),
         mouse_id = as.character(mouse_id),
         across(all_of(c(cbc_number, 
                         cbc_percent)), ~ as.numeric(.)))

clean_cbc_data_2 <- cbc_data_2 %>%
  clean_names() %>%
  filter(id != "BG") %>%
  separate(col = id,
           into = c("experiment",
                    "cage_wedge",
                    "mouse_id",
                    "week",
                    "repeat"),
           sep = "\\.",
           fill = "right",
           extra = "merge",
           remove = FALSE) %>%
  select(mouse_id,
         all_of(cbc_number),
         neu_percent_2,
         lym_percent_2,
         mon_percent,
         eos_percent_2,
         bas_percent_2,
         cage_wedge,
         week,
         `repeat`) %>%
  rename(neu_percent = neu_percent_2,
         lym_percent = lym_percent_2,
         eos_percent = eos_percent_2,
         bas_percent = bas_percent_2) %>%
  mutate(week = as.integer(week),
         cage_wedge = factor(cage_wedge),
         mouse_id = as.character(mouse_id),
         across(all_of(c(cbc_number, cbc_percent)), ~ as.numeric(.)))

```

# Combining Mouse CBC Data and Adding Metadata

```{r}

cleaned_meta_data <- meta_data %>%
  mutate(mouse_id = as.character(mouse_id))

cleaned_combined_cbc_data <- bind_rows(clean_cbc_data_1,
                                       clean_cbc_data_2) %>%
  select(-`repeat`) %>%
  left_join(cleaned_meta_data, by = "mouse_id") %>%
  mutate(cage_wedge = tolower(cage_wedge)) %>%
  mutate(week = as.integer(week),
         strain = factor(strain),
         lab_or_rw = factor(lab_or_rw),
         t_muris_dose = factor(t_muris_dose),
         mouse_id = factor(mouse_id)) %>%
  mutate(strain = factor(strain, 
                         levels = c("C57BL/6J",
                                    "B10.BR")),
         lab_or_rw = factor(lab_or_rw,
                            levels = c("Lab",
                                       "RW")),
         t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero",
                                          "Low",
                                          "High"))) %>%
  filter(!is.na(lab_or_rw)) %>%
  mutate(phase = case_when(
    week %in% c(-4, -3) ~ "Pre-RW",
    week == 0 ~ "During RW",
    week %in% c(3, 4) ~ "Post-RW"
  ) %>% factor(levels = c("Pre-RW", "During RW", "Post-RW")))

cbc_endpoint <- cleaned_combined_cbc_data %>%
  filter(phase == "Post-RW")

# [132, 193] was removed because it has different naming. 
# Check how to fix this issue soon.

```

```{r}

data_saver(data_name = cleaned_combined_cbc_data,
           path = file.path(clean_data_path,
                            "cleaned_combined_cbc_data.csv"))

```

# Sample sizes per time point

```{r}

cbc_sample_counts <- cleaned_combined_cbc_data %>%
  group_by(phase, strain, lab_or_rw, t_muris_dose) %>%
  summarise(n = n(), .groups = "drop")

kable(cbc_sample_counts) %>%
  kable_styling(full_width = FALSE)

```

# Cleaning NHANES Human Data

```{r}

cleaned_nhanes_data_2017_18 <- nhanes_data_2017_18 %>%
  clean_names() %>%
  rename(respondent_id = seqn,
         lym_percent = lbxlypct,
         mon_percent = lbxmopct,
         neu_percent = lbxnepct,
         bas_percent = lbxbapct,
         eos_percent = lbxeopct,
         lym_number = lbdlymno,
         neu_number = lbdneno,
         mon_number = lbdmono,
         bas_number = lbdbano,
         eos_number = lbdeono) %>%
  select(respondent_id,
         all_of(cbc_number),
         all_of(cbc_percent)) %>%
  drop_na() %>%
  mutate(respondent_id = as.character(respondent_id),
         across(all_of(c(cbc_number, cbc_percent)), ~ as.numeric(.)))

cleaned_nhanes_data_2021_23 <- nhanes_data_2021_23 %>%
  clean_names() %>%
  rename(respondent_id = seqn,
         lym_percent = lbxlypct,
         mon_percent = lbxmopct,
         neu_percent = lbxnepct,
         bas_percent = lbxbapct,
         eos_percent = lbxeopct,
         lym_number = lbdlymno,
         neu_number = lbdneno,
         mon_number = lbdmono,
         bas_number = lbdbano,
         eos_number = lbdeono) %>%
  select(respondent_id,
         all_of(cbc_number),
         all_of(cbc_percent)) %>%
  drop_na() %>%
  mutate(respondent_id = as.character(respondent_id),
         across(all_of(c(cbc_number, 
                         cbc_percent)), ~ as.numeric(.)))

data_saver(data_name = cleaned_nhanes_data_2017_18,
           path = file.path(clean_data_path,
                            "cleaned_nhanes_data_2017_18.csv"))

data_saver(data_name = cleaned_nhanes_data_2021_23,
           path = file.path(clean_data_path,
                            "cleaned_nhanes_data_2021_23.csv"))

```

# Sample size of NHANES 2021-23 data

```{r}

nrow(cleaned_nhanes_data_2021_23)

```

# Pairwise comparison of human NHANES data (justification for pooling)

```{r}

human_year_comparison <- bind_rows(
  cleaned_nhanes_data_2017_18 %>% mutate(nhanes_year = "2017_18"),
  cleaned_nhanes_data_2021_23 %>% mutate(nhanes_year = "2021_23")
) %>%
  mutate(nhanes_year = factor(nhanes_year))

# Neutrophil
neu_year_lm <- lm(neu_percent ~ nhanes_year,
                   data = human_year_comparison)

kable(pairs(emmeans(neu_year_lm, specs = "nhanes_year")),
      digits = 3,
      caption = "Neutrophil %: NHANES 2017-18 vs 2021-23") %>%
  kable_styling(full_width = FALSE)

# Lymphocyte
lym_year_lm <- lm(lym_percent ~ nhanes_year,
                   data = human_year_comparison)

kable(pairs(emmeans(lym_year_lm, specs = "nhanes_year")),
      digits = 3,
      caption = "Lymphocyte %: NHANES 2017-18 vs 2021-23") %>%
  kable_styling(full_width = FALSE)

# Monocyte
mon_year_lm <- lm(mon_percent ~ nhanes_year,
                   data = human_year_comparison)

kable(pairs(emmeans(mon_year_lm, specs = "nhanes_year")),
      digits = 3,
      caption = "Monocyte %: NHANES 2017-18 vs 2021-23") %>%
  kable_styling(full_width = FALSE)

# Eosinophil
eos_year_lm <- lm(eos_percent ~ nhanes_year,
                   data = human_year_comparison)

kable(pairs(emmeans(eos_year_lm, specs = "nhanes_year")),
      digits = 3,
      caption = "Eosinophil %: NHANES 2017-18 vs 2021-23") %>%
  kable_styling(full_width = FALSE)

# Basophil
bas_year_lm <- lm(bas_percent ~ nhanes_year,
                   data = human_year_comparison)

kable(pairs(emmeans(bas_year_lm, specs = "nhanes_year")),
      digits = 3,
      caption = "Basophil %: NHANES 2017-18 vs 2021-23") %>%
  kable_styling(full_width = FALSE)

# it seems that there are significant differences between the years. Let's pick 2021 and 2023.

```

# Making long-form mouse CBC data to be plotted

```{r}

cleaned_combined_cbc_data_long <- cleaned_combined_cbc_data %>%
  pivot_longer(cols = c(all_of(cbc_percent),
                        all_of(cbc_number)),
               names_to = c("cell_type",
                            ".value"),
               names_pattern = "(.*)_(number|percent)") %>%
  filter(!is.na(lab_or_rw))

data_saver(data_name = cleaned_combined_cbc_data_long,
           path = file.path(clean_data_path, 
                            "cleaned_combined_cbc_data_long.csv"))

```

# Composition Plot for Human vs. Mouse understanding

```{r}

human_summary <- cbc_summarizer(cleaned_nhanes_data_2021_23,
                                "Human") %>%
  mutate(strain = "Human")

mean_mouse <- cbc_endpoint %>%
  group_by(strain,
           t_muris_dose,
           lab_or_rw) %>%
  summarise(neu_percent = mean(neu_percent,
                               na.rm = TRUE) / 100,
            lym_percent = mean(lym_percent,
                               na.rm = TRUE) / 100,
            mon_percent = mean(mon_percent,
                               na.rm = TRUE) / 100,
            eos_percent = mean(eos_percent,
                               na.rm = TRUE) / 100,
            bas_percent = mean(bas_percent,
                               na.rm = TRUE) / 100,
            .groups = "drop") %>%
  mutate(group = paste(strain,
                       t_muris_dose,
                       lab_or_rw)) %>%
  pivot_longer(cols = neu_percent:bas_percent,
               names_to = "cell_type",
               values_to = "cbc_percent")

bar_plot_input <- bind_rows(human_summary,
                            mean_mouse) %>%
  mutate(group = factor(group,
                        levels = c(
                          "Human",
                          "C57BL/6J Zero Lab",
                          "B10.BR Zero Lab",
                          "C57BL/6J Low Lab",
                          "B10.BR Low Lab",
                          "C57BL/6J High Lab",
                          "B10.BR High Lab",
                          "C57BL/6J Zero RW",
                          "B10.BR Zero RW",
                          "C57BL/6J Low RW",
                          "B10.BR Low RW",
                          "C57BL/6J High RW",
                          "B10.BR High RW")),
         cell_type = factor(cell_type,
                            levels = c("lym_percent",
                                       "neu_percent",
                                       "mon_percent",
                                       "eos_percent",
                                       "bas_percent")),
         strain = factor(strain, levels = c("Human",
                                            "C57BL/6J",
                                            "B10.BR")),
)

data_saver(data_name = bar_plot_input,
           path = file.path(clean_data_path, 
                            "input_data_cbc_composition.csv"))

```

```{r}

comparative_cbc <- ggplot(bar_plot_input,
                    aes(x = group,
                        y = cbc_percent,
                        fill = cell_type)) +
  geom_col(position = "stack",
           color = "black",
           linewidth = 0.5,
           alpha = 0.8) +
  scale_fill_manual(values = c("neu_percent" = color_dark_purple,
                               "lym_percent" = color_light_purple,
                               "mon_percent" = color_sand,
                               "eos_percent" = color_cyan,
                               "bas_percent" = color_black),
                    labels = c("neu_percent" = "Neutrophils",
                               "lym_percent" = "Lymphocytes",
                               "mon_percent" = "Monocytes",
                               "eos_percent" = "Eosinophils",
                               "bas_percent" = "Basophils"),
                    name = expression(bold("Immune Cell Type"))) +
  scale_y_continuous(expand = expansion(mult = c(0, 
                                                 0.05)),
                     name = "CBC Composition") +
  scale_x_discrete(limits = c("Human",
                              "C57BL/6J Zero Lab",
                              "C57BL/6J Low Lab",
                              "C57BL/6J High Lab",
                              "C57BL/6J Zero RW",
                              "C57BL/6J Low RW",
                              "C57BL/6J High RW",
                              "B10.BR Zero Lab",
                              "B10.BR Low Lab",
                              "B10.BR High Lab",
                              "B10.BR Zero RW",
                              "B10.BR Low RW",
                              "B10.BR High RW"),
                   labels = c("Human" = "Human",
                              "C57BL/6J Zero Lab" = "Zero",
                              "C57BL/6J Low Lab" = "Low",
                              "C57BL/6J High Lab" = "High",
                              "C57BL/6J Zero RW" = "Zero",
                              "C57BL/6J Low RW" = "Low",
                              "C57BL/6J High RW" = "High",
                              "B10.BR Zero Lab" = "Zero",
                              "B10.BR Low Lab" = "Low",
                              "B10.BR High Lab" = "High",
                              "B10.BR Zero RW" = "Zero",
                              "B10.BR Low RW" = "Low",
                              "B10.BR High RW" = "High")) +
  coord_flip() +
  theme_minimal() +
  theme(panel.border = element_blank(),
        panel.grid = element_blank(),
        panel.spacing = unit(0.8, "cm"),
        axis.line = element_line(color = "black",
                                 linewidth = 0.5),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        plot.title = element_text(hjust = 0.5,
                                  face = "bold"),
        axis.text = element_text(color = "black"),
        axis.text.x = element_text(angle = 45,
                                   hjust = 1,
                                   size = 8),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm")) +
  xlab("") +
  ggtitle("CBC Composition: Mice vs. Humans")

comparative_cbc

plot_saver("cbc_composition_plot.png",
           plot = comparative_cbc)
svg_saver("cbc_composition_plot.svg",
          plot = comparative_cbc)

```

Running models and emmeans for each cell type

```{r}

mouse_cbc <- cbc_endpoint %>% 
  mutate(group = paste(strain, 
                       lab_or_rw, 
                       case_when(
                         t_muris_dose == "Zero" ~ "Zero Dose",
                         t_muris_dose == "Low" ~ "Low Dose",
                         t_muris_dose == "High" ~ "High Dose"))) %>%
  select(all_of(cbc_percent), 
         group,
         strain,
         lab_or_rw,
         t_muris_dose) %>%
  mutate(species = "Mouse")
  
human_cbc <- cleaned_nhanes_data_2021_23 %>%
  select(all_of(cbc_percent)) %>%
  mutate(species = "human",
         group = "Human")

human_mice_cbc <- bind_rows(mouse_cbc,
                            human_cbc) %>%
  mutate(group = factor(group,
                        levels = c("Human",
                                   "C57BL/6J Lab Zero Dose",
                                   "C57BL/6J Lab Low Dose",
                                   "C57BL/6J Lab High Dose",
                                   "C57BL/6J RW Zero Dose",
                                   "C57BL/6J RW Low Dose",
                                   "C57BL/6J RW High Dose",
                                   "B10.BR Lab Zero Dose",
                                   "B10.BR Lab Low Dose",
                                   "B10.BR Lab High Dose",
                                   "B10.BR RW Zero Dose",
                                   "B10.BR RW Low Dose",
                                   "B10.BR RW High Dose")))

lym_lm <- lm(lym_percent ~ group,
             data = human_mice_cbc)

tab_model(lym_lm,
          show.stat = TRUE)

lym_emm <- emmeans(lym_lm, 
                   specs = "group")

kable(pairs(lym_emm),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

```
```{r}

neu_lm <- lm(neu_percent ~ group,
             data = human_mice_cbc)

tab_model(neu_lm,
          show.stat = TRUE)

neu_emm <- emmeans(neu_lm, 
                   specs = "group")

kable(pairs(neu_emm),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

```

```{r}

mon_lm <- lm(mon_percent ~ group,
             data = human_mice_cbc)

tab_model(mon_lm,
          show.stat = TRUE)

mon_emm <- emmeans(mon_lm,
                   specs = "group")

kable(pairs(mon_emm),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

```

```{r}

eos_lm <- lm(eos_percent ~ group,
             data = human_mice_cbc)

tab_model(eos_lm,
          show.stat = TRUE)

eos_emm <- emmeans(eos_lm,
                   specs = "group")

kable(pairs(eos_emm),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

```

```{r}

bas_lm <- lm(bas_percent ~ group,
             data = human_mice_cbc)

tab_model(bas_lm,
          show.stat = TRUE)

bas_emm <- emmeans(bas_lm,
                   specs = "group")

kable(pairs(bas_emm),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

```

# Plotting the data as it is, looking at over time changes.

```{r}

lym_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(cell_type == "lym")

standard_error_lym <- lym_plot_data %>%
  group_by(phase, t_muris_dose, strain, lab_or_rw) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n),
            .groups = "drop")

lym_plot <- cbc_plotter(lym_plot_data,
                        standard_error_lym,
                        expression(bold("Lymphocyte Count (10"^3~"cells/"*mu*"L)")),
                        "Lymphocyte Count Over Time")

lym_plot

plot_saver("lym_number.png", plot = lym_plot)
svg_saver("lym_number.svg", plot = lym_plot)

```

```{r}

neu_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(cell_type == "neu")

standard_error_neu <- neu_plot_data %>%
  group_by(phase, t_muris_dose, strain, lab_or_rw) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n),
            .groups = "drop")

neu_plot <- cbc_plotter(neu_plot_data,
                        standard_error_neu,
                        expression(bold("Neutrophil Count (10"^3~"cells/"*mu*"L)")),
                        "Neutrophil Count Over Time")

neu_plot

plot_saver("neu_number.png", plot = neu_plot)
svg_saver("neu_number.svg", plot = neu_plot)

```

```{r}

mon_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(cell_type == "mon")

standard_error_mon <- mon_plot_data %>%
  group_by(phase, t_muris_dose, strain, lab_or_rw) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n),
            .groups = "drop")

mon_plot <- cbc_plotter(mon_plot_data,
                        standard_error_mon,
                        expression(bold("Monocyte Count (10"^3~"cells/"*mu*"L)")),
                        "Monocyte Count Over Time")

mon_plot

plot_saver("mon_number.png", plot = mon_plot)
svg_saver("mon_number.svg", plot = mon_plot)

```

```{r}

eos_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(cell_type == "eos")

standard_error_eos <- eos_plot_data %>%
  group_by(phase, t_muris_dose, strain, lab_or_rw) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n),
            .groups = "drop")

eos_plot <- cbc_plotter(eos_plot_data,
                        standard_error_eos,
                        expression(bold("Eosinophil Count (10"^3~"cells/"*mu*"L)")),
                        "Eosinophil Count Over Time")

eos_plot

plot_saver("eos_number.png", plot = eos_plot)
svg_saver("eos_number.svg", plot = eos_plot)

```

```{r}

bas_plot_data <- cleaned_combined_cbc_data_long %>%
  filter(cell_type == "bas")

standard_error_bas <- bas_plot_data %>%
  group_by(phase, t_muris_dose, strain, lab_or_rw) %>%
  summarise(mean = mean(number),
            sd = sd(number),
            n = n(),
            se = sd / sqrt(n),
            .groups = "drop")

bas_plot <- cbc_plotter(bas_plot_data,
                        standard_error_bas,
                        expression(bold("Basophil Count (10"^3~"cells/"*mu*"L)")),
                        "Basophil Count Over Time")

bas_plot

plot_saver("bas_number.png", plot = bas_plot)
svg_saver("bas_number.svg", plot = bas_plot)

```

# Phase-split models and EMM plots for each CBC readout

## I'm thinking that since the pre-rw phase is where everyone is still just lab,
## there is no environment effect yet — so lab_or_rw is excluded.
## We still include t_muris_dose * strain so that the emmeans reflect the
## baseline differences between groups that will later receive different doses.
## Similarly, during-rw we add lab_or_rw since mice are now in different
## environments, giving us the full lab_or_rw * t_muris_dose * strain.
## Post-rw is the same full three-way interaction.

## We will also remove the mouse random effect since we're not sampling the same
## mouse over and over.

## Neutrophil

```{r}

# Pre-RW: dose x strain (no environment — all mice are lab-housed)
neu_pre <- lmer(log(neu_number) ~ t_muris_dose * strain +
                  (1 | cage_wedge) + (1 | cohort),
                data = filter(cleaned_combined_cbc_data,
                              phase == "Pre-RW"))

tab_model(neu_pre,
          show.stat = TRUE)
# statistic is t-value btw
summary(neu_pre)
neu_pre_sim <- simulateResiduals(fittedModel = neu_pre, n = 1000)
plot(neu_pre_sim)
testUniformity(neu_pre_sim)
testDispersion(neu_pre_sim)
testOutliers(neu_pre_sim)
plotResiduals(neu_pre_sim, form = filter(cleaned_combined_cbc_data, 
                                         phase == "Pre-RW")$t_muris_dose)
plotResiduals(neu_pre_sim, form = filter(cleaned_combined_cbc_data, 
                                         phase == "Pre-RW")$strain)

# During RW: environment x dose x strain
neu_during <- lmer(log(neu_number) ~ lab_or_rw * t_muris_dose * strain +
                     (1 | cage_wedge) + (1 | cohort),
                   data = filter(cleaned_combined_cbc_data, phase == "During RW"))

tab_model(neu_during, show.stat = TRUE)
# statistic is t-value btw
summary(neu_during)
neu_during_sim <- simulateResiduals(fittedModel = neu_during, n = 1000)
plot(neu_during_sim)
testUniformity(neu_during_sim)
testDispersion(neu_during_sim)
testOutliers(neu_during_sim)
plotResiduals(neu_during_sim, form = filter(cleaned_combined_cbc_data, 
                                            phase == "During RW")$lab_or_rw)
plotResiduals(neu_during_sim, form = filter(cleaned_combined_cbc_data, 
                                            phase == "During RW")$t_muris_dose)
plotResiduals(neu_during_sim, form = filter(cleaned_combined_cbc_data, 
                                            phase == "During RW")$strain)

# Post-RW: environment x infection x strain
neu_post <- lmer(log(neu_number) ~ lab_or_rw * t_muris_dose * strain +
                   (1 | cage_wedge) + (1 | cohort),
                 data = filter(cleaned_combined_cbc_data, phase == "Post-RW"))

tab_model(neu_post, show.stat = TRUE)
# statistic is t-value btw
summary(neu_post)
neu_post_sim <- simulateResiduals(fittedModel = neu_post, n = 1000)
plot(neu_post_sim)
testUniformity(neu_post_sim)
testDispersion(neu_post_sim)
testOutliers(neu_post_sim)
plotResiduals(neu_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$lab_or_rw)
plotResiduals(neu_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$t_muris_dose)
plotResiduals(neu_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$strain)

```

```{r}
# Pre-RW (no environment effect yet)
emm_neu_pre <- emmeans(neu_pre,
                       specs = "t_muris_dose",
                       by = "strain",
                       type = "response")

pairs(emm_neu_pre)

kable(pairs(emm_neu_pre),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_neu_during <- emmeans(neu_during,
                          specs = c("t_muris_dose", "lab_or_rw"),
                          by = "strain",
                          type = "response")

pairs(emm_neu_during)

kable(pairs(emm_neu_during),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_neu_post <- emmeans(neu_post,
                        specs = c("t_muris_dose", "lab_or_rw"),
                        by = "strain",
                        type = "response")

pairs(emm_neu_post)

kable(pairs(emm_neu_post),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

```

```{r}

emm_neu_pre_df <- as.data.frame(emm_neu_pre) %>%
  crossing(lab_or_rw = factor(c("Lab", 
                                "RW"),
                              levels = c("Lab", 
                                         "RW"))) %>%
  mutate(phase = factor("Pre-RW",
                        levels = c("Pre-RW", 
                                   "During RW", 
                                   "Post-RW")))

emm_neu_during_df <- as.data.frame(emm_neu_during) %>%
  mutate(phase = factor("During RW", 
                        levels = c("Pre-RW", 
                                   "During RW", 
                                   "Post-RW")))

emm_neu_post_df <- as.data.frame(emm_neu_post) %>%
  mutate(phase = factor("Post-RW", 
                        levels = c("Pre-RW", 
                                   "During RW", 
                                   "Post-RW")))

# Combine
emm_neu_combined <- bind_rows(emm_neu_pre_df, 
                              emm_neu_during_df, 
                              emm_neu_post_df) %>%
  mutate(phase = factor(phase, 
                        levels = c("Pre-RW", 
                                   "During RW", 
                                   "Post-RW")),
         t_muris_dose = factor(t_muris_dose, 
                               levels = c("Zero",
                                          "Low", 
                                          "High")),
         strain = factor(strain, 
                         levels = c("C57BL/6J",
                                    "B10.BR")),
         lab_or_rw = factor(lab_or_rw, 
                            levels = c("Lab", 
                                       "RW")))

```


```{r}

emm_neu_plot <- emm_plotter(
  emm_df = emm_neu_combined,
  cbc_input = cleaned_combined_cbc_data,
  y_col = "neu_number",
  y_label = expression(bold("Neutrophil Count (10"^3~"cells/"*mu*"L)")),
  title = "Marginal Means of Neutrophil Count (95% CI)"
)

emm_neu_plot

plot_saver("emm_neu_number.png", emm_neu_plot)
svg_saver("emm_neu_number.svg", emm_neu_plot)

```

## Lymphocyte

```{r}

# Pre-RW: dose x strain (no environment — all mice are lab-housed)
lym_pre <- lmer(log(lym_number) ~ t_muris_dose * strain +
                  (1 | cage_wedge) + (1 | cohort),
                data = filter(cleaned_combined_cbc_data, phase == "Pre-RW"))

tab_model(lym_pre, show.stat = TRUE)
# statistic is t-value btw
summary(lym_pre)
lym_pre_sim <- simulateResiduals(fittedModel = lym_pre, n = 1000)
plot(lym_pre_sim)
testUniformity(lym_pre_sim)
testDispersion(lym_pre_sim)
testOutliers(lym_pre_sim)
plotResiduals(lym_pre_sim, form = filter(cleaned_combined_cbc_data, phase == "Pre-RW")$t_muris_dose)
plotResiduals(lym_pre_sim, form = filter(cleaned_combined_cbc_data, phase == "Pre-RW")$strain)

# During RW: environment x dose x strain
lym_during <- lmer(log(lym_number) ~ lab_or_rw * t_muris_dose * strain +
                     (1 | cage_wedge) + (1 | cohort),
                   data = filter(cleaned_combined_cbc_data, phase == "During RW"))

tab_model(lym_during, show.stat = TRUE)
# statistic is t-value btw
summary(lym_during)
lym_during_sim <- simulateResiduals(fittedModel = lym_during, n = 1000)
plot(lym_during_sim)
testUniformity(lym_during_sim)
testDispersion(lym_during_sim)
testOutliers(lym_during_sim)
plotResiduals(lym_during_sim, form = filter(cleaned_combined_cbc_data, phase == "During RW")$lab_or_rw)
plotResiduals(lym_during_sim, form = filter(cleaned_combined_cbc_data, phase == "During RW")$t_muris_dose)
plotResiduals(lym_during_sim, form = filter(cleaned_combined_cbc_data, phase == "During RW")$strain)

# Post-RW: environment x infection x strain
lym_post <- lmer(log(lym_number) ~ lab_or_rw * t_muris_dose * strain +
                   (1 | cage_wedge) + (1 | cohort),
                 data = filter(cleaned_combined_cbc_data, phase == "Post-RW"))

tab_model(lym_post, show.stat = TRUE)
# statistic is t-value btw
summary(lym_post)
lym_post_sim <- simulateResiduals(fittedModel = lym_post, n = 1000)
plot(lym_post_sim)
testUniformity(lym_post_sim)
testDispersion(lym_post_sim)
testOutliers(lym_post_sim)
plotResiduals(lym_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$lab_or_rw)
plotResiduals(lym_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$t_muris_dose)
plotResiduals(lym_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$strain)

```

```{r}

# Pre-RW (no environment effect yet)
emm_lym_pre <- emmeans(lym_pre,
                       specs = "t_muris_dose",
                       by = "strain",
                       type = "response")

pairs(emm_lym_pre)

kable(pairs(emm_lym_pre),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_lym_during <- emmeans(lym_during,
                          specs = c("t_muris_dose", "lab_or_rw"),
                          by = "strain",
                          type = "response")

pairs(emm_lym_during)

kable(pairs(emm_lym_during),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_lym_post <- emmeans(lym_post,
                        specs = c("t_muris_dose", "lab_or_rw"),
                        by = "strain",
                        type = "response")

pairs(emm_lym_post)

kable(pairs(emm_lym_post),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

```

```{r}

emm_lym_pre_df <- as.data.frame(emm_lym_pre) %>%
  crossing(lab_or_rw = factor(c("Lab",
                                "RW"),
                              levels = c("Lab",
                                         "RW"))) %>%
  mutate(phase = factor("Pre-RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

emm_lym_during_df <- as.data.frame(emm_lym_during) %>%
  mutate(phase = factor("During RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

emm_lym_post_df <- as.data.frame(emm_lym_post) %>%
  mutate(phase = factor("Post-RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

# Combine
emm_lym_combined <- bind_rows(emm_lym_pre_df,
                              emm_lym_during_df,
                              emm_lym_post_df) %>%
  mutate(phase = factor(phase,
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")),
         t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero",
                                          "Low",
                                          "High")),
         strain = factor(strain,
                         levels = c("C57BL/6J",
                                    "B10.BR")),
         lab_or_rw = factor(lab_or_rw,
                            levels = c("Lab",
                                       "RW")))

```

```{r}

emm_lym_plot <- emm_plotter(
  emm_df = emm_lym_combined,
  cbc_input = cleaned_combined_cbc_data,
  y_col = "lym_number",
  y_label = expression(bold("Lymphocyte Count (10"^3~"cells/"*mu*"L)")),
  title = "Marginal Means of Lymphocyte Count (95% CI)"
)

emm_lym_plot

plot_saver("emm_lym_number.png", emm_lym_plot)
svg_saver("emm_lym_number.svg", emm_lym_plot)

```

## Monocyte

```{r}

# Pre-RW: dose x strain (no environment — all mice are lab-housed)
mon_pre <- lmer(log(mon_number) ~ t_muris_dose * strain +
                  (1 | cage_wedge) + (1 | cohort),
                data = filter(cleaned_combined_cbc_data, phase == "Pre-RW"))

tab_model(mon_pre, show.stat = TRUE)
# statistic is t-value btw
summary(mon_pre)
mon_pre_sim <- simulateResiduals(fittedModel = mon_pre, n = 1000)
plot(mon_pre_sim)
testUniformity(mon_pre_sim)
testDispersion(mon_pre_sim)
testOutliers(mon_pre_sim)
plotResiduals(mon_pre_sim, form = filter(cleaned_combined_cbc_data, phase == "Pre-RW")$t_muris_dose)
plotResiduals(mon_pre_sim, form = filter(cleaned_combined_cbc_data, phase == "Pre-RW")$strain)

# During RW: environment x dose x strain
mon_during <- lmer(log(mon_number) ~ lab_or_rw * t_muris_dose * strain +
                     (1 | cage_wedge) + (1 | cohort),
                   data = filter(cleaned_combined_cbc_data, phase == "During RW"))

tab_model(mon_during, show.stat = TRUE)
# statistic is t-value btw
summary(mon_during)
mon_during_sim <- simulateResiduals(fittedModel = mon_during, n = 1000)
plot(mon_during_sim)
testUniformity(mon_during_sim)
testDispersion(mon_during_sim)
testOutliers(mon_during_sim)
plotResiduals(mon_during_sim, form = filter(cleaned_combined_cbc_data, phase == "During RW")$lab_or_rw)
plotResiduals(mon_during_sim, form = filter(cleaned_combined_cbc_data, phase == "During RW")$t_muris_dose)
plotResiduals(mon_during_sim, form = filter(cleaned_combined_cbc_data, phase == "During RW")$strain)

# Post-RW: environment x infection x strain
mon_post <- lmer(log(mon_number) ~ lab_or_rw * t_muris_dose * strain +
                   (1 | cage_wedge) + (1 | cohort),
                 data = filter(cleaned_combined_cbc_data, phase == "Post-RW"))

tab_model(mon_post, show.stat = TRUE)
# statistic is t-value btw
summary(mon_post)
mon_post_sim <- simulateResiduals(fittedModel = mon_post, n = 1000)
plot(mon_post_sim)
testUniformity(mon_post_sim)
testDispersion(mon_post_sim)
testOutliers(mon_post_sim)
plotResiduals(mon_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$lab_or_rw)
plotResiduals(mon_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$t_muris_dose)
plotResiduals(mon_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$strain)

```

```{r}

# Pre-RW (no environment effect yet)
emm_mon_pre <- emmeans(mon_pre,
                       specs = "t_muris_dose",
                       by = "strain",
                       type = "response")

pairs(emm_mon_pre)

kable(pairs(emm_mon_pre),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_mon_during <- emmeans(mon_during,
                          specs = c("t_muris_dose", "lab_or_rw"),
                          by = "strain",
                          type = "response")

pairs(emm_mon_during)

kable(pairs(emm_mon_during),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_mon_post <- emmeans(mon_post,
                        specs = c("t_muris_dose", "lab_or_rw"),
                        by = "strain",
                        type = "response")

pairs(emm_mon_post)

kable(pairs(emm_mon_post),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

```

```{r}

emm_mon_pre_df <- as.data.frame(emm_mon_pre) %>%
  crossing(lab_or_rw = factor(c("Lab",
                                "RW"),
                              levels = c("Lab",
                                         "RW"))) %>%
  mutate(phase = factor("Pre-RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

emm_mon_during_df <- as.data.frame(emm_mon_during) %>%
  mutate(phase = factor("During RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

emm_mon_post_df <- as.data.frame(emm_mon_post) %>%
  mutate(phase = factor("Post-RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

# Combine
emm_mon_combined <- bind_rows(emm_mon_pre_df,
                              emm_mon_during_df,
                              emm_mon_post_df) %>%
  mutate(phase = factor(phase,
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")),
         t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero",
                                          "Low",
                                          "High")),
         strain = factor(strain,
                         levels = c("C57BL/6J",
                                    "B10.BR")),
         lab_or_rw = factor(lab_or_rw,
                            levels = c("Lab",
                                       "RW")))

```

```{r}

emm_mon_plot <- emm_plotter(
  emm_df = emm_mon_combined,
  cbc_input = cleaned_combined_cbc_data,
  y_col = "mon_number",
  y_label = expression(bold("Monocyte Count (10"^3~"cells/"*mu*"L)")),
  title = "Marginal Means of Monocyte Count (95% CI)"
)

emm_mon_plot

plot_saver("emm_mon_number.png", emm_mon_plot)
svg_saver("emm_mon_number.svg", emm_mon_plot)

```

## Eosinophil

```{r}

# Pre-RW: dose x strain (log1p for zeros, no environment — all mice are lab-housed)
eos_pre <- lmer(log1p(eos_number) ~ t_muris_dose * strain +
                  (1 | cage_wedge) + (1 | cohort),
                data = filter(cleaned_combined_cbc_data, phase == "Pre-RW"))

tab_model(eos_pre, show.stat = TRUE)
# statistic is t-value btw
summary(eos_pre)
eos_pre_sim <- simulateResiduals(fittedModel = eos_pre, n = 1000)
plot(eos_pre_sim)
testUniformity(eos_pre_sim)
testDispersion(eos_pre_sim)
testOutliers(eos_pre_sim)
plotResiduals(eos_pre_sim, form = filter(cleaned_combined_cbc_data, phase == "Pre-RW")$t_muris_dose)
plotResiduals(eos_pre_sim, form = filter(cleaned_combined_cbc_data, phase == "Pre-RW")$strain)

# During RW: environment x dose x strain
eos_during <- lmer(log1p(eos_number) ~ lab_or_rw * t_muris_dose * strain +
                     (1 | cage_wedge) + (1 | cohort),
                   data = filter(cleaned_combined_cbc_data, phase == "During RW"))

tab_model(eos_during, show.stat = TRUE)
# statistic is t-value btw
summary(eos_during)
eos_during_sim <- simulateResiduals(fittedModel = eos_during, n = 1000)
plot(eos_during_sim)
testUniformity(eos_during_sim)
testDispersion(eos_during_sim)
testOutliers(eos_during_sim)
plotResiduals(eos_during_sim, form = filter(cleaned_combined_cbc_data, phase == "During RW")$lab_or_rw)
plotResiduals(eos_during_sim, form = filter(cleaned_combined_cbc_data, phase == "During RW")$t_muris_dose)
plotResiduals(eos_during_sim, form = filter(cleaned_combined_cbc_data, phase == "During RW")$strain)

# Post-RW: environment x infection x strain
eos_post <- lmer(log1p(eos_number) ~ lab_or_rw * t_muris_dose * strain +
                   (1 | cage_wedge) + (1 | cohort),
                 data = filter(cleaned_combined_cbc_data, phase == "Post-RW"))

tab_model(eos_post, show.stat = TRUE)
# statistic is t-value btw
summary(eos_post)
eos_post_sim <- simulateResiduals(fittedModel = eos_post, n = 1000)
plot(eos_post_sim)
testUniformity(eos_post_sim)
testDispersion(eos_post_sim)
testOutliers(eos_post_sim)
plotResiduals(eos_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$lab_or_rw)
plotResiduals(eos_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$t_muris_dose)
plotResiduals(eos_post_sim, form = filter(cleaned_combined_cbc_data, phase == "Post-RW")$strain)

```

```{r}

# Pre-RW (no environment effect yet)
emm_eos_pre <- emmeans(eos_pre,
                       specs = "t_muris_dose",
                       by = "strain")

pairs(emm_eos_pre)

kable(pairs(emm_eos_pre),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_eos_during <- emmeans(eos_during,
                          specs = c("t_muris_dose", "lab_or_rw"),
                          by = "strain")

pairs(emm_eos_during)

kable(pairs(emm_eos_during),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_eos_post <- emmeans(eos_post,
                        specs = c("t_muris_dose", "lab_or_rw"),
                        by = "strain")

pairs(emm_eos_post)

kable(pairs(emm_eos_post),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

```

```{r}

# log1p back-transform: expm1() reverses log1p()
emm_eos_pre_df <- as.data.frame(emm_eos_pre) %>%
  mutate(response = expm1(emmean),
         lower.CL = expm1(lower.CL),
         upper.CL = expm1(upper.CL)) %>%
  crossing(lab_or_rw = factor(c("Lab",
                                "RW"),
                              levels = c("Lab",
                                         "RW"))) %>%
  mutate(phase = factor("Pre-RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

emm_eos_during_df <- as.data.frame(emm_eos_during) %>%
  mutate(response = expm1(emmean),
         lower.CL = expm1(lower.CL),
         upper.CL = expm1(upper.CL)) %>%
  mutate(phase = factor("During RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

emm_eos_post_df <- as.data.frame(emm_eos_post) %>%
  mutate(response = expm1(emmean),
         lower.CL = expm1(lower.CL),
         upper.CL = expm1(upper.CL)) %>%
  mutate(phase = factor("Post-RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

# Combine
emm_eos_combined <- bind_rows(emm_eos_pre_df,
                              emm_eos_during_df,
                              emm_eos_post_df) %>%
  mutate(phase = factor(phase,
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")),
         t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero",
                                          "Low",
                                          "High")),
         strain = factor(strain,
                         levels = c("C57BL/6J",
                                    "B10.BR")),
         lab_or_rw = factor(lab_or_rw,
                            levels = c("Lab",
                                       "RW")))

```

```{r}

emm_eos_plot <- emm_plotter(
  emm_df = emm_eos_combined,
  cbc_input = cleaned_combined_cbc_data,
  y_col = "eos_number",
  y_label = expression(bold("Eosinophil Count (10"^3~"cells/"*mu*"L)")),
  title = "Marginal Means of Eosinophil Count (95% CI)",
  floor_zero = TRUE
)

emm_eos_plot

plot_saver("emm_eos_number.png", emm_eos_plot)
svg_saver("emm_eos_number.svg", emm_eos_plot)

```

## Basophil

```{r}

# Pre-RW: dose x strain (log1p for zeros, no environment — all mice are lab-housed)
bas_pre <- lmer(log1p(bas_number) ~ t_muris_dose * strain +
                  (1 | cage_wedge) + (1 | cohort),
                data = filter(cleaned_combined_cbc_data, phase == "Pre-RW"))

tab_model(bas_pre, show.stat = TRUE)
# statistic is t-value btw
summary(bas_pre)
bas_pre_sim <- simulateResiduals(fittedModel = bas_pre, n = 1000)
plot(bas_pre_sim)
testUniformity(bas_pre_sim)
testDispersion(bas_pre_sim)
testOutliers(bas_pre_sim)
plotResiduals(bas_pre_sim, form = filter(cleaned_combined_cbc_data, phase == "Pre-RW")$t_muris_dose)
plotResiduals(bas_pre_sim, form = filter(cleaned_combined_cbc_data, phase == "Pre-RW")$strain)

# During RW: environment x dose x strain
bas_during <- lmer(log1p(bas_number) ~ lab_or_rw * t_muris_dose * strain +
                     (1 | cage_wedge) + (1 | cohort),
                   data = filter(cleaned_combined_cbc_data, phase == "During RW"))

tab_model(bas_during, show.stat = TRUE)
# statistic is t-value btw
summary(bas_during)
bas_during_sim <- simulateResiduals(fittedModel = bas_during, n = 1000)
plot(bas_during_sim)
testUniformity(bas_during_sim)
testDispersion(bas_during_sim)
testOutliers(bas_during_sim)
plotResiduals(bas_during_sim, form = filter(cleaned_combined_cbc_data, 
                                            phase == "During RW")$lab_or_rw)
plotResiduals(bas_during_sim, form = filter(cleaned_combined_cbc_data, 
                                            phase == "During RW")$t_muris_dose)
plotResiduals(bas_during_sim, form = filter(cleaned_combined_cbc_data, 
                                            phase == "During RW")$strain)

# Post-RW: environment x infection x strain
bas_post <- lmer(log1p(bas_number) ~ lab_or_rw * t_muris_dose * strain +
                   (1 | cage_wedge) + (1 | cohort),
                 data = filter(cleaned_combined_cbc_data, phase == "Post-RW"))

tab_model(bas_post, show.stat = TRUE)
# statistic is t-value btw
summary(bas_post)
bas_post_sim <- simulateResiduals(fittedModel = bas_post, n = 1000)
plot(bas_post_sim)
testUniformity(bas_post_sim)
testDispersion(bas_post_sim)
testOutliers(bas_post_sim)
plotResiduals(bas_post_sim, form = filter(cleaned_combined_cbc_data, 
                                          phase == "Post-RW")$lab_or_rw)
plotResiduals(bas_post_sim, form = filter(cleaned_combined_cbc_data, 
                                          phase == "Post-RW")$t_muris_dose)
plotResiduals(bas_post_sim, form = filter(cleaned_combined_cbc_data, 
                                          phase == "Post-RW")$strain)

```

```{r}

# Pre-RW (no environment effect yet)
emm_bas_pre <- emmeans(bas_pre,
                       specs = "t_muris_dose",
                       by = "strain")

pairs(emm_bas_pre)

kable(pairs(emm_bas_pre),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_bas_during <- emmeans(bas_during,
                          specs = c("t_muris_dose", "lab_or_rw"),
                          by = "strain")

pairs(emm_bas_during)

kable(pairs(emm_bas_during),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_bas_post <- emmeans(bas_post,
                        specs = c("t_muris_dose", "lab_or_rw"),
                        by = "strain")

pairs(emm_bas_post)

kable(pairs(emm_bas_post),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

```

```{r}

# log1p back-transform: expm1() reverses log1p()
emm_bas_pre_df <- as.data.frame(emm_bas_pre) %>%
  mutate(response = expm1(emmean),
         lower.CL = expm1(lower.CL),
         upper.CL = expm1(upper.CL)) %>%
  crossing(lab_or_rw = factor(c("Lab",
                                "RW"),
                              levels = c("Lab",
                                         "RW"))) %>%
  mutate(phase = factor("Pre-RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

emm_bas_during_df <- as.data.frame(emm_bas_during) %>%
  mutate(response = expm1(emmean),
         lower.CL = expm1(lower.CL),
         upper.CL = expm1(upper.CL)) %>%
  mutate(phase = factor("During RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

emm_bas_post_df <- as.data.frame(emm_bas_post) %>%
  mutate(response = expm1(emmean),
         lower.CL = expm1(lower.CL),
         upper.CL = expm1(upper.CL)) %>%
  mutate(phase = factor("Post-RW",
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")))

# Combine
emm_bas_combined <- bind_rows(emm_bas_pre_df,
                              emm_bas_during_df,
                              emm_bas_post_df) %>%
  mutate(phase = factor(phase,
                        levels = c("Pre-RW",
                                   "During RW",
                                   "Post-RW")),
         t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero",
                                          "Low",
                                          "High")),
         strain = factor(strain,
                         levels = c("C57BL/6J",
                                    "B10.BR")),
         lab_or_rw = factor(lab_or_rw,
                            levels = c("Lab",
                                       "RW")))

```

```{r}

emm_bas_plot <- emm_plotter(
  emm_df = emm_bas_combined,
  cbc_input = cleaned_combined_cbc_data,
  y_col = "bas_number",
  y_label = expression(bold("Basophil Count (10"^3~"cells/"*mu*"L)")),
  title = "Marginal Means of Basophil Count (95% CI)",
  floor_zero = TRUE
)

emm_bas_plot

plot_saver("emm_bas_number.png", emm_bas_plot)
svg_saver("emm_bas_number.svg", emm_bas_plot)

```
