---
title: "ELISA analysis"
author: "Zehao Wu"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

# First time run order: Spleen → CBC → ELISA
# Requires clean_meta_data.csv from the spleen analysis file.

# Loading the libraries

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(janitor)
library(factoextra)
library(skimr)
library(naniar)
library(ggplot2)
library(ggpubr)
library(emmeans)
library(lme4)
library(lmerTest)
library(scales)
library(haven)
library(broom.mixed)
library(readr)
library(here)
library(DHARMa) 
library(rstatix)
library(kableExtra)
library(sjPlot) # does not work when knitting to PDF

```

# Setting the paths

```{r}

base_path <- "/Users/zehao/Documents/Research Projects/Graham Group Thesis Research (Summer 2025)"

data_path <- file.path(base_path, "Metadata", "elisadata")

clean_data_path <- file.path(base_path, "Cleaned data")

figure_path       <- file.path(base_path, "Figures")
elisa_figure_path <- file.path(figure_path, "ELISA Figures")
dir.create(elisa_figure_path, showWarnings = FALSE, recursive = TRUE)

```

# Setting up the functions for repeated codes

# Mean calculator

```{r}

mean_calculator <- function(input_data, 
                            final_concentration, 
                            cv_data,
                            groupings = c("mouse_id",
                                          "treatment", 
                                          "plate", 
                                          "strain", 
                                          "t_muris_dose",
                                          "lab_or_rw",
                                          "cohort",
                                          "experiment",
                                          "source",
                                          "experiment_day")) {
input_data %>%
  group_by(across(all_of(groupings))) %>% 
  summarize(
    mean_concentration = mean({{final_concentration}}, 
                    na.rm = TRUE),
    cv_percent = first(na.omit({{cv_data}})),
    .groups = "drop"
  )}

```

# Setting the colors (Okabe-Ito palette and Paul Tol's Bright)

```{r}

color_black <- "#000000"
color_mustard <- "#E69F00"
color_light_blue <- "#56B4E9"
color_green <- "#009E73"
color_yellow <- "#F0E442"
color_dark_blue <- "#0072B2"
color_orange <- "#D55E00"
color_purple <- "#CC79A7"

color_grey <- "#BBBBBB"
color_dark_purple <- "#AA3377"
color_light_purple <- "#EE6677"
color_sand <- "#CCBB44"
color_cyan <- "#66CCEE"
color_olive <- "#228833"
color_navy <- "#4477AA"

```

# Basic elisa summary plotter

```{r}

elisa_summary_plotter <- function(input_data,
                                  y_label,
                                  title = NULL) {
  ggplot(input_data %>% mutate(treatment = factor(treatment, levels = c("IgG", "LPS", "CD28")),
                               lab_or_rw = factor(lab_or_rw, levels = c("Lab", "RW")),
                               t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High"))),
         aes(x = lab_or_rw,
             y = mean_concentration,
             fill = t_muris_dose)) +
    geom_boxplot(position = position_dodge(width = 0.75),
               alpha = 0.8,
               outlier.shape = NA,
               width = 0.5,
               color = "black",
               linewidth = 0.6) +
    geom_jitter(position = position_jitterdodge(jitter.width = 0.15,
                                              dodge.width = 0.75),
              alpha = 0.8,
              size = 1.8,
              shape = 21,
              color = "black") +
    facet_wrap(~ treatment,
               labeller = as_labeller(c("CD28" = "CD28SA",
                                        "LPS" = "LPS",
                                        "IgG" = "IgG1"))) +
    scale_fill_manual(values = c("Zero" = color_dark_blue,
                                 "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
    scale_color_manual(values = c("Zero" = color_dark_blue,
                                  "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
    scale_x_discrete(labels = c("Lab" = "Lab",
                                "RW"  = "Rewilded"),
                     name = expression(bold("Environment"))) +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
    ylab(y_label) +
    ggtitle(title)
}

```

# Treatment specific plotter

```{r}

elisa_treatment_plotter <- function(input_data,
                                   y_label,
                                   title = NULL) {
  ggplot(input_data %>% mutate(lab_or_rw = factor(lab_or_rw, levels = c("Lab", "RW")),
                               t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High"))),
         aes(x = lab_or_rw,
             y = mean_concentration,
             fill = t_muris_dose)) +
    geom_boxplot(alpha = 0.8,
                 outlier.shape = NA,
                 width = 0.5,
                 color = "black",
                 linewidth = 0.6,
                 position = position_dodge(width = 0.6)) +
    geom_jitter(alpha = 0.8,
                size = 1.8,
                shape = 21,
                color = "black",
                position = position_jitterdodge(jitter.width = 0.1,
                                                dodge.width = 0.6)) +
    scale_fill_manual(values = c("Zero" = color_dark_blue,
                                 "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
    scale_color_manual(values = c("Zero" = color_dark_blue,
                                  "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
    scale_x_discrete(labels = c("Lab" = "Lab",
                                "RW"  = "Rewilded"),
                     name = expression(bold("Environment"))) +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
    theme_minimal() +
    theme(panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
    ylab(y_label) +
    ggtitle(title)
}


```

# Emmeans per treatment data frame maker
# NOTE: emm_treatment function was removed — it contained a bug where
# filter(treatment == treatment) compared the column to itself (always TRUE).
# The function was never called anywhere, so it was dead code.

# Plot saver

```{r}

plot_saver <- function(filename,
                            plot,
                            path = elisa_figure_path,
                            width = 6, 
                            height = 4, 
                            dpi = 300) {
  ggsave(filename = filename, 
       plot = plot, 
       path = path,
       width = width, 
       height = height, 
       units = "in", 
       dpi = dpi)
}

```

## Loading the data

```{r}

elisa_ifn_1 <- read_xlsx(file.path(data_path, 
                                   "IFNg_Plate1_20260121.xlsx"), 
                         sheet = "Plate 1 Data")

elisa_ifn_2 <- read_xlsx(file.path(data_path,
                                   "IFNg_Plate2_20260123.xlsx"), 
                         sheet = "Plate 2 Data")

elisa_tnf_1 <- read_xlsx(file.path(data_path,
                                   "TNFa_Plate1_20260121.xlsx"), 
                         sheet = "Plate 1 Data")

elisa_tnf_2 <- read_xlsx(file.path(data_path,
                                   "TNFa_Plate2_20260123.xlsx"), 
                         sheet = "Plate 2 Data")

plate1_map <- read_xlsx(file.path(data_path,
                                  "SF ELISA Plate Mapping.xlsx"), 
                        sheet = "Plate 1 for both IFN and TNF")

plate2_map <- read_xlsx(file.path(data_path,
                                  "SF ELISA Plate Mapping.xlsx"), 
                        sheet = "Plate 2 for both IFN and TNF")

meta_data <- read_csv(file.path(clean_data_path,
                                "clean_meta_data.csv"))

```

# data cleaning and organization

# Four original cell culture plates, each stimulated on a different day:
#   Original plate 1 stimulated July 22, 2025 -> experiment_day = 1
#   Original plate 2 stimulated July 23, 2025 -> experiment_day = 2
#   Original plate 3 stimulated July 29, 2025 -> experiment_day = 3
#   Original plate 4 stimulated July 30, 2025 -> experiment_day = 4
# Each mouse appears on exactly one stimulation plate (one experiment_day).
# For ELISA detection, original plates were combined into 2 detection plates:
#   Detection plate 1 = original plates 1 + 2 (July 22 + 23)
#   Detection plate 2 = original plates 3 + 4 (July 29 + 30)
# The detection plate number is coded as `plate` in the cleaning chunks below.
# experiment_day (1-4) must be added as a column in the plate map Excel files
# so that it joins in automatically via left_join(plate1_map/plate2_map).

# ifn data cleaning for plate 1

```{r}

clean_ifn_1 <- elisa_ifn_1 %>%
  clean_names() %>% 
  rename("final_ifn" = "final_if_ng") %>% # cleaner names
  left_join(plate1_map,  by = "well_id") %>% # join elisa data and mouse number using plate number
  mutate(plate = 1) %>% # create a new column to signal that this is plate 1
  left_join(meta_data, by = "mouse_id") %>% # add the rest of the meta data
  mutate(
    across(c(concentration, 
             final_ifn, mean, 
             std_dev, 
             cv_percent),
            # apply the same transformation across these columns
           ~ case_when(
              # ~ is like a mini function. case when is checking for a condition and applying the first operation if the first condition matches.
             str_detect(., "^<") ~ 0,
             # . is the current value in the column and the rest of the function is asking if this value starts with a <. If yes, then replace it with 0. This is just turning all the <0.0000 values into actual numbers of 0.
             TRUE ~ as.numeric(str_remove(., "^>"))
           )) # the TRUE means that if none of the previous conditinos matched, then do these. Same logic applies where we remove > from values if present. The as numeric converts the text to number.
  ) %>%
  drop_na(mouse_id)

```

# ifn data cleaning for plate 2

```{r}

clean_ifn_2 <- elisa_ifn_2 %>% #same as before...
  clean_names() %>%
  rename("final_ifn" = "final_if_ng") %>%
  left_join(plate2_map,  by = "well_id") %>%
  mutate(plate = 2) %>%
  left_join(meta_data, by = "mouse_id") %>%
   mutate(
    across(c(concentration, final_ifn, mean, std_dev, cv_percent),
           ~ case_when(
             str_detect(., "^<") ~ 0,
             TRUE ~ as.numeric(str_remove(., "^>"))
           ))
  ) %>%
  drop_na(mouse_id)


total_ifn <- bind_rows(clean_ifn_1, 
                       clean_ifn_2)

write_csv(total_ifn, file.path(clean_data_path, "combined_ifn_elisa_data.csv"))

```

For now, I just kept all of the values that are over the detection limit 
without their ">" and all of the small values as zero. Discuss with David.

# tnf data cleaning for plate 1

```{r}

clean_tnf_1 <- elisa_tnf_1 %>% #same as before...
  clean_names() %>%
  rename("final_tnf" = "concentration") %>%
  left_join(plate1_map,  by = "well_id") %>%
  mutate(plate = 1) %>%
  left_join(meta_data, by = "mouse_id") %>%
   mutate(
    across(c(final_tnf, mean, std_dev, cv_percent),
           ~ case_when(
             str_detect(., "^<") ~ 0,
             TRUE ~ as.numeric(str_remove(., "^>"))
           ))
  ) %>%
  drop_na(mouse_id)

```

# tnf data cleaning for plate 2

```{r}

clean_tnf_2 <- elisa_tnf_2 %>% #same as before...
  clean_names() %>%
  rename("final_tnf" = "concentration") %>%
  left_join(plate2_map,  by = "well_id") %>%
  mutate(plate = 2) %>%
  left_join(meta_data, by = "mouse_id") %>%
   mutate(
    across(c(final_tnf, mean, std_dev, cv_percent),
           ~ case_when(
             str_detect(., "^<") ~ 0,
             TRUE ~ as.numeric(str_remove(., "^>"))
           ))
  ) %>%
  drop_na(mouse_id)

total_tnf <- bind_rows(clean_tnf_1, 
                       clean_tnf_2)

write_csv(total_tnf, file.path(clean_data_path, 
                               "combined_tnf_elisa_data.csv"))

```

# Calculating the mean concentration value

```{r}

ifn_input <- mean_calculator(total_ifn,
                             final_ifn,
                             cv_percent) %>%
  mutate(treatment = factor(treatment, levels = c("CD28", "LPS", "IgG")))

tnf_input <- mean_calculator(total_tnf,
                             final_tnf,
                             cv_percent) %>%
  mutate(treatment = factor(treatment, levels = c("CD28", "LPS", "IgG")))

```

I wonder if I should use the original data points instead of the mean data.

# Filter the data for treatment specific plots

```{r}

ifn_input_cd28 <- ifn_input %>% filter(treatment == "CD28")

ifn_input_lps <- ifn_input %>% filter(treatment == "LPS")

ifn_input_igg <- ifn_input %>% filter(treatment == "IgG")

tnf_input_cd28 <- tnf_input %>% filter(treatment == "CD28")

tnf_input_lps <- tnf_input %>% filter(treatment == "LPS")

tnf_input_igg <- tnf_input %>% filter(treatment == "IgG")

```

# General ifn plot to get an overview

```{r}

ifn_input <- ifn_input %>%
  mutate(t_muris_dose = factor(t_muris_dose, 
                               levels = c("Zero", "High")))

ifn_plot <- elisa_summary_plotter(ifn_input, 
                                  expression(bold("IFN-" * gamma ~ "(pg mL"^"-1"*")")),
                                  title = expression(bold("IFN-"*gamma~"by Stimulation"))) 

ifn_plot

plot_saver("general_ifn_plot.png",
           ifn_plot)

```

# General tnf plots

```{r}

tnf_input <- tnf_input %>%
  mutate(t_muris_dose = factor(t_muris_dose, 
                               levels = c("Zero", "High")))

tnf_plot <- elisa_summary_plotter(tnf_input, 
                                  expression(bold("TNF-"*alpha ~ "(pg mL"^"-1"*")")),
                                  title = expression(bold("TNF-"*alpha~"by Stimulation")))

tnf_plot

plot_saver("general_tnf_plot.png",
           tnf_plot)

```

# IFN faceted by T. muris dose: treatment on x, environment as color

```{r}

ifn_dose_facet_plot <- ggplot(ifn_input %>%
                                mutate(treatment = factor(treatment,
                                                          levels = c("IgG",
                                                                     "LPS",
                                                                     "CD28"))),
                              aes(x = treatment,
                                  y = mean_concentration,
                                  fill = lab_or_rw)) +
  geom_boxplot(position = position_dodge(width = 0.75),
               alpha = 0.8,
               outlier.shape = NA,
               width = 0.5,
               color = "black",
               linewidth = 0.6) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.15,
                                              dodge.width = 0.75),
              alpha = 0.8,
              size = 1.8,
              shape = 21,
              color = "black") +
  facet_wrap(~ t_muris_dose,
             labeller = as_labeller(c("Zero" = "Zero Dose",
                                      "High" = "High Dose"))) +
  scale_fill_manual(values = c("Lab" = color_mustard,
                               "RW"  = color_green),
                    labels = c("Lab" = "Lab",
                               "RW"  = "Rewilded"),
                    name = expression(bold("Environment"))) +
  scale_color_manual(values = c("Lab" = color_mustard,
                                "RW"  = color_green),
                     labels = c("Lab" = "Lab",
                                "RW"  = "Rewilded"),
                     name = expression(bold("Environment"))) +
  scale_x_discrete(name = expression(bold("Treatment")),
                   labels = c("CD28" = "CD28SA",
                              "LPS" = "LPS",
                              "IgG" = "IgG1")) +
  scale_y_continuous(name = expression(bold("IFN-" * gamma ~
                                              "(pg mL"^"-1" * ")")),
                     expand = expansion(mult = c(0.05, 0.3))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle(expression(bold("IFN-" * gamma ~ "by Treatment and Dose")))

ifn_dose_facet_plot

plot_saver("ifn_dose_facet_plot.png",
           ifn_dose_facet_plot)

```

# TNF faceted by T. muris dose: treatment on x, environment as color

```{r}

tnf_dose_facet_plot <- ggplot(tnf_input %>%
                                mutate(treatment = factor(treatment,
                                                          levels = c("IgG",
                                                                     "LPS",
                                                                     "CD28"))),
                              aes(x = treatment,
                                  y = mean_concentration,
                                  fill = lab_or_rw)) +
  geom_boxplot(position = position_dodge(width = 0.75),
               alpha = 0.8,
               outlier.shape = NA,
               width = 0.5,
               color = "black",
               linewidth = 0.6) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.15,
                                              dodge.width = 0.75),
              alpha = 0.8,
              size = 1.8,
              shape = 21,
              color = "black") +
  facet_wrap(~ t_muris_dose,
             labeller = as_labeller(c("Zero" = "Zero Dose",
                                      "High" = "High Dose"))) +
  scale_fill_manual(values = c("Lab" = color_mustard,
                               "RW"  = color_green),
                    labels = c("Lab" = "Lab",
                               "RW"  = "Rewilded"),
                    name = expression(bold("Environment"))) +
  scale_color_manual(values = c("Lab" = color_mustard,
                                "RW"  = color_green),
                     labels = c("Lab" = "Lab",
                                "RW"  = "Rewilded"),
                     name = expression(bold("Environment"))) +
  scale_x_discrete(name = expression(bold("Treatment")),
                   labels = c("CD28" = "CD28SA",
                              "LPS" = "LPS",
                              "IgG" = "IgG1")) +
  scale_y_continuous(name = expression(bold("TNF-" * alpha ~
                                              "(pg mL"^"-1" * ")")),
                     expand = expansion(mult = c(0.05, 0.3))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle(expression(bold("TNF-" * alpha ~ "by Treatment and Dose")))

tnf_dose_facet_plot

plot_saver("tnf_dose_facet_plot.png",
           tnf_dose_facet_plot)

```

# IFN faceted by environment: treatment on x, dose as color (#3)

```{r}

ifn_env_facet_plot <- ggplot(ifn_input %>%
                               mutate(treatment = factor(treatment,
                                                         levels = c("IgG",
                                                                    "LPS",
                                                                    "CD28")),
                                      lab_or_rw = factor(lab_or_rw,
                                                         levels = c("Lab", "RW")),
                                      t_muris_dose = factor(t_muris_dose,
                                                            levels = c("Zero", "High"))),
                             aes(x = treatment,
                                 y = mean_concentration,
                                 fill = t_muris_dose)) +
  geom_boxplot(position = position_dodge(width = 0.75),
               alpha = 0.8,
               outlier.shape = NA,
               width = 0.5,
               color = "black",
               linewidth = 0.6) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.15,
                                              dodge.width = 0.75),
              alpha = 0.8,
              size = 1.8,
              shape = 21,
              color = "black") +
  facet_wrap(~ lab_or_rw,
             labeller = as_labeller(c("Lab" = "Lab",
                                      "RW"  = "Rewilded"))) +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_x_discrete(name = expression(bold("Treatment")),
                   labels = c("CD28" = "CD28SA",
                              "LPS" = "LPS",
                              "IgG" = "IgG1")) +
  scale_y_continuous(name = expression(bold("IFN-" * gamma ~
                                              "(pg mL"^"-1" * ")")),
                     expand = expansion(mult = c(0.05, 0.3))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle(expression(bold("IFN-" * gamma ~ "by Treatment and Environment")))

ifn_env_facet_plot

plot_saver("ifn_env_facet_plot.png",
           ifn_env_facet_plot)

```

# TNF faceted by environment: treatment on x, dose as color (#3)

```{r}

tnf_env_facet_plot <- ggplot(tnf_input %>%
                               mutate(treatment = factor(treatment,
                                                         levels = c("IgG",
                                                                    "LPS",
                                                                    "CD28")),
                                      lab_or_rw = factor(lab_or_rw,
                                                         levels = c("Lab", "RW")),
                                      t_muris_dose = factor(t_muris_dose,
                                                            levels = c("Zero", "High"))),
                             aes(x = treatment,
                                 y = mean_concentration,
                                 fill = t_muris_dose)) +
  geom_boxplot(position = position_dodge(width = 0.75),
               alpha = 0.8,
               outlier.shape = NA,
               width = 0.5,
               color = "black",
               linewidth = 0.6) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.15,
                                              dodge.width = 0.75),
              alpha = 0.8,
              size = 1.8,
              shape = 21,
              color = "black") +
  facet_wrap(~ lab_or_rw,
             labeller = as_labeller(c("Lab" = "Lab",
                                      "RW"  = "Rewilded"))) +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_x_discrete(name = expression(bold("Treatment")),
                   labels = c("CD28" = "CD28SA",
                              "LPS" = "LPS",
                              "IgG" = "IgG1")) +
  scale_y_continuous(name = expression(bold("TNF-" * alpha ~
                                              "(pg mL"^"-1" * ")")),
                     expand = expansion(mult = c(0.05, 0.3))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle(expression(bold("TNF-" * alpha ~ "by Treatment and Environment")))

tnf_env_facet_plot

plot_saver("tnf_env_facet_plot.png",
           tnf_env_facet_plot)

```

# Treatment specific plots for ifn: CD28

```{r}

ifn_input_cd28 <- ifn_input_cd28 %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High")))

ifn_cd28_plot <- elisa_treatment_plotter(ifn_input_cd28,
                                       expression(bold("IFN-" * gamma ~ "(pg mL"^"-1"*")")),
                                       title = expression(bold("CD28SA Stimulation")))

ifn_cd28_plot

plot_saver("ifn_cd28_treatment_plot.png",
           ifn_cd28_plot)

```

# Treatment specific plots for tnf: CD28

```{r}

tnf_input_cd28 <- tnf_input_cd28 %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High")))

tnf_cd28_plot <- elisa_treatment_plotter(tnf_input_cd28,
                                       expression(bold("TNF-"*alpha ~ "(pg mL"^"-1"*")")),
                                       title = expression(bold("CD28SA Stimulation")))

tnf_cd28_plot

plot_saver("tnf_cd28_treatment_plot.png",
           tnf_cd28_plot)

```

# Treatment specific plots for ifn: LPS

```{r}

ifn_input_lps <- ifn_input_lps %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High")))

ifn_lps_plot <- elisa_treatment_plotter(ifn_input_lps,
                                       expression(bold("IFN-" * gamma ~ "(pg mL"^"-1"*")")),
                                       title = expression(bold("LPS Stimulation")))

ifn_lps_plot

plot_saver("ifn_lps_treatment_plot.png",
           ifn_lps_plot)

```

# Treatment specific plots for tnf: LPS

```{r}

tnf_input_lps <- tnf_input_lps %>%
  mutate(t_muris_dose = factor(t_muris_dose, 
                               levels = c("Zero", "High")))

tnf_lps_plot <- elisa_treatment_plotter(tnf_input_lps,
                                       expression(bold("TNF-"*alpha ~ "(pg mL"^"-1"*")")),
                                       title = expression(bold("LPS Stimulation")))

tnf_lps_plot

plot_saver("tnf_lps_treatment_plot.png",
           tnf_lps_plot)

```

# Treatment specific plots for ifn: IgG

```{r}

ifn_input_igg <- ifn_input_igg %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High")))

ifn_igg_plot <- elisa_treatment_plotter(ifn_input_igg,
                                       expression(bold("IFN-" * gamma ~ "(pg mL"^"-1"*")")),
                                       title = expression(bold("IgG1 Stimulation")))

ifn_igg_plot

plot_saver("ifn_igg_treatment_plot.png",
           ifn_igg_plot)

```

# Treatment specific plots for tnf: IgG

```{r}

tnf_input_igg <- tnf_input_igg %>%
  mutate(t_muris_dose = factor(t_muris_dose, 
                               levels = c("Zero", "High")))

tnf_igg_plot <- elisa_treatment_plotter(tnf_input_igg,
                                       expression(bold("TNF-"*alpha ~ "(pg mL"^"-1"*")")),
                                       title = expression(bold("IgG1 Stimulation")))

tnf_igg_plot

plot_saver("tnf_igg_treatment_plot.png",
           tnf_igg_plot)

```

# Setting reference levels for ifn

# Mixed effect model (gamma) for ifn and diagnostics
# Note to self: should I just assume that all zeros are at the detection level?
# Maybe we can just take half of the detection limit (standard practice)
# Detection level for IFN is 1.797 pg/mL so we use this to handle zero values.

```{r}

ifn_input <- ifn_input %>%
  mutate(
    t_muris_dose = relevel(factor(t_muris_dose), ref = "Zero"),
    treatment = relevel(factor(treatment), ref = "IgG"),
    lab_or_rw = relevel(factor(lab_or_rw), ref = "Lab")
  ) 

ifn_glmer_input <- ifn_input %>%
  mutate(mean_concentration = ifelse(mean_concentration == 0,
                                     1.797 / 2, 
                                     mean_concentration))


ifn_glmer <- glmer(mean_concentration ~ lab_or_rw *
                     t_muris_dose *
                     treatment +
                     (1 | mouse_id) +
                     (1 | experiment_day),
                  family = Gamma(link = "log"),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list(maxfun = 2e5)),
                  data = ifn_glmer_input)

tab_model(ifn_glmer, show.stat = TRUE)
# statistic is t-value btw

```


# Diagnostics for ifn

```{r}

ifn_sim_res <- simulateResiduals(fittedModel = ifn_glmer, n = 1000)

plot(ifn_sim_res)

testUniformity(ifn_sim_res)
testDispersion(ifn_sim_res)
testOutliers(ifn_sim_res)

plotResiduals(ifn_sim_res, form = ifn_glmer_input$lab_or_rw)
plotResiduals(ifn_sim_res, form = ifn_glmer_input$t_muris_dose)
plotResiduals(ifn_sim_res, form = ifn_glmer_input$treatment)

ranef(ifn_glmer)

VarCorr(ifn_glmer)

```

# Run emmeans for ifn

```{r}

emm_ifn <- emmeans(ifn_glmer,
                   specs = c("t_muris_dose","lab_or_rw"),
                   by = c("treatment"),
                   type = "response")
emm_ifn

kable(pairs(emm_ifn))

# Pairwise contrast: Lab vs Rewilded within each treatment
emm_ifn_env <- emmeans(ifn_glmer,
                       specs = "t_muris_dose",
                       by = c("treatment", "lab_or_rw"),
                       type = "response")

kable(pairs(emm_ifn_env))

```

# Doing the same thing here
# The lowest detection level is 3.7 for TNF so halving that.

```{r}

tnf_input <- tnf_input %>%
  mutate(
    t_muris_dose = relevel(factor(t_muris_dose), 
                           ref = "Zero"),
    treatment = relevel(factor(treatment), 
                        ref = "IgG"),
    lab_or_rw = relevel(factor(lab_or_rw), 
                        ref = "Lab")
  )

tnf_glmer_input <- tnf_input %>%
  mutate(mean_concentration = ifelse(mean_concentration == 0,
                                     3.7 / 2, 
                                     mean_concentration))

# we want to set the zero dose, IgG and lab groups to be the basis of all our comparisons.

tnf_glmer <- glmer(mean_concentration ~ lab_or_rw *
                     t_muris_dose *
                     treatment +
                     (1 | mouse_id) +
                     (1 | experiment_day),
                  family = Gamma(link = "log"),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list(maxfun = 2e5)),
                 data = tnf_glmer_input)

tab_model(tnf_glmer, show.stat = TRUE)
# statistic is t-value btw

```

# Diagnostics for tnf

```{r}

tnf_sim_res <- simulateResiduals(fittedModel = tnf_glmer, n = 1000)

plot(tnf_sim_res)

testUniformity(tnf_sim_res)
testDispersion(tnf_sim_res)
testOutliers(tnf_sim_res)

plotResiduals(tnf_sim_res, form = tnf_glmer_input$lab_or_rw)
plotResiduals(tnf_sim_res, form = tnf_glmer_input$t_muris_dose)
plotResiduals(tnf_sim_res, form = tnf_glmer_input$treatment)

ranef(tnf_glmer)

VarCorr(tnf_glmer)

```

# Run emmeans for tnf

```{r}
emm_tnf <- emmeans(tnf_glmer,
                   specs = c("t_muris_dose", "lab_or_rw"),
                   by = c("treatment"),
                   type = "response")

emm_tnf

kable(pairs(emm_tnf))

# Pairwise contrast: Lab vs Rewilded within each treatment
emm_tnf_env <- emmeans(tnf_glmer,
                       specs = "t_muris_dose",
                       by = c("treatment", "lab_or_rw"),
                       type = "response")

kable(pairs(emm_tnf_env))

```

# EMM data frames

```{r}

emm_ifn_df <- as.data.frame(emm_ifn) %>%
  mutate(treatment = factor(treatment, levels = c("IgG", "LPS", "CD28")),
         t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High")),
         lab_or_rw = factor(lab_or_rw, levels = c("Lab", "RW")))

emm_tnf_df <- as.data.frame(emm_tnf) %>%
  mutate(treatment = factor(treatment, levels = c("IgG", "LPS", "CD28")),
         t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High")),
         lab_or_rw = factor(lab_or_rw, levels = c("Lab", "RW")))

```

# EMM IFN-gamma: #1 — Facet = Treatment, x = Environment, Color = Dose
# "Within each treatment, how does infection shift the response across environments?"

```{r}

# We're plotting the LOD/2 values as the data points as well.
# That is what those clustered points are. Those were zeros.

emm_ifn_plot <- ggplot(emm_ifn_df) +
  geom_point(data = ifn_glmer_input %>%
               mutate(treatment = factor(treatment,
                                         levels = c("IgG", "LPS", "CD28")),
                      lab_or_rw = factor(lab_or_rw, levels = c("Lab", "RW")),
                      t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High"))),
             aes(x = lab_or_rw,
                 y = mean_concentration,
                 color = t_muris_dose),
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.5),
             alpha = 0.3,
             size = 1.5) +
  geom_errorbar(aes(x = lab_or_rw,
                    ymin = asymp.LCL,
                    ymax = asymp.UCL,
                    color = t_muris_dose),
                width = 0.3,
                position = position_dodge(width = 0.5),
                alpha = 0.8) +
  geom_point(aes(x = lab_or_rw,
                 y = response,
                 fill = t_muris_dose),
             position = position_dodge(width = 0.5),
             alpha = 1,
             size = 3,
             shape = 21,
             color = "black") +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                     name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_y_continuous(name = expression(bold("IFN-" * gamma ~
                                              "(pg mL"^"-1" * ")")),
                     expand = expansion(mult = 0),
                     trans = "log10",
                     breaks = 10^seq(-1, 4),
                     labels = expression(10^-1, 1, 10^1,
                                         10^2, 10^3, 10^4),
                     limits = c(0.1, 10000),
                     guide = guide_axis_logticks(long = 1.8,
                                                mid = 1.3,
                                                short = 0.8,
                                                expanded = FALSE)) +
  scale_x_discrete(name = expression(bold("Environment")),
                   labels = c("Lab" = "Lab",
                              "RW"  = "Rewilded")) +
  facet_wrap(~ treatment,
             labeller = as_labeller(c("CD28" = "CD28SA",
                                      "LPS" = "LPS",
                                      "IgG" = "IgG1"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.1, "cm")) +
  ggtitle(expression(bold("Marginal Means IFN-" * gamma ~ "(95% C.I.)")))

emm_ifn_plot

plot_saver("emm_ifn_plot.png",
           emm_ifn_plot)

```

# EMM IFN-gamma: #5 — Facet = Dose, x = Treatment, Color = Environment
# "Among infected (or uninfected) mice, which treatment shows the biggest environment effect?"

```{r}

emm_ifn_dose_plot <- ggplot(emm_ifn_df) +
  geom_point(data = ifn_glmer_input %>%
               mutate(treatment = factor(treatment,
                                         levels = c("IgG", "LPS", "CD28")),
                      lab_or_rw = factor(lab_or_rw, levels = c("Lab", "RW")),
                      t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High"))),
             aes(x = treatment,
                 y = mean_concentration,
                 color = lab_or_rw),
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.5),
             alpha = 0.3,
             size = 1.5) +
  geom_errorbar(aes(x = treatment,
                    ymin = asymp.LCL,
                    ymax = asymp.UCL,
                    color = lab_or_rw),
                width = 0.3,
                position = position_dodge(width = 0.5),
                alpha = 0.8) +
  geom_point(aes(x = treatment,
                 y = response,
                 fill = lab_or_rw),
             position = position_dodge(width = 0.5),
             alpha = 1,
             size = 3,
             shape = 21,
             color = "black") +
  scale_fill_manual(values = c("Lab" = color_mustard,
                               "RW"  = color_green),
                    name = expression(bold("Environment")),
                    labels = c("Lab" = "Lab",
                               "RW"  = "Rewilded")) +
  scale_color_manual(values = c("Lab" = color_mustard,
                                "RW"  = color_green),
                     name = expression(bold("Environment")),
                     labels = c("Lab" = "Lab",
                                "RW"  = "Rewilded")) +
  scale_y_continuous(name = expression(bold("IFN-" * gamma ~
                                              "(pg mL"^"-1" * ")")),
                     expand = expansion(mult = 0),
                     trans = "log10",
                     breaks = 10^seq(-1, 4),
                     labels = expression(10^-1, 1, 10^1,
                                         10^2, 10^3, 10^4),
                     limits = c(0.1, 10000),
                     guide = guide_axis_logticks(long = 1.8,
                                                mid = 1.3,
                                                short = 0.8,
                                                expanded = FALSE)) +
  scale_x_discrete(name = expression(bold("Treatment")),
                   labels = c("CD28" = "CD28SA",
                              "LPS" = "LPS",
                              "IgG" = "IgG1")) +
  facet_wrap(~ t_muris_dose,
             labeller = as_labeller(c("Zero" = "Zero Dose",
                                      "High" = "High Dose"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.1, "cm")) +
  ggtitle(expression(bold("Marginal Means of IFN-" * gamma ~ "by Dose (95% CI)")))

emm_ifn_dose_plot

plot_saver("emm_ifn_dose_plot.png",
           emm_ifn_dose_plot)

```

# EMM IFN-gamma: #3 — Facet = Environment, x = Treatment, Color = Dose
# "Within Lab (or RW), which treatment shows the strongest infection effect?"

```{r}

emm_ifn_treatment_plot <- ggplot(emm_ifn_df) +
  geom_point(data = ifn_glmer_input %>%
               mutate(treatment = factor(treatment,
                                         levels = c("IgG", "LPS", "CD28")),
                      lab_or_rw = factor(lab_or_rw, levels = c("Lab", "RW")),
                      t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High"))),
             aes(x = treatment,
                 y = mean_concentration,
                 color = t_muris_dose),
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.5),
             alpha = 0.3,
             size = 1.5) +
  geom_errorbar(aes(x = treatment,
                    ymin = asymp.LCL,
                    ymax = asymp.UCL,
                    color = t_muris_dose),
                width = 0.3,
                position = position_dodge(width = 0.5),
                alpha = 0.8) +
  geom_point(aes(x = treatment,
                 y = response,
                 fill = t_muris_dose),
             position = position_dodge(width = 0.5),
             alpha = 1,
             size = 3,
             shape = 21,
             color = "black") +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                     name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_y_continuous(name = expression(bold("IFN-" * gamma ~
                                              "(pg mL"^"-1" * ")")),
                     expand = expansion(mult = 0),
                     trans = "log10",
                     breaks = 10^seq(-1, 4),
                     labels = expression(10^-1, 1, 10^1,
                                         10^2, 10^3, 10^4),
                     limits = c(0.1, 10000),
                     guide = guide_axis_logticks(long = 1.8,
                                                mid = 1.3,
                                                short = 0.8,
                                                expanded = FALSE)) +
  scale_x_discrete(name = expression(bold("Treatment")),
                   labels = c("CD28" = "CD28SA",
                              "LPS" = "LPS",
                              "IgG" = "IgG1")) +
  facet_wrap(~ lab_or_rw,
             labeller = as_labeller(c("Lab" = "Lab",
                                      "RW"  = "Rewilded"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.1, "cm")) +
  ggtitle(expression(bold("Marginal Means of IFN-" * gamma ~ "by Environment (95% CI)")))

emm_ifn_treatment_plot

plot_saver("emm_ifn_treatment_plot.png",
           emm_ifn_treatment_plot)

```

# EMM TNF-alpha: #1 — Facet = Treatment, x = Environment, Color = Dose

```{r}

emm_tnf_plot <- ggplot(emm_tnf_df) +
  geom_point(data = tnf_glmer_input %>%
               mutate(treatment = factor(treatment,
                                         levels = c("IgG", "LPS", "CD28")),
                      lab_or_rw = factor(lab_or_rw, levels = c("Lab", "RW")),
                      t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High"))),
             aes(x = lab_or_rw,
                 y = mean_concentration,
                 color = t_muris_dose),
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.5),
             alpha = 0.3,
             size = 1.5) +
  geom_errorbar(aes(x = lab_or_rw,
                    ymin = asymp.LCL,
                    ymax = asymp.UCL,
                    color = t_muris_dose),
                width = 0.3,
                position = position_dodge(width = 0.5),
                alpha = 0.8) +
  geom_point(aes(x = lab_or_rw,
                 y = response,
                 fill = t_muris_dose),
             position = position_dodge(width = 0.5),
             alpha = 1,
             size = 3,
             shape = 21,
             color = "black") +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                     name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_y_continuous(name = expression(bold("TNF-" * alpha ~
                                              "(pg mL"^"-1" * ")")),
                     expand = expansion(mult = 0),
                     trans = "log10",
                     breaks = 10^seq(-1, 4),
                     labels = expression(10^-1, 1, 10^1,
                                         10^2, 10^3, 10^4),
                     limits = c(0.1, 10000),
                     guide = guide_axis_logticks(long = 1.8,
                                                mid = 1.3,
                                                short = 0.8,
                                                expanded = FALSE)) +
  scale_x_discrete(name = expression(bold("Environment")),
                   labels = c("Lab" = "Lab",
                              "RW"  = "Rewilded")) +
  facet_wrap(~ treatment,
             labeller = as_labeller(c("CD28" = "CD28SA",
                                      "LPS" = "LPS",
                                      "IgG" = "IgG1"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.1, "cm")) +
  ggtitle(expression(bold("Marginal Means of TNF-" * alpha ~ "(95% CI)")))

emm_tnf_plot

plot_saver("emm_tnf_plot.png",
           emm_tnf_plot)

```

# EMM TNF-alpha: #5 — Facet = Dose, x = Treatment, Color = Environment

```{r}

emm_tnf_dose_plot <- ggplot(emm_tnf_df) +
  geom_point(data = tnf_glmer_input %>%
               mutate(treatment = factor(treatment,
                                         levels = c("IgG", "LPS", "CD28")),
                      lab_or_rw = factor(lab_or_rw, levels = c("Lab", "RW")),
                      t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High"))),
             aes(x = treatment,
                 y = mean_concentration,
                 color = lab_or_rw),
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.5),
             alpha = 0.3,
             size = 1.5) +
  geom_errorbar(aes(x = treatment,
                    ymin = asymp.LCL,
                    ymax = asymp.UCL,
                    color = lab_or_rw),
                width = 0.3,
                position = position_dodge(width = 0.5),
                alpha = 0.8) +
  geom_point(aes(x = treatment,
                 y = response,
                 fill = lab_or_rw),
             position = position_dodge(width = 0.5),
             alpha = 1,
             size = 3,
             shape = 21,
             color = "black") +
  scale_fill_manual(values = c("Lab" = color_mustard,
                               "RW"  = color_green),
                    name = expression(bold("Environment")),
                    labels = c("Lab" = "Lab",
                               "RW"  = "Rewilded")) +
  scale_color_manual(values = c("Lab" = color_mustard,
                                "RW"  = color_green),
                     name = expression(bold("Environment")),
                     labels = c("Lab" = "Lab",
                                "RW"  = "Rewilded")) +
  scale_y_continuous(name = expression(bold("TNF-" * alpha ~
                                              "(pg mL"^"-1" * ")")),
                     expand = expansion(mult = 0),
                     trans = "log10",
                     breaks = 10^seq(-1, 4),
                     labels = expression(10^-1, 1, 10^1,
                                         10^2, 10^3, 10^4),
                     limits = c(0.1, 10000),
                     guide = guide_axis_logticks(long = 1.8,
                                                mid = 1.3,
                                                short = 0.8,
                                                expanded = FALSE)) +
  scale_x_discrete(name = expression(bold("Treatment")),
                   labels = c("CD28" = "CD28SA",
                              "LPS" = "LPS",
                              "IgG" = "IgG1")) +
  facet_wrap(~ t_muris_dose,
             labeller = as_labeller(c("Zero" = "Zero Dose",
                                      "High" = "High Dose"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.1, "cm")) +
  ggtitle(expression(bold("Marginal Means of TNF-" * alpha ~ "by Dose (95% CI)")))

emm_tnf_dose_plot

plot_saver("emm_tnf_dose_plot.png",
           emm_tnf_dose_plot)

```

# EMM TNF-alpha: #3 — Facet = Environment, x = Treatment, Color = Dose

```{r}

emm_tnf_treatment_plot <- ggplot(emm_tnf_df) +
  geom_point(data = tnf_glmer_input %>%
               mutate(treatment = factor(treatment,
                                         levels = c("IgG", "LPS", "CD28")),
                      lab_or_rw = factor(lab_or_rw, levels = c("Lab", "RW")),
                      t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High"))),
             aes(x = treatment,
                 y = mean_concentration,
                 color = t_muris_dose),
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.5),
             alpha = 0.3,
             size = 1.5) +
  geom_errorbar(aes(x = treatment,
                    ymin = asymp.LCL,
                    ymax = asymp.UCL,
                    color = t_muris_dose),
                width = 0.3,
                position = position_dodge(width = 0.5),
                alpha = 0.8) +
  geom_point(aes(x = treatment,
                 y = response,
                 fill = t_muris_dose),
             position = position_dodge(width = 0.5),
             alpha = 1,
             size = 3,
             shape = 21,
             color = "black") +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                     name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_y_continuous(name = expression(bold("TNF-" * alpha ~
                                              "(pg mL"^"-1" * ")")),
                     expand = expansion(mult = 0),
                     trans = "log10",
                     breaks = 10^seq(-1, 4),
                     labels = expression(10^-1, 1, 10^1,
                                         10^2, 10^3, 10^4),
                     limits = c(0.1, 10000),
                     guide = guide_axis_logticks(long = 1.8,
                                                mid = 1.3,
                                                short = 0.8,
                                                expanded = FALSE)) +
  scale_x_discrete(name = expression(bold("Treatment")),
                   labels = c("CD28" = "CD28SA",
                              "LPS" = "LPS",
                              "IgG" = "IgG1")) +
  facet_wrap(~ lab_or_rw,
             labeller = as_labeller(c("Lab" = "Lab",
                                      "RW"  = "Rewilded"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.1, "cm")) +
  ggtitle(expression(bold("Marginal Means of TNF-" * alpha ~ "by Environment (95% CI)")))

emm_tnf_treatment_plot

plot_saver("emm_tnf_treatment_plot.png",
           emm_tnf_treatment_plot)

```

# Comparing IFN and TNF levels to see if there is a correlation

```{r}

ifn_tnf_combined <- left_join(ifn_glmer_input, 
                              tnf_glmer_input,
                              by = c("mouse_id",
                                     "treatment"),
                              suffix = c("_ifn",
                                         "_tnf")) %>%
  select("mouse_id",
         "plate_ifn",
         "lab_or_rw_ifn",
         "strain_ifn",
         "t_muris_dose_ifn",
         "cohort_ifn",
         "experiment_ifn",
         "source_ifn",
         "experiment_day_ifn",
         "treatment",
         "mean_concentration_ifn",
         "mean_concentration_tnf") %>%
  rename("plate" = "plate_ifn",
         "lab_or_rw" = "lab_or_rw_ifn",
         "strain" = "strain_ifn",
         "t_muris_dose" = "t_muris_dose_ifn",
         "cohort" = "cohort_ifn",
         "experiment" = "experiment_ifn",
         "source" = "source_ifn",
         "experiment_day" = "experiment_day_ifn") %>%
  mutate(treatment = factor(treatment, 
                            levels = c("CD28", "LPS", "IgG")))
  

```

# We used log scale here to make sure that the whole graph fits.

```{r}

#TODO need to find out what hte line is telling me.

ifn_tnf_combined_plot <- ggplot(ifn_tnf_combined %>% 
                                  mutate(treatment = factor(treatment, 
                                                            levels = c("IgG", 
                                                                       "LPS", 
                                                                       "CD28"))),
                                aes(x = mean_concentration_ifn,
                                    y = mean_concentration_tnf,
                                    fill = lab_or_rw,
                                    color = lab_or_rw,
                                    group = lab_or_rw)) +
  geom_point(aes(shape = t_muris_dose),
             size = 2,
             alpha = 0.8) +
  geom_smooth(method = "lm",
              se = TRUE,
              alpha = 0.1,
              linewidth = 0.8) +
  facet_grid(~ treatment,
             labeller = as_labeller(c("CD28" = "CD28SA",
                                      "LPS" = "LPS",
                                      "IgG" = "IgG1"))) +
  scale_fill_manual(values = c("RW" = color_green,
                               "Lab" = color_mustard),
                   name = expression(bold("Environment")),
                   labels = c("Lab" = "Lab",
                              "RW" = "Rewilded")) +
  scale_color_manual(values = c("RW" = color_green,
                               "Lab" = color_mustard),
                     name = expression(bold("Environment")),
                     labels = c("Lab" = "Lab",
                              "RW" = "Rewilded")) +
  scale_shape_manual(values = c("Zero" = 1,
                                "High" = 3),
                     name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_x_continuous(name = expression(bold("IFN-"*gamma~"(pg mL"^"-1"*")")),
                     trans = "log10",
                     breaks = 10^seq(-1, 3),
                     labels = expression(10^-1, 
                                         1, 
                                         10^1, 
                                         10^2, 
                                         10^3),
                     limits = c(0.1, 1000),
                     expand = expansion(mult = 0),
                     guide = guide_axis_logticks(long = 1.8,
                                                mid = 1.3,
                                                short = 0.8,
                                                expanded = FALSE)) +
  scale_y_continuous(name = expression(bold("TNF-"*alpha~"(pg mL"^"-1"*")")),
                     trans = "log10",
                     breaks = 10^seq(-1, 3),
                     labels = expression(10^-1, 
                                         1, 
                                         10^1, 
                                         10^2, 
                                         10^3),
                     limits = c(0.1, 1000),
                     expand = expansion(mult = 0),
                     guide = guide_axis_logticks(long = 1.8,
                                                mid = 1.3,
                                                short = 0.8,
                                                expanded = FALSE)) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black", 
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.1, "cm"),
        plot.title = element_text(hjust = 0.5,
                                  face = "bold"),
        panel.spacing = unit(0.6, "cm")) +
  ggtitle(expression(bold("IFN-"*gamma~"vs. TNF-"*alpha~"Concentration")))

ifn_tnf_combined_plot

plot_saver("ifn_tnf_combined_plot.png",
           ifn_tnf_combined_plot)

# Highest IFN
ifn_tnf_combined %>% 
  slice_max(mean_concentration_ifn, n = 10) %>% 
  select(mouse_id, mean_concentration_ifn)

# Highest TNF
ifn_tnf_combined %>% 
  slice_max(mean_concentration_tnf, n = 10) %>% 
  select(mouse_id, mean_concentration_tnf)
```
