---
title: "ELISA analysis"
author: "Zehao Wu"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

# Loading the Libraries

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(janitor)
library(factoextra)
library(skimr)
library(naniar)
library(ggplot2)
library(ggpubr)
library(emmeans)
library(lme4)
library(lmerTest)
library(scales)
library(haven)
library(broom.mixed)
library(readr)
library(here)
library(DHARMa) 
library(rstatix)

```

# Setting the paths

```{r}

base_path <- "/Users/zehao/Documents/Research Projects/Graham Group Thesis Research (Summer 2025)"

data_path <- file.path(base_path, "Metadata", "elisadata")

clean_data_path <- file.path(base_path, "Cleaned data")

figure_path       <- file.path(base_path, "Figures")
elisa_figure_path <- file.path(figure_path, "ELISA Figures")
dir.create(elisa_figure_path, showWarnings = FALSE, recursive = TRUE)

```

# Setting up the functions for repeated codes

# Mean calculator

```{r}

mean_calculator <- function(input_data, 
                            final_concentration, 
                            cv_data,
                            groupings = c("mouse_id",
                                          "treatment", 
                                          "plate", 
                                          "strain", 
                                          "t_muris_dose",
                                          "lab_or_rw",
                                          "cohort",
                                          "experiment",
                                          "source",
                                          "experiment_day")) {
input_data %>%
  group_by(across(all_of(groupings))) %>% 
  summarize(
    mean_concentration = mean({{final_concentration}}, 
                    na.rm = TRUE),
    cv_percent = first(na.omit({{cv_data}})),
    .groups = "drop"
  )}

```

# Setting the colors (Okabe-Ito palette and IBM palette)

```{r}

color_black      <- "#000000"
color_mustard    <- "#E69F00"
color_light_blue <- "#56B4E9"
color_green      <- "#009E73"
color_yellow     <- "#F0E442"
color_dark_blue  <- "#0072B2"
color_orange     <- "#D55E00"
color_purple     <- "#CC79A7"

ibm_yellow <- "#ffb000"
ibm_orange <- "#fe6100"
ibm_reddish_purple <- "#dc267f"
ibm_purple <- "#785ef0"
ibm_blue <- "#648fff"

```

# Basic elisa summary plotter

```{r}

elisa_summary_plotter <- function(input_data, 
                                  y_label, 
                                  title = NULL) {
  ggplot(input_data, aes(x = treatment,
                        y = mean_concentration,
                        fill = t_muris_dose)) +
    geom_boxplot(position = position_dodge(width = 0.75),
               alpha = 0.8,
               outlier.shape = NA,
               width = 0.5,
               color = "black",
               linewidth = 0.6) +
    geom_jitter(position = position_jitterdodge(jitter.width = 0.15,
                                              dodge.width = 0.75),
              alpha = 0.8,
              size = 1.8,
              shape = 21,
              color = "black") +
    facet_wrap(~ lab_or_rw,
               labeller = as_labeller(c("Lab" = "Laboratory",
                                        "RW" = "Rewilded"))) +
    scale_fill_manual(values = c("Zero" = color_dark_blue, 
                               "High" = color_orange),
                    labels = c("Zero" = "Uninfected", 
                               "High" = "Infected"),
                    name = "Infection Status") +
    scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                     labels = c("Zero" = "Uninfected", 
                                "High" = "Infected"),
                    name = "Infection Status") +
    scale_x_discrete(limits = c("CD28", "IgG", "LPS")) +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "white", 
                                        color = "black", 
                                        linewidth = 0.8),
        strip.text = element_text(size = 10, 
                                  face = "bold"),
        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
    xlab("Types of Stimulation") +
    ylab(y_label) +
    ggtitle(title)
}

```

# Treatment specific plotter

```{r}

elisa_treatment_plotter <- function(input_data, 
                                   y_label, 
                                   title = NULL) {
  ggplot(input_data, 
         aes(x = lab_or_rw,
             y = mean_concentration,
             fill = t_muris_dose)) + 
    geom_boxplot(alpha = 0.8,
                 outlier.shape = NA,
                 width = 0.5,
                 color = "black",
                 linewidth = 0.6,
                 position = position_dodge(width = 0.6)) +
    geom_jitter(alpha = 0.8,
                size = 1.8,
                shape = 21,
                color = "black",
                position = position_jitterdodge(jitter.width = 0.1,
                                                dodge.width = 0.6)) +
    scale_fill_manual(values = c("Zero" = color_dark_blue, 
                                 "High" = color_orange),
                    labels = c("Zero" = "Uninfected", 
                               "High" = "Infected"),
                    name = "Infection Status") +
    scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                     labels = c("Zero" = "Uninfected", 
                                "High" = "Infected"),
                     name = "Infection Status") +
    scale_x_discrete(limits = c("Lab", "RW"),
                     labels = c("Lab" = "Laboratory", 
                                "RW" = "Rewilded")) +
    scale_y_continuous(expand = expansion(mult = c(0.05, 0.3))) +
    theme_minimal() +
    theme(strip.background = element_rect(fill = "white", 
                                        color = "black", 
                                        linewidth = 0.8),
        strip.text = element_text(size = 10, 
                                  face = "bold"),
        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
    xlab("Environment") +
    ylab(y_label) +
    ggtitle(title)
}


```

# Emmeans per treatment data frame maker
# NOTE: emm_treatment function was removed — it contained a bug where
# filter(treatment == treatment) compared the column to itself (always TRUE).
# The function was never called anywhere, so it was dead code.

# Plot saver

```{r}

plot_saver <- function(filename,
                            plot,
                            path = elisa_figure_path,
                            width = 6, 
                            height = 4, 
                            dpi = 300) {
  ggsave(filename = filename, 
       plot = plot, 
       path = path,
       width = width, 
       height = height, 
       units = "in", 
       dpi = dpi)
}

```

# Diagnostics for data

```{r}

qq_plotter <- function(model,
                       input,
                       condition) {
  qqnorm(resid(model))
  qqline(resid(model))
  plot(resid(model) ~ input[[condition]])
}

# NOTE: qq_plotter is no longer used for GLMER diagnostics.
# DHARMa (simulateResiduals) is used instead in each diagnostics chunk.

```

## Loading the data

```{r}

elisa_ifn_1 <- read_xlsx(file.path(data_path, 
                                   "IFNg_Plate1_20260121.xlsx"), 
                         sheet = "Plate 1 Data")

elisa_ifn_2 <- read_xlsx(file.path(data_path,
                                   "IFNg_Plate2_20260123.xlsx"), 
                         sheet = "Plate 2 Data")

elisa_tnf_1 <- read_xlsx(file.path(data_path,
                                   "TNFa_Plate1_20260121.xlsx"), 
                         sheet = "Plate 1 Data")

elisa_tnf_2 <- read_xlsx(file.path(data_path,
                                   "TNFa_Plate2_20260123.xlsx"), 
                         sheet = "Plate 2 Data")

plate1_map <- read_xlsx(file.path(data_path,
                                  "SF ELISA Plate Mapping.xlsx"), 
                        sheet = "Plate 1 for both IFN and TNF")

plate2_map <- read_xlsx(file.path(data_path,
                                  "SF ELISA Plate Mapping.xlsx"), 
                        sheet = "Plate 2 for both IFN and TNF")

meta_data <- read_csv(file.path(clean_data_path,
                                "clean_meta_data.csv"))

```

# data cleaning and organization

# Four original cell culture plates, each stimulated on a different day:
#   Original plate 1 stimulated July 22, 2025 -> experiment_day = 1
#   Original plate 2 stimulated July 23, 2025 -> experiment_day = 2
#   Original plate 3 stimulated July 29, 2025 -> experiment_day = 3
#   Original plate 4 stimulated July 30, 2025 -> experiment_day = 4
# Each mouse appears on exactly one stimulation plate (one experiment_day).
# For ELISA detection, original plates were combined into 2 detection plates:
#   Detection plate 1 = original plates 1 + 2 (July 22 + 23)
#   Detection plate 2 = original plates 3 + 4 (July 29 + 30)
# The detection plate number is coded as `plate` in the cleaning chunks below.
# experiment_day (1-4) must be added as a column in the plate map Excel files
# so that it joins in automatically via left_join(plate1_map/plate2_map).

# ifn data cleaning for plate 1

```{r}

clean_ifn_1 <- elisa_ifn_1 %>%
  clean_names() %>% 
  rename("final_ifn" = "final_if_ng") %>% # cleaner names
  left_join(plate1_map,  by = "well_id") %>% # join elisa data and mouse number using plate number
  mutate(plate = 1) %>% # create a new column to signal that this is plate 1
  left_join(meta_data, by = "mouse_id") %>% # add the rest of the meta data
  mutate(
    across(c(concentration, 
             final_ifn, mean, 
             std_dev, 
             cv_percent),
            # apply the same transformation across these columns
           ~ case_when(
              # ~ is like a mini function. case when is checking for a condition and applying the first operation if the first condition matches.
             str_detect(., "^<") ~ 0,
             # . is the current value in the column and the rest of the function is asking if this value starts with a <. If yes, then replace it with 0. This is just turning all the <0.0000 values into actual numbers of 0.
             TRUE ~ as.numeric(str_remove(., "^>"))
           )) # the TRUE means that if none of the previous conditinos matched, then do these. Same logic applies where we remove > from values if present. The as numeric converts the text to number.
  ) %>%
  drop_na(mouse_id)

```

# ifn data cleaning for plate 2

```{r}

clean_ifn_2 <- elisa_ifn_2 %>% #same as before...
  clean_names() %>%
  rename("final_ifn" = "final_if_ng") %>%
  left_join(plate2_map,  by = "well_id") %>%
  mutate(plate = 2) %>%
  left_join(meta_data, by = "mouse_id") %>%
   mutate(
    across(c(concentration, final_ifn, mean, std_dev, cv_percent),
           ~ case_when(
             str_detect(., "^<") ~ 0,
             TRUE ~ as.numeric(str_remove(., "^>"))
           ))
  ) %>%
  drop_na(mouse_id)


total_ifn <- bind_rows(clean_ifn_1, 
                       clean_ifn_2)

write_csv(total_ifn, file.path(clean_data_path, "combined_ifn_elisa_data.csv"))

```

For now, I just kept all of the values that are over the detection limit 
without their ">" and all of the small values as zero. Discuss with David.

# tnf data cleaning for plate 1

```{r}

clean_tnf_1 <- elisa_tnf_1 %>% #same as before...
  clean_names() %>%
  rename("final_tnf" = "concentration") %>%
  left_join(plate1_map,  by = "well_id") %>%
  mutate(plate = 1) %>%
  left_join(meta_data, by = "mouse_id") %>%
   mutate(
    across(c(final_tnf, mean, std_dev, cv_percent),
           ~ case_when(
             str_detect(., "^<") ~ 0,
             TRUE ~ as.numeric(str_remove(., "^>"))
           ))
  ) %>%
  drop_na(mouse_id)

```

# tnf data cleaning for plate 2

```{r}

clean_tnf_2 <- elisa_tnf_2 %>% #same as before...
  clean_names() %>%
  rename("final_tnf" = "concentration") %>%
  left_join(plate2_map,  by = "well_id") %>%
  mutate(plate = 2) %>%
  left_join(meta_data, by = "mouse_id") %>%
   mutate(
    across(c(final_tnf, mean, std_dev, cv_percent),
           ~ case_when(
             str_detect(., "^<") ~ 0,
             TRUE ~ as.numeric(str_remove(., "^>"))
           ))
  ) %>%
  drop_na(mouse_id)

total_tnf <- bind_rows(clean_tnf_1, 
                       clean_tnf_2)

write_csv(total_tnf, file.path(clean_data_path, 
                               "combined_tnf_elisa_data.csv"))

```

# Calculating the mean concentration value

```{r}

ifn_input <- mean_calculator(total_ifn, 
                             final_ifn, 
                             cv_percent)

tnf_input <- mean_calculator(total_tnf, 
                             final_tnf, 
                             cv_percent)

```

I wonder if I should use the original data points instead of the mean data.

# Filter the data for treatment specific plots

```{r}

ifn_input_cd28 <- ifn_input %>% filter(treatment == "CD28")

ifn_input_lps <- ifn_input %>% filter(treatment == "LPS")

ifn_input_igg <- ifn_input %>% filter(treatment == "IgG")

tnf_input_cd28 <- tnf_input %>% filter(treatment == "CD28")

tnf_input_lps <- tnf_input %>% filter(treatment == "LPS")

tnf_input_igg <- tnf_input %>% filter(treatment == "IgG")

```

# General ifn plot to get an overview

```{r}

ifn_input <- ifn_input %>%
  mutate(t_muris_dose = factor(t_muris_dose, 
                               levels = c("Zero", "High")))

ifn_plot <- elisa_summary_plotter(ifn_input, 
                                  expression(bold("IFN-" * gamma ~ "(pg mL"^"-1"*")")),
                                  title = expression(bold("IFN-"*gamma~"Concentration by Stimulation"))) 

ifn_plot

plot_saver("general_ifn_plot.png",
           ifn_plot)

```

# General tnf plots

```{r}

tnf_input <- tnf_input %>%
  mutate(t_muris_dose = factor(t_muris_dose, 
                               levels = c("Zero", "High")))

tnf_plot <- elisa_summary_plotter(tnf_input, 
                                  expression(bold("TNF-"*alpha ~ "(pg mL"^"-1"*")")),
                                  title = expression(bold("Summary TNF-"*alpha~"Plot for All Conditions")))

tnf_plot

plot_saver("general_tnf_plot.png",
           tnf_plot)

```

# Treatment specific plots for ifn: CD28

```{r}

ifn_input_cd28 <- ifn_input_cd28 %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High")))

ifn_cd28_plot <- elisa_treatment_plotter(ifn_input_cd28,
                                       expression(bold("IFN-" * gamma ~ "(pg mL"^"-1"*")")),
                                       title = "CD28 Stimulation")

ifn_cd28_plot

plot_saver("ifn_cd28_treatment_plot.png",
           ifn_cd28_plot)

```

# Treatment specific plots for tnf: CD28

```{r}

tnf_input_cd28 <- tnf_input_cd28 %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High")))

tnf_cd28_plot <- elisa_treatment_plotter(tnf_input_cd28,
                                       expression(bold("TNF-"*alpha ~ "(pg mL"^"-1"*")")),
                                       title = "CD28 Stimulation")

tnf_cd28_plot

plot_saver("tnf_cd28_plot.png",
           tnf_cd28_plot)

```

# Treatment specific plots for ifn: LPS

```{r}

ifn_input_lps <- ifn_input_lps %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High")))

ifn_lps_plot <- elisa_treatment_plotter(ifn_input_lps,
                                       expression(bold("IFN-" * gamma ~ "(pg mL"^"-1"*")")),
                                       title = "LPS Stimulation")

ifn_lps_plot

plot_saver("ifn_lps_treatment_plot.png",
           ifn_lps_plot)

```

# Treatment specific plots for tnf: LP

```{r}

tnf_input_lps <- tnf_input_lps %>%
  mutate(t_muris_dose = factor(t_muris_dose, 
                               levels = c("Zero", "High")))

tnf_lps_plot <- elisa_treatment_plotter(tnf_input_lps,
                                       expression(bold("TNF-"*alpha ~ "(pg mL"^"-1"*")")),
                                       title = "LPS Stimulation")

tnf_lps_plot

plot_saver("tnf_lps_plot.png",
           tnf_lps_plot)

```

# Treatment specifiic plots for ifn: IgG

```{r}

ifn_input_igg <- ifn_input_igg %>%
  mutate(t_muris_dose = factor(t_muris_dose, levels = c("Zero", "High")))

ifn_igg_plot <- elisa_treatment_plotter(ifn_input_igg,
                                       expression(bold("IFN-" * gamma ~ "(pg mL"^"-1"*")")),
                                       title = "IgG Stimulation")

ifn_igg_plot

plot_saver("ifn_igg_treatment_plot.png",
           ifn_igg_plot)

```

# Treatment specifiic plots for tnf: IgG

```{r}

tnf_input_igg <- tnf_input_igg %>%
  mutate(t_muris_dose = factor(t_muris_dose, 
                               levels = c("Zero", "High")))

tnf_igg_plot <- elisa_treatment_plotter(tnf_input_igg,
                                       expression(bold("TNF-"*alpha ~ "(pg mL"^"-1"*")")),
                                       title = "IgG Stimulation")

tnf_igg_plot

plot_saver("tnf_igg_plot.png",
           tnf_igg_plot)

```

# Setting reference levels for ifn

# Mixed effect model (gamma) for ifn and diagnostics
# Note to self: should I just assume that all zeros are at the detection level?
# Maybe we can just take half of the detection limit (standard practice)
# Detection level for IFN is 1.797 pg/mL so we use this to handle zero values.

```{r}

ifn_input <- ifn_input %>%
  mutate(
    t_muris_dose = relevel(factor(t_muris_dose), ref = "Zero"),
    treatment = relevel(factor(treatment), ref = "IgG"),
    lab_or_rw = relevel(factor(lab_or_rw), ref = "Lab")
  ) 

ifn_glmer_input <- ifn_input %>%
  mutate(mean_concentration = ifelse(mean_concentration == 0,
                                     1.797 / 2, 
                                     mean_concentration))


ifn_glmer <- glmer(mean_concentration ~ lab_or_rw *
                     t_muris_dose *
                     treatment +
                     (1 | mouse_id) +
                     (1 | experiment_day),
                  family = Gamma(link = "log"),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list(maxfun = 2e5)),
                  data = ifn_glmer_input)

summary(ifn_glmer)

```


# Diagnostics for ifn

```{r}

# DHARMa simulation-based diagnostics for Gamma GLMER
ifn_sim_res <- simulateResiduals(fittedModel = ifn_glmer, n = 1000)

# Standard DHARMa panel: uniform QQ plot + residuals vs. fitted
plot(ifn_sim_res)

# Formal tests
testUniformity(ifn_sim_res)   # tests whether residuals are uniformly distributed
testDispersion(ifn_sim_res)   # tests for over/under-dispersion
testOutliers(ifn_sim_res)     # tests for outliers

# Residuals by each predictor to check group-level misfit
plotResiduals(ifn_sim_res, form = ifn_glmer_input$lab_or_rw)
plotResiduals(ifn_sim_res, form = ifn_glmer_input$t_muris_dose)
plotResiduals(ifn_sim_res, form = ifn_glmer_input$treatment)

ranef(ifn_glmer)

VarCorr(ifn_glmer)

```

# Run emmeans for ifn

```{r}

emm_ifn <- emmeans(ifn_glmer, 
                   specs = c("t_muris_dose","lab_or_rw"), 
                   by = c("treatment"),
                   type = "response")
emm_ifn

pairs(emm_ifn)

```

# Doing the same thing here
# The lowest detection level is 3.7 for TNF so halving that.

```{r}

tnf_input <- tnf_input %>%
  mutate(
    t_muris_dose = relevel(factor(t_muris_dose), 
                           ref = "Zero"),
    treatment = relevel(factor(treatment), 
                        ref = "IgG"),
    lab_or_rw = relevel(factor(lab_or_rw), 
                        ref = "Lab")
  )

tnf_glmer_input <- tnf_input %>%
  mutate(mean_concentration = ifelse(mean_concentration == 0,
                                     3.7 / 2, 
                                     mean_concentration))

# we want to set the zero dose, IgG and lab groups to be the basis of all our comparisons.

tnf_glmer <- glmer(mean_concentration ~ lab_or_rw *
                     t_muris_dose *
                     treatment +
                     (1 | mouse_id) +
                     (1 | experiment_day),
                  family = Gamma(link = "log"),
                  control = glmerControl(optimizer = "bobyqa",
                                         optCtrl = list(maxfun = 2e5)),
                 data = tnf_glmer_input)

summary(tnf_glmer)

```

# Diagnostics for tnf

```{r}

# DHARMa simulation-based diagnostics for Gamma GLMER
tnf_sim_res <- simulateResiduals(fittedModel = tnf_glmer, n = 1000)

# Standard DHARMa panel: uniform QQ plot + residuals vs. fitted
plot(tnf_sim_res)

# Formal tests
testUniformity(tnf_sim_res)   # tests whether residuals are uniformly distributed
testDispersion(tnf_sim_res)   # tests for over/under-dispersion
testOutliers(tnf_sim_res)     # tests for outliers

# Residuals by each predictor to check group-level misfit
plotResiduals(tnf_sim_res, form = tnf_glmer_input$lab_or_rw)
plotResiduals(tnf_sim_res, form = tnf_glmer_input$t_muris_dose)
plotResiduals(tnf_sim_res, form = tnf_glmer_input$treatment)

ranef(tnf_glmer)

VarCorr(tnf_glmer)

```

# Run emmeans for tnf

```{r}
emm_tnf <- emmeans(tnf_glmer, 
                   specs = c("t_muris_dose"), 
                   by = c("lab_or_rw", "treatment"),
                   type = "response")
emm_tnf

pairs(emm_tnf)

```

# Plot interaction plots for ifn by emmeans (TODO)

```{r}

emm_ifn_df <- as.data.frame(emm_ifn) %>%
  mutate(treatment = factor(treatment, 
                            levels = c("CD28", "LPS", "IgG")))

emm_tnf_df <- as.data.frame(emm_tnf) %>%
  mutate(treatment = factor(treatment, 
                            levels = c("CD28", "LPS", "IgG")))

```

```{r}

emm_ifn_plot <- ggplot(emm_ifn_df) +
  geom_point(aes(x = treatment, 
                 y = response, 
                 fill = t_muris_dose),
             position = position_dodge(width = 0.5),
             alpha = 1,
             size = 2,
             shape = 21,
             color = "black") +
  geom_errorbar(aes(x = treatment,
                    ymin = asymp.LCL,
                    ymax = asymp.UCL,
                    color = t_muris_dose),
                width = 0.3,
             position = position_dodge(width = 0.5),
             alpha = 0.8) +
  geom_point(data = ifn_input,
             aes(x = treatment,
                 y = mean_concentration,
                 color = t_muris_dose),
                 position = position_jitterdodge(jitter.width = 0.1,
                                                 dodge.width = 0.5),
                 alpha = 0.3,
                 size = 1.5) +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "High" = color_orange),
                    name = "Infection Status",
                    labels = c("Zero" = "Uninfected",
                              "High" = "Infected")) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                     name = "Infection Status",
                     labels = c("Zero" = "Uninfected",
                              "High" = "Infected")) +
  scale_y_continuous(name = expression(bold("IFN-" * gamma ~ "(pg mL"^"-1"*")")),
                     expand = expansion(mult = c(0.05, 0.3))) +
  scale_x_discrete(name = "Types of Stimulation") +
  facet_wrap(~ lab_or_rw,
             labeller = as_labeller(c("Lab" = "Laboratory",
                                      "RW" = "Rewilded"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white", 
                                        color = "black", 
                                        linewidth = 0.8),
        strip.text = element_text(size = 10, 
                                  face = "bold"),
        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold")) +
  ggtitle(expression(bold("Estimated Marginal Means of IFN-" * gamma)))

emm_ifn_plot

plot_saver("emm_ifn_plot.png",
           emm_ifn_plot)

```


# run emmeans for tnf

```{r}

emm_tnf_plot <- ggplot(emm_tnf_df) + 
  geom_point(aes(x = treatment, 
                 y = response, 
                 fill = t_muris_dose),
             position = position_dodge(width = 0.5),
             alpha = 1,
             shape = 21,
             size = 2,
             color = "black") +
  geom_errorbar(aes(x = treatment, 
                    ymin = asymp.LCL,
                    ymax = asymp.UCL, 
                color = t_muris_dose), 
                width = 0.3,
                alpha = 0.8,
                position = position_dodge(width = 0.5)) +
  geom_point(data = tnf_input,
             aes(x = treatment,
                 y = mean_concentration,
                 color = t_muris_dose),
                 position = position_jitterdodge(jitter.width = 0.1,
                                                 dodge.width = 0.5),
                 alpha = 0.3,
                 size = 1.5) +
  scale_y_continuous(name = expression(bold("TNF-"*alpha~"(pg mL"^"-1"*")")),
                     expand = expansion(mult = c(0.05, 0.3))) +
  scale_x_discrete(name = "Types of Stimulation") +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "High" = color_orange),
                    name = "Infection Status",
                    labels = c("Zero" = "Uninfected",
                              "High" = "Infected")) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                     name = "Infection Status",
                     labels = c("Zero" = "Uninfected",
                              "High" = "Infected")) +
  facet_wrap(~ lab_or_rw,
             labeller = as_labeller(c("Lab" = "Laboratory",
                                      "RW" = "Rewilded"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white", 
                                        color = "black", 
                                        linewidth = 0.8),
        strip.text = element_text(size = 10, 
                                  face = "bold"),
        panel.border = element_rect(color = "black", 
                                    fill = NA, 
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold")) +
  ggtitle(expression(bold("Estimated Marginal Means of TNF-" * alpha)))

emm_tnf_plot

plot_saver("emm_tnf_plot.png",
           emm_tnf_plot)

```

