---
title: "Spleen_mass_analysis"
author: "Zehao Wu"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

# First time run order: Spleen → CBC → Spleen
This markdown produces clean_meta_data.csv used by CBC and ELISA analysis.
The section looking at the correlation between spleen size and lymphocyte counts
at the bottom requires cleaned_combined_cbc_data.csv from CBC. That's why CBC
needs to be ran on the second try. But, this is only applicable if this is the
first time this code is ran.

The analysis done in the ELISA code relies on the worm burden data cleaned here.

# Load the needed library first

```{r}
library(tidyverse)
library(readxl)
library(writexl)
library(janitor)
library(factoextra)
library(skimr)
library(naniar)
library(ggplot2)
library(ggpubr)
library(emmeans)
library(lme4)
library(lmerTest)
library(scales)
library(DHARMa)
library(kableExtra)
library(sjPlot) # does not work when knitting to PDF

```

# Setting the paths

```{r}

base_path <- "/Users/zehao/Documents/Research Projects/Graham Group Thesis Research (Summer 2025)"

data_path <- file.path(base_path, 
                       "Metadata")
clean_data_path <- file.path(base_path, 
                             "Cleaned data")
figure_path <- file.path(base_path, 
                         "Figures")
spleen_figure_path <- file.path(figure_path, 
                                "Spleen Figures")
dir.create(spleen_figure_path, 
           showWarnings = FALSE, 
           recursive = TRUE)

```

# Plot saver

```{r}

plot_saver <- function(filename,
                       plot,
                       path = spleen_figure_path,
                       width = 6,
                       height = 4,
                       dpi = 300) {
  ggsave(filename = filename,
         plot = plot,
         path = path,
         width = width,
         height = height,
         units = "in",
         dpi = dpi)
}

```

# Loading the data

```{r}

spleen_data <- read_excel(file.path(data_path,
                                    "SF Project Spleen Data (2025).xlsx"),
                          sheet = "spleen cell count data")
body_mass_data <- read_excel(file.path(data_path, 
                                       "metamouse25_12_16_25_ver.xlsx"),
                             sheet = "mass and fecal data")
spleen_lym_count <- read_excel(file.path(data_path, 
                                         "SF Project Spleen Data (2025).xlsx"),
                               sheet = "spleen cell concentration calcu")
meta_data <- read_excel(file.path(data_path, 
                                  "metamouse25_12_16_25_ver.xlsx"),
                        sheet = "metadata")

cbc_data <- read_csv(file.path(clean_data_path,
                               "cleaned_combined_cbc_data.csv"))

worm_burden <- read_excel(file.path(data_path,
                                    "Worm_burden.xlsx"),
                          sheet = "Worm burden")

```

# Setting the colors (Okabe-Ito palette and Paul Tol's Bright)

```{r}

color_black <- "#000000"
color_mustard <- "#E69F00"
color_light_blue <- "#56B4E9"
color_green <- "#009E73"
color_yellow <- "#F0E442"
color_dark_blue <- "#0072B2"
color_orange <- "#D55E00"
color_purple <- "#CC79A7"

color_grey <- "#BBBBBB"
color_dark_purple <- "#AA3377"
color_light_purple <- "#EE6677"
color_sand <- "#CCBB44"
color_cyan <- "#66CCEE"
color_olive <- "#228833"
color_navy <- "#4477AA"

```

# Cleaning the spleen data

```{r}

clean_spleen_data <- spleen_data %>% 
  filter(experiment == "NSF", 
         !is.na(`spleen mass (g)`)) %>%
  select(mouse.id, 
         strain, 
         T.muris.dose, 
         lab.or.rw, 
         `spleen mass (g)`) %>%
  clean_names() %>% 
  mutate(mouse_id = str_trim(as.character(mouse_id)),
         strain = as.character(strain),
         t_muris_dose = case_match(t_muris_dose, "0" ~ "Zero", .default = t_muris_dose),
         t_muris_dose = factor(t_muris_dose,
                                  levels = c("Zero", "Low", "High")),
         lab_or_rw = factor(lab_or_rw),
         spleen_mass_g = as.numeric(spleen_mass_g))

str(clean_spleen_data)

# Only using the NSF ones since I only have spleens from those

```

# Cleaning the mass data

```{r}
clean_body_mass_data <- body_mass_data %>% 
  filter(week == 3 | week == 4, 
         !is.na(mass)) %>% 
  select(mouse.id, 
         mass, 
         `cage/wedge`) %>%
  clean_names() %>%
  mutate(mouse_id = str_trim(as.character(mouse_id)),
         mass = as.numeric(mass),
         cage_wedge = if_else(cage_wedge == "spare", 
                              "25", 
                              cage_wedge),
         cage_wedge = factor(cage_wedge))

str(clean_body_mass_data)

# Let's assign the spares to cage 25 for ease of calculations.
# I'm going to only use week 3 and 4 since that's the end of the experiment
```

# Cleaning splenocyte data

```{r}

clean_spleen_lym_count <- spleen_lym_count %>%
  clean_names() %>%
  select(avg_total_cell_concentration_cell_m_l, 
         solution_volume_m_l, 
         mouse_id) %>%
  mutate(mouse_id = as.character(mouse_id) %>% str_trim(), 
         total_lym_number = avg_total_cell_concentration_cell_m_l * solution_volume_m_l,
         total_lym_number = as.numeric(total_lym_number)) %>%
  select(mouse_id,
         total_lym_number)

```

# Cleaning meta data

```{r}

clean_meta_data <- meta_data %>% 
  clean_names() %>%
  filter(experiment == "NSF", 
         !is.na(t_muris_dose)) %>%
  select(mouse_id, 
         strain, 
         experiment, 
         t_muris_dose, 
         lab_or_rw, 
         cohort, 
         source) %>%
   mutate(mouse_id = str_trim(as.character(mouse_id)),
         strain = as.character(strain),
         t_muris_dose = case_match(t_muris_dose, "0" ~ "Zero", 
                                   .default = t_muris_dose),
         t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero", 
                                          "Low", 
                                          "High")),
         lab_or_rw = factor(lab_or_rw),
         experiment = as.character(experiment))

str(clean_meta_data)

write_csv(clean_meta_data,
          file.path(clean_data_path, 
                    "clean_meta_data.csv"))

```
Cleaning worm burden data

```{r}

clean_worm_burden <- worm_burden %>%
  clean_names() %>%
  select(mouse_id,
         total_worms)

write_csv(clean_worm_burden,
          file.path(clean_data_path,
                    "clean_worm_burden.csv"))

```


We want to include in a file called "compiled data." But, before that I want to check to make sure the strains, ID, and the lab/rw conditions all match across the two data sets.

1.  Mouse ID
2.  Strain
3.  NSF Experiment
4.  T. muris Dosage
5.  Lab or RW
6.  Week 3 Body Mass
7.  Week 3 Spleen Mass

# Checking for mistakes

```{r}

compiled_data_strain_check <- clean_meta_data %>% 
  full_join(clean_spleen_data, 
            by = "mouse_id") %>%
  mutate(match = strain.x == strain.y) %>%
  filter(!match)

compiled_data_strain_check

compiled_data_lab_rw_check <- clean_meta_data %>%
  full_join(clean_spleen_data, 
            by = "mouse_id") %>%
  mutate(match = lab_or_rw.x == lab_or_rw.y) %>%
  filter(!match)

compiled_data_lab_rw_check

compiled_data_t_muris_dose_check <- clean_meta_data %>%
  full_join(clean_spleen_data, 
            by = "mouse_id") %>%
  mutate(match = t_muris_dose.x == t_muris_dose.y) %>%
  filter(!match)

compiled_data_t_muris_dose_check

#Mouse 1350 has different worm dose depending on the sheet, inner join will not include that data

```

# Creating compiled data

```{r}

spleen_mass_compiled_data <- clean_meta_data %>% 
  left_join(clean_spleen_data, 
            by = "mouse_id") %>%
  left_join(clean_body_mass_data, 
            by = "mouse_id") %>%
  filter(!is.na(spleen_mass_g), 
         !is.na(mass)) %>%
  select(mouse_id,
         strain = strain.x,
         lab_or_rw = lab_or_rw.x,
         t_muris_dose = t_muris_dose.x,
         spleen_mass_g,
         mass,
         cage_wedge,
         source,
         cohort) %>%
  mutate(strain = factor(strain),
         strain = relevel(strain, 
                          ref = "C57BL/6J"),
         t_muris_dose = factor(t_muris_dose,
                               levels = c("Zero",
                                          "Low",
                                          "High")),
         lab_or_rw = factor(lab_or_rw,
                            levels = c("Lab",
                                       "RW")))

```

Now I will calculate the spleen-mass ratio (SMR) and make a plot. Still thinking whether this is the correct way to do this. See the last plot.

# Calculating spleen-mass ratio (SMR)

```{r}
spleen_mass_compiled_data <- spleen_mass_compiled_data %>% 
  mutate(spleen_mass_ratio = spleen_mass_g/mass)

lym_number_spleen_data <- spleen_mass_compiled_data %>% 
  left_join(clean_spleen_lym_count, 
            by = "mouse_id") %>%
  filter(!is.na(total_lym_number))

write_csv(lym_number_spleen_data,
          file.path(clean_data_path, 
                    "lym_number_spleen_data.csv"))

write_csv(spleen_mass_compiled_data,
          file.path(clean_data_path, 
                    "clean_compiled_spleen_data.csv"))

```

# Filte the cbc data to join with spleen data

```{r}

cbc_lym <- cbc_data %>%
  filter(phase == "Post-RW") %>%
  select(mouse_id, 
         lym_number) %>%
  mutate(mouse_id = str_trim(as.character(mouse_id)))

spleen_cbc <- spleen_mass_compiled_data %>% 
  left_join(cbc_lym,
            by = "mouse_id") %>%
  filter(!is.na(lym_number),
         !is.na(spleen_mass_g))

```

# Plotting SMR data

```{r}

spleen_mass_ratio_plot <- ggplot(spleen_mass_compiled_data, 
                                 aes(x = t_muris_dose, 
                          y = spleen_mass_ratio, 
                          fill = lab_or_rw,
                          color = lab_or_rw,
                          group = interaction(t_muris_dose, 
                                              lab_or_rw))) + 
  geom_boxplot(position = position_dodge(width = 0.75),
               alpha = 0.8,
               outlier.shape = NA,
               width = 0.5,
               color = "black",
               linewidth = 0.6) +
 geom_jitter(position = position_jitterdodge(jitter.width = 0.15,
                                             dodge.width  = 0.75),
             alpha = 0.8,
             size = 1.8,
             shape = 21,
             color = "black") + 
  facet_wrap(~ strain) + 
  scale_fill_manual(values = c("RW" = color_green,
                               "Lab" = color_purple),
                    name = expression(bold("Environment")),
                    labels = c("Lab" = "Lab",
                               "RW" = "Rewilded")) +
  scale_color_manual(values = c("RW" = color_green,
                                "Lab" = color_purple),
                     name = expression(bold("Environment")),
                     labels = c("Lab" = "Lab",
                                "RW" = "Rewilded")) +
  scale_y_continuous(name = expression(bold("Spleen to Body Mass Ratio (SMR)")),
                     expand = expansion(mult = c(0.05, 0.3))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  xlab(expression(bolditalic("T. muris")~bold("Dose"))) +
  ggtitle("Spleen to Body Mass Ratio (SMR)")

spleen_mass_ratio_plot

plot_saver("smr_general.png",
           spleen_mass_ratio_plot)

```

# lmer modeling for SMR

```{r}
smr_lmer <- lmer(spleen_mass_ratio ~ 
                   strain * 
                   lab_or_rw * 
                   t_muris_dose + 
                   (1 | cage_wedge), 
                 data = spleen_mass_compiled_data)

tab_model(smr_lmer, show.stat = TRUE)
# statistic is t-value btw
summary(smr_lmer)

emm_smr <- emmeans(smr_lmer,
                   specs = c("t_muris_dose", "lab_or_rw"),
                   by = c("strain"),
                   type = "response")
emm_smr

pairs(emm_smr)

kable(pairs(emm_smr),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_smr_df <- as.data.frame(emm_smr)

emm_smr_plot <- ggplot(emm_smr_df) + 
  geom_point(aes(x = t_muris_dose, 
                 y = emmean, 
                 fill = lab_or_rw,
                 group = lab_or_rw),
             position = position_dodge(width = 0.5),
             alpha = 1,
             size = 3,
             shape = 21,
             color = "black") +
  geom_errorbar(aes(x = t_muris_dose, 
                    ymin = lower.CL,
                    ymax = upper.CL, 
                    color = lab_or_rw,
                    group = lab_or_rw), 
                width = 0.3,
                alpha = 0.8,
             position = position_dodge(width = 0.5)) +
  geom_point(data = spleen_mass_compiled_data,
             aes(x = t_muris_dose,
                 y = spleen_mass_ratio,
                 color = lab_or_rw),
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.5),
             alpha = 0.3,
             size = 1.5) +
  scale_y_continuous(name = expression(bold("Spleen to Body Mass Ratio (SMR)")),
                     expand = expansion(mult = c(0.05, 0.3))) +
  scale_x_discrete(name = expression(bolditalic("T. muris")~bold("Dose"))) +
  scale_fill_manual(values = c("Lab" = color_purple,
                               "RW" = color_green),
                    name = expression(bold("Environment")),
                    labels = c("Lab" = "Lab",
                              "RW" = "Rewilded")) +
  scale_color_manual(values = c("Lab" = color_purple,
                                "RW" = color_green),
                     name = expression(bold("Environment")),
                     labels = c("Lab" = "Lab",
                              "RW" = "Rewilded")) +
  facet_wrap(~ strain) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("Marginal Means of SMR (95% CI)")

emm_smr_plot

plot_saver("emm_smr_plot.png",
           emm_smr_plot)

```

```{r}

smr_sim_res <- simulateResiduals(fittedModel = smr_lmer, n = 1000)

plot(smr_sim_res)

testUniformity(smr_sim_res)
testDispersion(smr_sim_res)
testOutliers(smr_sim_res)

plotResiduals(smr_sim_res, form = spleen_mass_compiled_data$lab_or_rw)
plotResiduals(smr_sim_res, form = spleen_mass_compiled_data$t_muris_dose)
plotResiduals(smr_sim_res, form = spleen_mass_compiled_data$strain)

ranef(smr_lmer)

VarCorr(smr_lmer)

```


# lmer and emm of spleen mass only

```{r}

spleen_lmer <- lmer(spleen_mass_g ~ 
                   strain * 
                   lab_or_rw * 
                   t_muris_dose + 
                   (1 | cage_wedge), 
                 data = spleen_mass_compiled_data)

tab_model(spleen_lmer, show.stat = TRUE)
# statistic is t-value btw
summary(spleen_lmer)

emm_spleen_mass <- emmeans(spleen_lmer,
                   specs = c("t_muris_dose", "lab_or_rw"),
                   by = c("strain"),
                   type = "response")
emm_spleen_mass

pairs(emm_spleen_mass)

kable(pairs(emm_spleen_mass),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_spleen_mass_df <- as.data.frame(emm_spleen_mass)

emm_spleen_mass_plot <- ggplot(emm_spleen_mass_df) + 
  geom_point(aes(x = t_muris_dose, 
                 y = emmean, 
                 fill = lab_or_rw,
                 group = lab_or_rw),
             position = position_dodge(width = 0.5),
             alpha = 1,
             size = 3,
             shape = 21,
             color = "black") +
  geom_errorbar(aes(x = t_muris_dose, 
                    ymin = lower.CL,
                    ymax = upper.CL, 
                    color = lab_or_rw,
                    group = lab_or_rw), 
                width = 0.3,
                alpha = 0.8,
             position = position_dodge(width = 0.5)) +
  geom_point(data = spleen_mass_compiled_data,
             aes(x = t_muris_dose,
                 y = spleen_mass_g,
                 color = lab_or_rw),
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.5),
             alpha = 0.3,
             size = 1.5) +
  scale_y_continuous(name = expression(bold("Spleen Mass (g)")),
                     expand = expansion(mult = c(0.05, 0.3))) +
  scale_x_discrete(name = expression(bolditalic("T. muris")~bold("Dose"))) +
  scale_fill_manual(values = c("Lab" = color_purple,
                               "RW" = color_green),
                    name = expression(bold("Environment")),
                    labels = c("Lab" = "Lab",
                              "RW" = "Rewilded")) +
  scale_color_manual(values = c("Lab" = color_purple,
                                "RW" = color_green),
                     name = expression(bold("Environment")),
                     labels = c("Lab" = "Lab",
                              "RW" = "Rewilded")) +
  facet_wrap(~ strain) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("Marginal Means of Spleen Mass (95% CI)")

emm_spleen_mass_plot

plot_saver("emm_spleen_mass_plot.png",
           emm_spleen_mass_plot)

```

# We use log scale here to make sure that the whole thing fits.

```{r}

spleen_mass_reg <- ggplot(spleen_mass_compiled_data, 
                                 aes(x = mass, 
                          y = spleen_mass_g, 
                          fill = lab_or_rw,
                          color = lab_or_rw,
                          group = lab_or_rw)) +
  geom_point(aes(shape = t_muris_dose),
             alpha = 0.6,
             size = 1.8) +
  geom_smooth(method = "lm",
              se = TRUE,
              alpha = 0.15,
              linewidth = 0.8) +
  facet_wrap(~ strain) + 
  scale_fill_manual(values = c("Lab" = color_purple,
                               "RW" = color_green),
                    name = expression(bold("Environment")),
                    labels = c("Lab" = "Lab",
                               "RW" = "Rewilded")) +
  scale_color_manual(values = c("Lab" = color_purple,
                               "RW" = color_green),
                    name = expression(bold("Environment")),
                    labels = c("Lab" = "Lab",
                               "RW" = "Rewilded")) +
  scale_shape_manual(values = c("Zero" = 1,
                                "Low" = 13,
                                "High" = 3),
                     name = expression(bolditalic("T. muris") ~
                                         bold("Dose"))) +
  scale_x_continuous(trans = "log10",
                     name = expression(bold("Body Mass (g)"))) +
  scale_y_continuous(trans = "log10",
                     name = expression(bold("Spleen Mass (g)"))) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("Spleen vs Body Mass")

spleen_mass_reg

plot_saver("spleen_vs_body_mass_reg.png",
           spleen_mass_reg)

```

# Spleen mass adjusted for body mass for B6 and B10. I'm splitting the strains 
to not overfit the model.

```{r}

spleen_adj_lmer <- lmer(spleen_mass_g ~
                          lab_or_rw *
                          t_muris_dose * 
                          strain +
                          mass +
                          (1 | cage_wedge),
                        data = spleen_mass_compiled_data)

tab_model(spleen_adj_lmer,
          show.stat = TRUE)
# statistic is t-value btw
summary(spleen_adj_lmer)

qqnorm(resid(spleen_adj_lmer))
qqline(resid(spleen_adj_lmer))

plot(spleen_adj_lmer)

```

# Let's add the body mass as a co-variate to see if there is any change.

```{r}

emm_adj_spleen <- emmeans(spleen_adj_lmer,
                      specs = c("t_muris_dose", "lab_or_rw"),
                      by = c("strain"),
                      type = "response")

summary(emm_adj_spleen)

pairs(emm_adj_spleen)

kable(pairs(emm_adj_spleen),
      digits = 3) %>%
  kable_styling(full_width = FALSE)

emm_adj_spleen_df <- as.data.frame(emm_adj_spleen)

emm_adj_spleen_plot <- ggplot(emm_adj_spleen_df) +
  geom_point(aes(x = lab_or_rw,
                 y = emmean,
                 fill = t_muris_dose,
                 group = t_muris_dose),
             position = position_dodge(width = 0.5),
             alpha = 1,
             size = 3,
             shape = 21,
             color = "black") +
  geom_errorbar(aes(x = lab_or_rw,
                    ymin = lower.CL,
                    ymax = upper.CL,
                    color = t_muris_dose,
                    group = t_muris_dose),
                width = 0.3,
                alpha = 0.8,
             position = position_dodge(width = 0.5)) +
  geom_point(data = spleen_mass_compiled_data,
             aes(x = lab_or_rw,
                 y = spleen_mass_g,
                 color = t_muris_dose),
             position = position_jitterdodge(jitter.width = 0.1,
                                             dodge.width = 0.5),
             alpha = 0.3,
             size = 1.5) +
  scale_y_continuous(name = expression(bold("Spleen Mass Adjusted for Body Mass (g)")),
                     expand = expansion(mult = c(0.05, 0.3))) +
  scale_x_discrete(name = expression(bold("Environment")),
                   labels = c("Lab" = "Lab",
                              "RW" = "Rewilded")) +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "Low" = color_mustard,
                               "High" = color_orange),
                    name = expression(bolditalic("T. muris")~bold("Dose"))) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "Low" = color_mustard,
                                "High" = color_orange),
                     name = expression(bolditalic("T. muris")~bold("Dose"))) +
  facet_wrap(~ strain) +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  ggtitle("Marginal Means of Adjusted Spleen Mass (95% CI)")

emm_adj_spleen_plot

plot_saver("emm_adj_spleen_plot.png",
           emm_adj_spleen_plot)

```

```{r}
lym_lmer <- lmer(log(spleen_mass_ratio) ~ 
                   log(total_lym_number) +
                   lab_or_rw * 
                   t_muris_dose + 
                   (1 | cage_wedge), 
             data = lym_number_spleen_data)

tab_model(lym_lmer, show.stat = TRUE)
# statistic is t-value btw
summary(lym_lmer)

```

# Diagnostic for the lym-mass relationship

```{r}

qqnorm(resid(lym_lmer))
qqline(resid(lym_lmer))

plot(lym_lmer)

```

# Plotting the splenocyte count and spleen mass

```{r}

spleen_mass_lym_count_plot <- ggplot(lym_number_spleen_data %>%
                                       mutate(strain = factor(strain,
                                                              levels = c("C57BL/6J", "B10.BR")),
                                              lab_or_rw = factor(lab_or_rw,
                                                                 levels = c("Lab", "RW")),
                                              t_muris_dose = factor(t_muris_dose,
                                                                    levels = c("Zero", "High"))),
                                     aes(x = spleen_mass_g,
                                         y = total_lym_number,
                                         fill = t_muris_dose)) +
  geom_point(size = 2,
             alpha = 0.8,
             shape = 21,
             color = "black") +
  geom_smooth(method = "lm",
              se = TRUE,
              fill = "grey70",
              aes(group = t_muris_dose,
                  color = t_muris_dose),
              alpha = 0.2) +
  facet_grid(lab_or_rw ~ strain,
             labeller = labeller(lab_or_rw = c("Lab" = "Lab",
                                                "RW" = "Rewilded"))) +
  scale_color_manual(values = c("Zero" = color_dark_blue,
                                "High" = color_orange),
                     name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_fill_manual(values = c("Zero" = color_dark_blue,
                               "High" = color_orange),
                    name = expression(bolditalic("T. muris") ~ bold("Dose"))) +
  scale_x_continuous(trans = "log10",
                     name = expression(bold("Spleen Mass (g)"))) +
  scale_y_continuous(trans = "log10",
                     name = expression(bold("Extracted Lymphocyte Count"))) +
  guides(fill = guide_legend(override.aes = list(shape = 21,
                                                 color = "black")),
         color = "none") +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white", 
                                        color = "black", 
                                        linewidth = 0.8),
        strip.text = element_text(size = 10, face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA, 
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5, face = "bold")) +
  labs(title = "Spleen Mass vs. Lymphocyte Count")

spleen_mass_lym_count_plot
plot_saver("spleen_mass_lymphocyte.png",
           spleen_mass_lym_count_plot,
           width = 8,
           height = 6)
```

# CBC lym count and spleen mass at final time point regression

```{r}

spleen_cbc_reg <- ggplot(spleen_cbc,
                         aes(x = lym_number,
                             y = spleen_mass_g,
                             color = t_muris_dose)) +
  geom_point(size = 1.5,
             alpha = 0.6) +
  geom_smooth(method = "lm",
              se = TRUE,
              aes(group = t_muris_dose,
                  fill = t_muris_dose),
              alpha = 0.05,
              linewidth = 0.8,
              show.legend = FALSE) +
  facet_grid(lab_or_rw ~ strain,
             labeller = labeller(lab_or_rw = c("Lab" = "Lab",
                                               "RW" = "Rewilded"))) +
  scale_color_manual(values = c("Zero" = color_purple,
                                "Low" = color_green,
                                "High" = color_orange),
                     name = expression(bolditalic("T. muris") ~
                                         bold("Dose"))) +
  scale_fill_manual(values = c("Zero" = color_purple,
                                "Low" = color_green,
                                "High" = color_orange),
                     name = expression(bolditalic("T. muris") ~
                                         bold("Dose"))) +
  scale_x_continuous(name = expression(bold("Lymphocyte Count (10"^3~"cells/"*mu*"L)")),
                     trans = "log10") +
  scale_y_continuous(name = expression(bold("Spleen Mass (g)")),
                     trans = "log10") +
  theme_minimal() +
  theme(strip.background = element_rect(fill = "white",
                                        color = "black",
                                        linewidth = 0.8),
        strip.text = element_text(size = 10,
                                  face = "bold"),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    linewidth = 0.8),
        panel.grid = element_blank(),
        axis.title.x = element_text(face = "bold"),
        axis.title.y = element_text(face = "bold"),
        axis.text = element_text(color = "black"),
        axis.ticks = element_line(color = "black",
                                  linewidth = 0.3),
        axis.ticks.length = unit(0.15, "cm"),
        plot.title = element_text(hjust = 0.5,
                                  face = "bold")) +
  ggtitle("Spleen Mass vs. CBC Lymphocyte Count")

spleen_cbc_reg

plot_saver("spleen_cbc_reg.png",
           spleen_cbc_reg)


spleen_cbc <- spleen_cbc %>%
  mutate(lab_or_rw = relevel(factor(lab_or_rw), 
                             ref = "Lab"),
         strain = relevel(factor(strain), 
                          ref = "C57BL/6J"),
         t_muris_dose = relevel(factor(t_muris_dose),
                                ref = "Zero"))

spleen_cbc_lm <- lm(spleen_mass_g ~ lym_number + 
                      lab_or_rw + 
                      strain + 
                      t_muris_dose,
                    data = spleen_cbc)

summary(spleen_cbc_lm)

tab_model(spleen_cbc_lm)

```
